### 2020-07-10 nb/analysis_in_R.Rmd
calculate ExpectedScoringTime based on Transition Matrix

---
jupyter:
  jupytext:
    formats: ipynb,Rmd
    text_representation:
      extension: .Rmd
      format_name: rmarkdown
      format_version: '1.2'
      jupytext_version: 1.4.2
  kernelspec:
    display_name: R
    language: R
    name: ir
---

# Executive Summary

# Introduction

```{r}
library(plyr) # data manipulation @HWickham2009
library(dplyr) # data manipulation @HWickham@2017
library(knitr) # dynamic report generation @YXie2012
library(igraph) # for graph/network analysis @GCsardiTNepusz2006
library(ggraph) # graph visualization in the style of ggplot @TPederson2016a
library(netrankr) # network centrality @DSchoch2015
library(ggplot2) # for plotting visualizations @HWickham2007
library(ggforce) # adds functionality to ggplot2, eg. geom_circle @TPederson2016b
library(ggrepel) # adds 'repellent' non-overlapping labels to ggplot @KSlowikowski2016
library(pheatmap) # prettier heat maps
library(RColorBrewer) # color palettes
library(docstring) # allows ?help to display function descriptions analogous to Pythonic docstrings
```

```{r}
DrawPitch <- function(lengthPitch=105, widthPitch=68, arrow=c("none", "r", "l"), theme=c("light", "dark", "grey", "grass")) {
    #' Draw regulation football pitch with penalty areas and centre circle.
    #' Adapted from https://github.com/JoGall/soccermatics

  # define colours by theme
  if(theme[1] == "grass") {
    fill1 <- "#008000"
    fill2 <- "#328422"
    colPitch <- "grey85"
    arrowCol <- "white"
    colText <- "white"
  } else if(theme[1] == "light") {
    fill1 <- "grey98"
    fill2 <- "grey98"
    colPitch <- "grey60"
    arrowCol = "black"
    colText <- "black"
  } else if(theme[1] %in% c("grey", "gray")) {    
    fill1 <- "#A3A1A3"
    fill2 <- "#A3A1A3"
    colPitch <- "white"
    arrowCol <- "white"
    colText <- "black"
  } else if(theme[1] == "dark") {  
    fill1 <- "#1C1F26"
    fill2 <- "#1C1F26"
    colPitch <- "white"
    arrowCol <- "white"
    colText <- "white"
  } else if(theme[1] == "blank") {
    fill1 <- "white"
    fill2 <- "white"
    colPitch <- "white"
    arrowCol <- "black"
    colText <- "black"
  }
  lwd <- 0.5

  border <- c(5, 5, 5, 5)

  # mowed grass lines
  lines <- (lengthPitch + border[2] + border[4]) / 13
  boxes <- data.frame(start = lines * 0:12 - border[4], end = lines * 1:13 - border[2])[seq(2, 12, 2),]

  # draw pitch
  p <- ggplot() +
    # background
    geom_rect(aes(xmin = -border[4], xmax = lengthPitch + border[2], ymin = -border[3], ymax = widthPitch + border[1]), fill = fill1) +
    # mowed pitch lines
    geom_rect(data = boxes, aes(xmin = start, xmax = end, ymin = -border[3], ymax = widthPitch + border[1]), fill = fill2) +
    # perimeter line
    geom_rect(aes(xmin = 0, xmax = lengthPitch, ymin = 0, ymax = widthPitch), fill = NA, col = colPitch, lwd = lwd) +
    # centre circle
    geom_circle(aes(x0 = lengthPitch/2, y0 = widthPitch/2, r = 9.15), col = colPitch, lwd = lwd) +
    # kick off spot
    geom_circle(aes(x0 = lengthPitch/2, y0 = widthPitch/2, r = 0.25), fill = colPitch, col = colPitch, lwd = lwd) +
    # halfway line
    geom_segment(aes(x = lengthPitch/2, y = 0, xend = lengthPitch/2, yend = widthPitch), col = colPitch, lwd = lwd) +
    # penalty arcs
    geom_arc(aes(x0= 11, y0 = widthPitch/2, r = 9.15, start = pi/2 + 0.9259284, end = pi/2 - 0.9259284), col = colPitch, lwd = lwd) +
    geom_arc(aes(x0 = lengthPitch - 11, y0 = widthPitch/2, r = 9.15, start = pi/2*3 - 0.9259284, end = pi/2*3 + 0.9259284), 
             col = colPitch, lwd = lwd) +
    # penalty areas
    geom_rect(aes(xmin = 0, xmax = 16.5, ymin = widthPitch/2 - 20.15, ymax = widthPitch/2 + 20.15), fill = NA, col = colPitch, lwd = lwd) +
    geom_rect(aes(xmin = lengthPitch - 16.5, xmax = lengthPitch, ymin = widthPitch/2 - 20.15, ymax = widthPitch/2 + 20.15), 
              fill = NA, col = colPitch, lwd = lwd) +
    # penalty spots
    geom_circle(aes(x0 = 11, y0 = widthPitch/2, r = 0.25), fill = colPitch, col = colPitch, lwd = lwd) +
    geom_circle(aes(x0 = lengthPitch - 11, y0 = widthPitch/2, r = 0.25), fill = colPitch, col = colPitch, lwd = lwd) +
    # six yard boxes
    geom_rect(aes(xmin = 0, xmax = 5.5, ymin = (widthPitch/2) - 9.16, ymax = (widthPitch/2) + 9.16), fill = NA, col = colPitch, lwd = lwd) +
    geom_rect(aes(xmin = lengthPitch - 5.5, xmax = lengthPitch, ymin = (widthPitch/2) - 9.16, ymax = (widthPitch/2) + 9.16), 
              fill = NA, col = colPitch, lwd = lwd) +
    # goals
    geom_rect(aes(xmin = -2, xmax = 0, ymin = (widthPitch/2) - 3.66, ymax = (widthPitch/2) + 3.66), fill = NA, col = colPitch, lwd = lwd) +
    geom_rect(aes(xmin = lengthPitch, xmax = lengthPitch + 2, ymin = (widthPitch/2) - 3.66, ymax = (widthPitch/2) + 3.66), 
              fill = NA, col = colPitch, lwd = lwd) +
    coord_fixed() +
    theme(rect = element_blank(), 
          line = element_blank(),
          axis.text = element_blank(),
          axis.title = element_blank())

  # add arrow
  if(arrow[1] == "r") {
    p <- p + 
      geom_segment(aes(x = 0, y = -2, xend = lengthPitch / 3, yend = -2), 
                   colour = arrowCol, size = 1.5, arrow = arrow(length = unit(0.2, "cm"), type="closed"), linejoin='mitre')
  } else if(arrow[1] == "l") {
    p <- p + 
      geom_segment(aes(x = lengthPitch, y = -2, xend = lengthPitch / 3 * 2, yend = -2), 
                   colour = arrowCol, size = 1.5, arrow = arrow(length = unit(0.2, "cm"), type="closed"), linejoin='mitre')
  }

  return(p)
}
```

# Data

```{r}
LoadAndScaleData <- function(national_league) {
    #' Load preprocessed event logs for given national league.

    events <- read.csv(paste0('../data/processed/', national_league, '.csv'))

    x_scale <- 105
    y_scale <- 68

    events$location_x <- events$location_x/100 * x_scale
    events$target_x <- events$target_x/100 * x_scale

    events$location_y <- y_scale - (events$location_y/100 * y_scale)
    events$target_y <- y_scale - (events$target_y/100 * y_scale)

    return(events)
}

```

```{r}
events <- LoadAndScaleData('England')
```

```{r}
ShowMatchEvents <- function(events, match_id, team_name, home_or_away='Home', flip=F) {
    #' Visualize events on pitch for given match and team.

    # select match
    game <- events[events$matchId==match_id,]
    goals <- game[game$subEventName=='Goal',]

    # note teams
    teams <- unique(game$team)

    # limit events to team specified by name or home/away
    if(!missing(team_name)){
        game <- game[game$team==team_name,]
        home_or_away <- unique(game$home_or_away)
    } else {
        game <- game[game$home_or_away==home_or_away,]
        team_name <- unique(game$team)
    }

    # note opposition
    opposition <- teams[teams!=team_name]

    # get score
    team_score <- table(goals$team)[team_name]
    opposition_score <- table(goals$team)[opposition]

    # flip coordinates if desired
    if (flip==T) {
        game$location_x <- 105 - game$location_x
        game$location_y <- 68 - game$location_y
        direction_of_play = 'l'
    } else { 
        direction_of_play = 'r'
    }

    # limit attention to main events
    game_events <- game[game$eventName == 'Pass'| 
                        game$eventName == 'Shot' | 
                        game$eventName == 'On the Ball' | 
                        game$eventName == 'Challenge' | 
                        game$eventName == 'Free Kick',]

    # draw pitch
    p <- DrawPitch(theme='grey', arrow=direction_of_play) + 
        geom_point(data = game_events, 
               aes(location_x , location_y, fill=eventName, shape=eventName), pch=21, alpha=0.6, size=2 ) + 
        geom_point(data=game_events[game_events$subEventName=='Goal',], 
                   aes(location_x, location_y), shape=13, size=5) +
        geom_label_repel(data=game_events[game_events$subEventName=='Goal',], 
                         aes(location_x, location_y, label = paste0(source,'(',matchPeriod,' ',time,')')), label.padding=0.1, size=2.3, alpha=1) +
        theme(legend.direction='horizontal', legend.position=c(0.5,0)) +
        scale_fill_manual(values=c("red", "black", 'white', "yellow", 'blue'), name='Event Type') +
        ggtitle(paste('Event Map:', team_name,'playing',home_or_away,'versus',opposition,'(',team_score,'-',opposition_score,')'))

    return(p)

}
```

```{r}
match_id <- '2500032'
```

```{r}
ShowMatchEvents(events, match_id, home_or_away='Home')
ShowMatchEvents(events, match_id, home_or_away='Away', flip=T)
```

```{r}
ShowPassesAndShots <- function(events, match_id, team_name, home_or_away='Home', flip=F) {
    #' Visualize passes on pitch for given match and team.

    # select match
    game <- events[events$matchId==match_id,]
    goals <- game[game$subEventName=='Goal',]

    # note teams
    teams <- unique(game$team)

    # limit events to team specified by name or home/away
    if(!missing(team_name)){
        game <- game[game$team==team_name,]
        home_or_away <- unique(game$home_or_away)
    } else {
        game <- game[game$home_or_away==home_or_away,]
        team_name <- unique(game$team)
    }

    # note opposition
    opposition <- teams[teams!=team_name]

    # get score
    team_score <- table(goals$team)[team_name]
    opposition_score <- table(goals$team)[opposition]

    # flip coordinates if desired
    if (flip==T) {
        game$location_x <- 105 - game$location_x
        game$target_x <- 105 - game$target_x
        game$location_y <- 68 - game$location_y
        game$target_y <- 68 - game$target_y
        direction_of_play = 'l'
    } else { 
        direction_of_play = 'r'
    }

    # limit attention to passes and shots
    passes <- game[game$eventName == 'Pass' & game$team == team_name,]
    shots <- game[(game$eventName=='Shot' | game$subEventName=='Goal') & game$team==team_name,]

    # draw pitch
    p <- DrawPitch(theme='grass', arrow=direction_of_play) + 
        geom_segment(data=na.exclude(passes), 
                     aes(x=location_x, y=location_y, xend=target_x, yend=target_y, color=subEventName), 
                     alpha=0.6, arrow = arrow(length = unit(0.1,"cm"))) +
        geom_label_repel(data=game[game$subEventName=='Goal',], 
                     aes(location_x, location_y, label = paste0(source,'(',matchPeriod,' ',time,')')), label.padding=0.1, size=2.3, alpha=1) + 
        geom_segment(data=na.exclude(shots), 
                 aes(x=location_x, y=location_y, xend=target_x, yend=target_y, color=subEventName), 
                 alpha=0.7, arrow = arrow(length = unit(0.1,"cm"))) +
        theme(legend.position=c(0.5,-0.01), legend.direction='horizontal',
              legend.background=element_rect(fill='#008000', linetype='solid')) +
        scale_color_manual(values=c("red", "black", 'blue', 'orange', "yellow", 'grey', 'purple', 'white','brown'), name='Pass Type') +
        ggtitle(paste('Pass Map:',team_name,'playing',home_or_away,'versus',opposition,'(',team_score,'-',opposition_score,')'))

    return(p)
} 
```

```{r}
ShowPassesAndShots(events, match_id, home_or_away='Home')
ShowPassesAndShots(events, match_id, home_or_away='Away', flip=T)
```

# Methodology

```{r}
PassNetwork <- function(events, match_id, team_name, team_colour='red', home_or_away='Home', flip=F, lower_threshold=1, high_threshold=10) {
    #' Draw Pass Map of First XI with nodes placed on mean (x,y) pitch-coordinates.

    game <- events[events$matchId == match_id,]
    goals <- game[game$subEventName=='Goal',]

    # note teams
    teams <- unique(game$team)

    # limit events to team specified by name or home/away
    if(!missing(team_name)){
        game <- game[game$team==team_name,]
        home_or_away <- unique(game$home_or_away)
    } else {
        game <- game[game$home_or_away==home_or_away,]
        team_name <- unique(game$team)
    }

    # note opposition
    opposition <- teams[teams!=team_name]

    # get score
    team_score <- table(goals$team)[team_name]
    opposition_score <- table(goals$team)[opposition]

    firstXI <- game[game$FirstXI == 'True',]
    mean_positions <- firstXI[firstXI$location_x>0 & firstXI$location_y>0 & firstXI$location_x<105 & firstXI$location_y<68,] %>% 
        group_by(team, matchId, source) %>%
          dplyr::summarise(x_mean = mean(location_x), y_mean = mean(location_y)) %>% 
          ungroup() %>%
          mutate(team = as.factor(team), id = as.factor(matchId)) %>%
          as.data.frame()

    pass_counts <- ddply(data.frame(game$source, game$target),.(game.source,game.target),nrow)
    names(pass_counts) <- c('source','target','passcount')
    step1 <- merge(mean_positions, pass_counts, by='source')
    step2 <- step1[,c(1,4,5,7,8)]
    names(step2)[2:3] <- c('source_x','source_y')
    names(mean_positions)[3] <- 'target'
    step3 <- merge(mean_positions, step2, by='target')
    team <- step3[step3$team==team_name,]

    if (flip==T) {
        team$source_x <- 105 - team$source_x
        team$x_mean <- 105 - team$x_mean
        team$source_y <- 68 - team$source_y
        team$y_mean <- 68 - team$y_mean
        mean_positions$x_mean <- 105 - mean_positions$x_mean
        mean_positions$y_mean <- 68 - mean_positions$y_mean
    }

    return (DrawPitch() +
        geom_segment(data=team[team$passcount>=lower_threshold,], size=1, colour=team_colour,
                     aes(x=source_x, y=source_y, xend=x_mean, yend=y_mean, alpha=passcount)) +
        geom_segment(data=team[team$passcount>=high_threshold,], size=1.5, colour='black',
                     aes(x=source_x, y=source_y, xend=x_mean, yend=y_mean, alpha=passcount)) +
        geom_label_repel(data = mean_positions[mean_positions$team==team_name,], 
                         aes(x_mean, y_mean, label = target), label.padding=0.1, size=2.3, alpha=1) +
        geom_point(data=team, aes(x_mean, y_mean,), fill=team_colour, colour='black', pch=21, size=3) +
        ggtitle(paste('Passing Network:',team_name,'playing',home_or_away,'versus',opposition,'(',team_score,'-',opposition_score,')')) +
        theme(legend.position=c(0.5,0.07), legend.direction='horizontal'))

}
```

```{r}
PassNetwork(events, match_id, team_colour='blue')
PassNetwork(events, match_id, home_or_away='Away', flip=T)
```

```{r}
GameGraph <- function(events, match_id, team_name, home_or_away='Home') {
    #' Return Possession Graph for given Match and Team.

    game <- events[events$matchId == match_id,]
    game <- game[game$source!='' & game$target!='nan',]

     # note teams
    teams <- unique(game$team)

    # limit events to team specified by name or home/away
    if(!missing(team_name)){
        team_game <- game[game$team==team_name,]
        home_or_away <- unique(team_game$home_or_away)
    } else {
        team_game <- game[game$home_or_away==home_or_away,]
        team_name <- as.character(unique(team_game$team))
    }

    opposition_game <- game[game$team != team_name,]
    opposition_team <- as.character(unique(opposition_game$team))

    team_nodes <- unique(c(as.character(team_game$source),as.character(team_game$target)))
    team_edges <- data.frame(team_game$source, team_game$target)
    team_graph <- graph_from_data_frame(d=team_edges, vertices=team_nodes, directed=TRUE)
    team_adj <- as.matrix(team_graph[])

    possession_seqs <- unique(team_game$possession)

    seq_start <- c()

    for (pseq in possession_seqs){
        if (sum(team_game$possession==pseq)>1){
            player <- as.character(team_game[team_game$possession==pseq,][1,]$source)
            if (player!='' & player!='nan'){
                seq_start <- append(seq_start, player)
            }
        }
    }

    start_counts <- table(as.factor(seq_start))

    for (i in (2:(length(start_counts)-1))){
        team_adj[nrow(team_adj)-1,][names(start_counts[i])] <- as.numeric(start_counts[i])
    }

    opposition_status <- as.character(unique(opposition_game$home_or_away))
    opp_to_opp <- as.numeric(summary(game[game$team != team_name,]$to_team)[opposition_status])

    opp_nodes <- unique(c(as.character(opposition_game$source),as.character(opposition_game$target)))
    opp_edges <- data.frame(opposition_game$source, opposition_game$target)
    opp_graph <- graph_from_data_frame(d=opp_edges, vertices=opp_nodes, directed=TRUE)
    opp_adj <- as.matrix(opp_graph[])

    opp_to_opp <- sum(opp_adj[,ncol(opp_adj)-1])
    team_adj[(nrow(team_adj)-1),(ncol(team_adj)-1)] <- opp_to_opp

    team_status <- unique(team_game$home_or_away)
    opposition_own_goals <- as.numeric(nrow(game[game$home_or_away==opposition_status & game$to_team==paste(team_status,'Goal'),]))
    team_adj[(nrow(team_adj)-1),(ncol(team_adj))] <- opposition_own_goals

    team_graph <- graph_from_adjacency_matrix(team_adj, mode='directed')

    return(team_graph)
}
```

```{r}
AdjacencyMatrix <- function(graph){
    #' Return Adjacency Matrix for Graph

    return(as.matrix(graph[]))
}
```

```{r}
TransitionMatrix <- function(graph){
    #' Return Transition Matrix for Graph considered as time-homogeneous Markov Process

    matrix <- as.matrix(graph[])

    for (i in 1:nrow(matrix)){
        matrix[i,] <- matrix[i,]/sum(matrix[i,])
    }

    matrix[nrow(matrix),ncol(matrix)] <- 1
    matrix[nrow(matrix),] <- rep(0, ncol(matrix))

    # treat 'Goal' as absorbtion state
    matrix[nrow(matrix),ncol(matrix)] <- 1

    return(matrix)
}
```

```{r}
ExpectedScoringTime <- function(transition_matrix){
    #' Calculate expected scoring time based on transition matrix.

    max_steps <- 10000
#     transition_matrix <- TransitionMatrix(GameGraph(events, match_id))
    hitting_times <- rep(0,(ncol(transition_matrix)-2))

    for (i in (1:(ncol(transition_matrix)-2))){

        state_probabilities <- matrix(NA,nrow=max_steps+1,ncol=ncol(transition_matrix),dimnames=list(0:max_steps,(ncol(transition_matrix)-1):0))
        vector <- rep(0,ncol(transition_matrix))

        vector[i] <- 1
        state_probabilities[1,] <- vector

        for ( kk in 1:max_steps ) {
            state_probabilities[kk+1,] <- t(transition_matrix)%*%state_probabilities[kk,]
        }

        probs <- diff(state_probabilities[,ncol(transition_matrix)])
        hitting_time <- sum(probs*seq_along(probs))

        hitting_times[i] <- hitting_time
    }

    names(hitting_times) <- names(transition_matrix[1,])[1:(nrow(transition_matrix)-2)]

    return(hitting_times)

}
```

```{r}
sort(ExpectedScoringTime(
    TransitionMatrix(
        GameGraph(events, match_id, home_or_away='Away'))

))
```

```{r}
names(transition_matrix[1,])[1:(nrow(transition_matrix)-2)]
```

```{r}
max_steps <- 10000
transition_matrix <- TransitionMatrix(GameGraph(events, match_id))
state_probabilities <- matrix(NA,nrow=max_steps+1,ncol=ncol(transition_matrix),dimnames=list(0:max_steps,(ncol(transition_matrix)-1):0))
vector <- rep(0,ncol(transition_matrix))
vector[6] <- 1
state_probabilities[1,] <- vector

for ( kk in 1:max_steps ) {
    state_probabilities[kk+1,] <- t(transition_matrix)%*%state_probabilities[kk,]
}

probs <- diff(state_probabilities[,ncol(transition_matrix)])
sum(probs*seq_along(probs))
```

```{r}
transition_matrix
```

```{r}
MatrixHeatMap <- function(matrix, number_format='%.d', title='Matrix Heatmap'){
    #' Show Matrix as HeatMap

    hm <- pheatmap(matrix,color=brewer.pal(15,'YlGnBu'),cluster_rows=F,cluster_cols=F,legend=F,
                    display_numbers=T,number_format=number_format,fontsize_number=9,angle_col='315',
                    main=title)

    return(hm)
}
```

```{r}
MatrixHeatMap(
    AdjacencyMatrix(g)[1:11,1:11]

)
```

```{r}
MatrixHeatMap(
    AdjacencyMatrix(
        GameGraph(events, match_id, home_or_away='Away'))[1:11,1:11]
)
```

```{r}
MatrixHeatMap(
    TransitionMatrix(
        GameGraph(events, match_id, home_or_away='Home')), 
    number_format='%.2f',
    title=''
)
```

```{r}
MatrixHeatMap(
    TransitionMatrix(
        GameGraph(events, match_id, home_or_away='Away')), 
    number_format='%.2f',
    title=''
)
```

```{r}
GoalSeqs <- function(events, match_id) {
    match <- events[events$matchId==match_id,]
    goals <- match[match$subEventName=='Goal',]
    goal_seqs <- unique(goals$possession)   
    return(goal_seqs)
}
```

```{r}
TabulateSequence <- function(events, possession_sequence) {
    # Return a markdown table starting from the row before a possession sequence begins, and continuing until it ends.

    start <- min(events[events$possession == possession_sequence,]$X)
    stop <- max(events[events$possession == possession_sequence,]$X) + 1
    table <- events[start:stop,]
    df <- data.frame(table$matchPeriod, table$time, table$team, table$source, table$subEventName)
    names(df) <- c('Half', 'Time', 'Team', 'Event Description', 'Player')

    return(kable(df))
}
```

```{r}
sequence <- '2500032-91-Liv-0'
```

```{r}
TabulateSequence(events, sequence)
```

```{r}
SequenceOnPitch <- function(events, possession_sequence){
    #' Draw Possession Sequence on Pitch

    data <- events[events$possession == possession_sequence,]

    p <- DrawPitch(theme='grass') + 
        geom_label_repel(data = data, aes(location_x, location_y, label = paste(source,time)), label.padding=0.1, size=2.3, alpha=1) +
        geom_curve(data = data, aes(x = location_x, xend = target_x, y = location_y, yend = target_y*.99, col = X), 
                   show.legend=FALSE, size=1, alpha = 1) +
        geom_point(data = data, aes(location_x , location_y, shape=eventName), size=2) + 
        ggtitle(paste0(nrow(data),'-part ', unique(data$team)[1], ' Possession Sequence ending with ', 
                       data[nrow(data),]$subEventName, ' by ', data[nrow(data),]$source))  +
        theme(legend.position=c(0.5,0.08), legend.direction='horizontal', legend.title=element_blank())

    return(p)
}
```

```{r}
SequenceOnPitch(events, sequence)
```

youtube clip
<img src="../figures/youtube_clip.gif" width='700'>

```{r}
SequenceGraph <- function(events, possession_sequence){
    #' Return Graph of Possession Sequence

    data <- events[events$possession == possession_sequence,]
    data <- data[data$source!='' & data$target!='nan',]
    nodes <- unique(c(as.character(data$source),as.character(data$target)))
    edges <- data.frame(data$source, data$target)
    g <- graph_from_data_frame(d=edges, vertices=nodes, directed=TRUE)
    g$id <- possession_sequence
    g$data <- data
    g$team <- as.character(unique(data$team))

    return(g)
}
```

```{r}
VisualizeGraph <- function(possession_sequence_graph) {
    #' Visualize Possession Sequence as Linear Graph with Looping Edges

    g <- possession_sequence_graph
    data <- possession_sequence_graph$data

    visualization <- ggraph(g, 'linear') + 
        geom_edge_arc(aes(color=data$eventName), 
                      arrow=arrow(length=unit(4,'mm')), 
                      fold=F,
                     width=1) +
        geom_edge_loop(aes(color=data$eventName),
                      width=1) +
        geom_node_point(color='black', 
                        size=2,
                        alpha=0.5) +
        geom_node_text(aes(label = name), 
                        repel=T, 
                        angle=90, hjust=2, ) +
        scale_edge_colour_manual(
            values=c('indianred3', 'wheat4', 'grey', 'grey30',
                    'red','blue','green','orange','purple','brown','pink'
                    ),
            name='Event Type') +
        theme_void()

    return(visualization)

}
```

```{r}
g <- SequenceGraph(events, sequence)
```

```{r}
VisualizeGraph(g)
```

```{r}
GetGraphMetrics <- function(possession_graph) {
    #' Return basic metrics for given Graph

    g <- possession_graph
    path_metrics <- c(g$id, g$team, length(V(g)), length(E(g)), 
                      diameter(g), mean_distance(g), edge_density(g, loops=T), 
                      girth(g)$girth, transitivity(g), authority_score(g)$value, hub_score(g)$value)
    names(path_metrics) <- c('Label', 'Team', 'Nodes', 'Edges', 'Diameter', 'Mean_Distance', 
                             'Edge_Density', 'Girth', 'Transitivity', 'Authority_Score', 'Hub_Score')

    return(path_metrics)
}
```

```{r}
GetNodeMetrics <- function(possession_graph) {
    #' Return metrics for each Node of given Graph

    g <- possession_graph
    df <- data.frame(page_rank(g)$vector)
    names(df) <- 'pagerank'
    df$in_degree <- degree(g, mode='in')
    df$out_degree <- degree(g, mode='out')
    df$eccentricity <- eccentricity(g)
    df$closeness <- closeness(g)
    df$betweenness <- betweenness(g)
    df$norm_betweenness <- betweenness(g, normalized=T)
    df$alpha_centrality <- alpha_centrality(g)
    df$power_centrality <- power_centrality(g)
    df$eigen_centrality <- eigen_centrality(g)$vector
    df$authority <- authority_score(g)$vector
    df$hub <- hub_score(g)$vector

    return(df)
}
```

```{r}
GetEdgeMetrics <- function(possession_graph) {
    #' Return metrics for each Edge of given Graph

    g <- possession_graph
    df <- data.frame(as_edgelist(g))
    names(df) <- c('Source','Target')
    df$sequence <- 1:length(E(g))
    df$edge_betweenness <- edge_betweenness(g)

    return(df)
}
```

```{r}
GetGraphMetrics(g)
```

```{r}
GetNodeMetrics(match_graph)
```

```{r}
GetEdgeMetrics(g)
```

```{r}
sequence2 <- '2499858-18-Liv-5'
```

```{r}
TabulateSequence(sequence2)
```

```{r}
SequenceOnPitch(sequence2)
```

<img src='../figures/mane_vs_stoke.gif'>

```{r}
g2 <- SequenceGraph(sequence2)
VisualizeGraph(g2)
```

```{r}
GetGraphMetrics(g2)
```

```{r}
sequence3 <-  '2499889-18-Liv-0'
```

```{r}
SequenceOnPitch(sequence3)
```

```{r}
g3 <- SequenceGraph(sequence3)
VisualizeGraph(g3)
```

```{r}
GetGraphMetrics(g3)
```

<img src="../figures/youtube_clip2.gif">

```{r}

```

### 2020-07-10 nb/analysis_in_R.Rmd
start making LeagueResults function

@@
 events <- LoadAndScaleData('England')
+```
+
+```{r}
+LeagueResults <- function(league){
+
+    teams <- sort(unique(league$team))
+
+    home_score <- matrix(0, nrow=20, ncol=20)
+    colnames(home_score) <- teams
+    rownames(home_score) <- teams
+
+    away_score <- home_score
+
+    league_table <- matrix(0, nrow=20, ncol=8)
+    rownames(league_table) <- teams
+    colnames(league_table) <- c('POS','PTS','W','D','L','GF','GA','GD')
+
+    return(list(league_table, home_score, away_score))
+}
+```
+
+```{r}
+
+```
+
+```{r}
+v <- LeagueResults(events)
+```
+
+```{r}
+hs <- v[[2]]
+```
+
+```{r}
+league <- events
+```
+
+```{r}
+teams <- sort(unique(league$team))
+
+home_score <- matrix(0, nrow=20, ncol=20)
+colnames(home_score) <- teams
+rownames(home_score) <- teams
+
+away_score <- home_score
+
+league_table <- matrix(0, nrow=20, ncol=8)
+rownames(league_table) <- teams
+colnames(league_table) <- c('POS','PTS','W','D','L','GF','GA','GD')
+```
+
+```{r}
+for m in unique(league$matchId){
+
+    match <- league[league$matchId==m,]
+    home_team <- unique(match[match$home_or_away=='Home']$team)
+    away_team <- unique(match[match$home_or_away=='Away']$team)
+    home_goals <- nrow(match[match$to_team=='Home Goal',])
+    away_goals <- nrow(match[match$to_team=='Away Goal',])
+
+    home_score[home_team, away_team] <- home_goals
+    away_score[home_team, away_team] <- away_goals
+
+    league_table[]
+
+        league_table[team]
+    }
+
+}
+```
+
+```{r}
+league_table
+```
+
+```{r}
+
+```
+
+```{r}
+for (match in unique(events$matchId)){
+
+    goals <- 0
+
+}
+```
+
+```{r}
+
+```
+
+```{r}
+
+```
+
+```{r}
+mId <- unique(events$matchId)[1]
+```
+
+```{r}
+m <- events[events$matchId==mId,]
+```
+
+```{r}
+home_goals <- m[m$to_team=='Home Goal',]
+```
+
+```{r}
+away_goals <- m[m$to_team=='Away Goal',]
+```
+
+```{r}
+home_goals <- matrix(0, nrow=20, ncol=20)
+```
+
+```{r}
+away_goals <- matrix(0, nrow=20, ncol=20)
+```
+
+```{r}
+colnames(home_goals) <- teams
+```
+
+```{r}
+rownames(home_goals) <- teams
+```
+
+```{r}
+home
+```
+
+```{r}
+home_goals
+```
+
+```{r}
+teams <- sort(unique(events$team))
 ```
@@
 ```{r}
-sort(ExpectedScoringTime(
+x_score_time <- ExpectedScoringTime(
     TransitionMatrix(
-        GameGraph(events, match_id, home_or_away='Away'))
+        GameGraph(events, match_id, home_or_away='Home'))
 
-))
+)
+```
+
+```{r}
+x_score_time/mean(x_score_time)
+```
+
+```{r}
+x_score_time/mean(x_score_time)
 ```

### 2020-07-10 nb/preprocessing.Rmd
start making LeagueResults function

---
jupyter:
  jupytext:
    formats: ipynb,Rmd
    text_representation:
      extension: .Rmd
      format_name: rmarkdown
      format_version: '1.2'
      jupytext_version: 1.4.2
  kernelspec:
    display_name: Python 3
    language: python
    name: python3
---

```{python}
import pandas as pd
import matplotlib.pyplot as plt
import json
import numpy as np

import warnings
warnings.filterwarnings('ignore')
```

```{python}
tag_df = pd.read_csv('../data/raw/tagcodes.csv')
tag_dict = dict(zip(tag_df.Tag, tag_df.Description))
```

```{python}
def load_json(file):
    "Load raw JSON data."
    with open(f'../data/raw/{file}.json','r') as f:
        return json.loads(f.read().replace('\\\\','\\').encode('utf-8'))
```

```{python}
def preprocess(national_league):
    "Create processed CSV from raw JSON for WyScout national league event-logs."

    # load JSON event-logs
    events = load_json(f'events_{national_league}')

    # create dataframe from JSON event-logs
    ev_df = pd.DataFrame(events)

    # decode tags
    for i in range(6):
        ev_df[f'tag{i}'] = ev_df['tags'].str[i].str['id'].map(tag_dict).str.lower()

    # decode player ids
    players = load_json('players')
    players_dict = dict(zip(pd.DataFrame(players).wyId,pd.DataFrame(players).shortName))
    ev_df['player'] = ev_df['playerId'].map(players_dict)

    # decode team ids
    teams = load_json('teams')
    teams_dict = dict(zip(pd.DataFrame(teams).wyId,pd.DataFrame(teams).name))
    ev_df['team'] = ev_df['teamId'].map(teams_dict)

    # unpack location and target
    ev_df['location'] = ev_df['positions'].str[0]
    ev_df['target'] = ev_df['positions'].str[1]

    # create simpler dataframe
    e_df = ev_df[['matchId','eventSec','matchPeriod','team','location','target','player','eventName','subEventName','tag0','tag1','tag2','tag3','tag4','tag5']]
    e_df['notes'] = e_df[['tag0','tag1','tag2','tag3','tag4','tag5']].values.tolist()
    e_df['tags'] = e_df['notes'].astype(str)
    e_df['accurate'] = e_df.tags.str.contains('accurate')
    e_df['accurate'] = np.where(e_df.tags.str.contains('not accurate'), 0, e_df.accurate)
    e_df.drop(columns=['tag0','tag1','tag2','tag3','tag4','tag5','notes'],inplace=True)

    # label in_play sequences
    play = 0
    plays = []
    next_action = []
    match_play = []
    matchId = e_df.iloc[0].matchId
    for i in range(len(e_df)):
        if i>1:
            row_minus2 = row_minus1
        if i>0:
            row_minus1 = row
        row = e_df.iloc[i]
        if i == 0:
            row_minus1 = row
            row_minus2 = row

        if row.eventName == 'Free Kick':
            # when the ball goes 'out of play', the game resumes with a 'free kick' (here includes Throw-Ins)
            play += 1
        elif 'goal' in row_minus2.tags and 'goal' in row_minus1.tags:
            # (there are generally two events marked goal -- the shot and the attempted save)
            play +=1

        if row.matchId == matchId:
            pass
        else:
            play = 0
            matchId = row.matchId
        plays.append(play)
        match_play.append(f'{matchId}-{play}')

    e_df['in_play'] = pd.Series(plays)
    e_df['match_play'] = pd.Series(match_play)

    # label possession sequences
    ps_sequence = np.full(len(e_df),'0000000000000000000000000000000000')
    played_to = np.full(len(e_df), '                                    ')
    to_team = np.full(len(e_df), '                                       ')

    for match in e_df.matchId.unique():
        match_df = e_df.loc[e_df.matchId == match]
        for team in match_df.team.unique():
            for play in match_df.in_play.unique():
                team_play_df = match_df.loc[match_df.in_play == play].loc[match_df.team == team]
                possession_sequence = 0
                for i in range(len(team_play_df)):
                    if i>0:
                        previous_row = row

                    row = team_play_df.iloc[i]

                    if i > 0:
                        if row.location != target:
                            possession_sequence += 1
                            played_to[previous_row.name] = 'Opposition' 
                            to_team[previous_row.name] = 'Opposition'
                        else:
                            played_to[previous_row.name] = row.player
                            to_team[previous_row.name] = row.team
                    team_name = team.split(' ')
                    team_abbr = ''
                    for word in team_name:
                        team_abbr = team_abbr + word[:3]
                    ps_sequence[row.name] = f'{match}-{play}-{team_abbr}-{possession_sequence}'
                    target = row.target

    e_df['possession'] = ps_sequence            
    e_df['played_to'] = played_to
    e_df['to_team'] = to_team

    e_df['played_to'] = np.where(e_df['played_to'] == '                                    ', 'Opposition', e_df['played_to'])
    e_df['to_team'] = np.where(e_df['to_team'] == '                                       ', 'Opposition', e_df['to_team'])

    # unpack x and y coorinates
    for foo in ('location','target'):
        for goo in ('x','y'):
            e_df[f'{foo}_{goo}'] = e_df[foo].str[goo]

    # change target coordinates of goal to inside goal
    e_df['target_x'] = np.where((e_df.tags.str.contains("'goal'")) & (e_df.eventName != 'Save attempt'),
                                100, e_df.target_x )
    e_df['target_y'] = np.where((e_df.tags.str.contains("'goal'")) & (e_df.eventName != 'Save attempt'),
                                50, e_df.target_y )

    e_df['subEventName'] = np.where((e_df.tags.str.contains("'goal'")) & (e_df.eventName != 'Save attempt'),
                                'Goal', e_df.subEventName )

    e_df['eventName'] = np.where(e_df['eventName'] == 'Others on the ball', 'On the Ball', e_df['eventName'])

    e_df['eventName'] = np.where(e_df['eventName'] == 'Duel', 'Challenge', e_df['eventName'])

    e_df['played_to'] = np.where(e_df.subEventName == 'Goal', 'Goal', e_df.played_to)
    e_df['to_team'] = np.where(e_df.subEventName == 'Goal', 'Goal', e_df.to_team)

    # change location coordinates of goal kick
    e_df['location_x'] = np.where((e_df.subEventName == 'Goal kick'), 5, e_df.location_x)
    e_df['location_y'] = np.where((e_df.subEventName == 'Goal kick'), 50, e_df.location_y)

    # add time in minutes and seconds
    e_df['minute'] = e_df.eventSec//60
    e_df.minute = e_df.minute.astype(int)
    e_df['seconds'] = e_df.eventSec % 60
    e_df.seconds = e_df.seconds.astype(int)
    e_df['time'] = e_df.minute.astype(str).str.zfill(2) + ':' + e_df.seconds.astype(str).str.zfill(2)
    e_df.drop(columns=['minute','seconds'],inplace=True)

    e_df.drop(columns=['location','target'], inplace=True)
    e_df.rename(columns={'player':'source','played_to':'target'},inplace=True)

    # include starting lineup data
    matches = load_json(f'matches_{national_league}')
    match_df = pd.DataFrame(matches)
    match_lineups = {}
    for match in match_df.wyId.unique():
        match_lineups[match] = {}
        for team in match_df.loc[match_df.wyId == match].teamsData.values[0].keys():
            match_lineups[match][teams_dict[int(team)]] = list(pd.Series(pd.DataFrame(match_df.loc[
                match_df.wyId == match].teamsData.str[str(team)].str['formation'].str['lineup'].values[0]).playerId.values).map(players_dict))

    def firstXI(row):
        return row.source in match_lineups[row.matchId][row.team]

    e_df['FirstXI'] = e_df.apply(firstXI, axis=1)

    # include home_or_away data
    h_a_map = {}
    for match in match_df.wyId.unique():
        h_a_map[match] = {}
        for team in match_df.loc[match_df.wyId == match].teamsData.values[0].keys():
            status = match_df.loc[match_df.wyId==match].teamsData.str[str(team)].str['side'].values[0].capitalize()
            h_a_map[match][team] = status

    def home_or_away(row):
        return h_a_map[row.matchId][str(row.teamId)]

    e_df['home_or_away'] = ev_df.apply(home_or_away, axis=1)
    e_df['to_team'] = np.where(e_df.to_team == e_df.team, e_df.home_or_away, e_df.to_team)
    opposition = {'Home':'Away','Away':'Home'}
    e_df['to_team'] = np.where(e_df.to_team == 'Opposition', e_df.home_or_away.map(opposition), e_df.to_team)
    e_df['to_team'] = np.where(e_df.tags.str.contains('own goal'), 'Own Goal', e_df.to_team)
    whose_goal = {'Home':'Home Goal', 'Away':'Away Goal'}
    e_df['to_team'] = np.where(e_df.to_team == 'Goal', e_df.home_or_away.map(whose_goal), e_df.to_team)
    whose_own_goal = {'Home':'Away Goal', 'Away':'Home Goal'}
    e_df['to_team'] = np.where(e_df.to_team == 'Own Goal', e_df.home_or_away.map(whose_own_goal), e_df.to_team)

    e_df.drop(columns=['tags','in_play'],inplace=True)

    e_df.to_csv(f'../data/processed/{national_league}FootballLeague2017-18.csv')
```

```{python}
national_leagues = ('England','France','Germany','Italy','Spain')
```

```{python}
for league in national_leagues:
    preprocess(league)
```

```{python}

```

```{python}
national_league='England'

# load JSON event-logs
events = load_json(f'events_{national_league}')

# create dataframe from JSON event-logs
ev_df = pd.DataFrame(events)

# decode tags
for i in range(6):
    ev_df[f'tag{i}'] = ev_df['tags'].str[i].str['id'].map(tag_dict).str.lower()

# decode player ids
players = load_json('players')
players_dict = dict(zip(pd.DataFrame(players).wyId,pd.DataFrame(players).shortName))
ev_df['player'] = ev_df['playerId'].map(players_dict)

# decode team ids
teams = load_json('teams')
teams_dict = dict(zip(pd.DataFrame(teams).wyId,pd.DataFrame(teams).name))
ev_df['team'] = ev_df['teamId'].map(teams_dict)

# unpack location and target
ev_df['location'] = ev_df['positions'].str[0]
ev_df['target'] = ev_df['positions'].str[1]

# create simpler dataframe
e_df = ev_df[['matchId','eventSec','matchPeriod','team','location','target','player','eventName','subEventName','tag0','tag1','tag2','tag3','tag4','tag5']]
e_df['notes'] = e_df[['tag0','tag1','tag2','tag3','tag4','tag5']].values.tolist()
e_df['tags'] = e_df['notes'].astype(str)
e_df['accurate'] = e_df.tags.str.contains('accurate')
e_df['accurate'] = np.where(e_df.tags.str.contains('not accurate'), 0, e_df.accurate)
e_df.drop(columns=['tag0','tag1','tag2','tag3','tag4','tag5','notes'],inplace=True)

# label in_play sequences
play = 0
plays = []
next_action = []
match_play = []
matchId = e_df.iloc[0].matchId
for i in range(len(e_df)):
    if i>1:
        row_minus2 = row_minus1
    if i>0:
        row_minus1 = row
    row = e_df.iloc[i]
    if i == 0:
        row_minus1 = row
        row_minus2 = row

    if row.eventName == 'Free Kick':
        # when the ball goes 'out of play', the game resumes with a 'free kick' (here includes Throw-Ins)
        play += 1
    elif 'goal' in row_minus2.tags and 'goal' in row_minus1.tags:
        # (there are generally two events marked goal -- the shot and the attempted save)
        play +=1

    if row.matchId == matchId:
        pass
    else:
        play = 0
        matchId = row.matchId
    plays.append(play)
    match_play.append(f'{matchId}-{play}')

e_df['in_play'] = pd.Series(plays)
e_df['match_play'] = pd.Series(match_play)

# label possession sequences
ps_sequence = np.full(len(e_df),'0000000000000000000000000000000000')
played_to = np.full(len(e_df), '                                    ')
to_team = np.full(len(e_df), '                                       ')

for match in e_df.matchId.unique():
    match_df = e_df.loc[e_df.matchId == match]
    for team in match_df.team.unique():
        for play in match_df.in_play.unique():
            team_play_df = match_df.loc[match_df.in_play == play].loc[match_df.team == team]
            possession_sequence = 0
            for i in range(len(team_play_df)):
                if i>0:
                    previous_row = row

                row = team_play_df.iloc[i]

                if i > 0:
                    if row.location != target:
                        possession_sequence += 1
                        played_to[previous_row.name] = 'Opposition' 
                        to_team[previous_row.name] = 'Opposition'
                    else:
                        played_to[previous_row.name] = row.player
                        to_team[previous_row.name] = row.team
                team_name = team.split(' ')
                team_abbr = ''
                for word in team_name:
                    team_abbr = team_abbr + word[:3]
                ps_sequence[row.name] = f'{match}-{play}-{team_abbr}-{possession_sequence}'
                target = row.target

e_df['possession'] = ps_sequence            
e_df['played_to'] = played_to
e_df['to_team'] = to_team

e_df['played_to'] = np.where(e_df['played_to'] == '                                    ', 'Opposition', e_df['played_to'])
e_df['to_team'] = np.where(e_df['to_team'] == '                                       ', 'Opposition', e_df['to_team'])

# unpack x and y coorinates
for foo in ('location','target'):
    for goo in ('x','y'):
        e_df[f'{foo}_{goo}'] = e_df[foo].str[goo]

# change target coordinates of goal to inside goal
e_df['target_x'] = np.where((e_df.tags.str.contains("'goal'")) & (e_df.eventName != 'Save attempt'),
                            100, e_df.target_x )
e_df['target_y'] = np.where((e_df.tags.str.contains("'goal'")) & (e_df.eventName != 'Save attempt'),
                            50, e_df.target_y )

e_df['subEventName'] = np.where((e_df.tags.str.contains("'goal'")) & (e_df.eventName != 'Save attempt'),
                            'Goal', e_df.subEventName )

e_df['eventName'] = np.where(e_df['eventName'] == 'Others on the ball', 'On the Ball', e_df['eventName'])

e_df['eventName'] = np.where(e_df['eventName'] == 'Duel', 'Challenge', e_df['eventName'])

e_df['played_to'] = np.where(e_df.subEventName == 'Goal', 'Goal', e_df.played_to)
e_df['to_team'] = np.where(e_df.subEventName == 'Goal', 'Goal', e_df.to_team)

# change location coordinates of goal kick
e_df['location_x'] = np.where((e_df.subEventName == 'Goal kick'), 5, e_df.location_x)
e_df['location_y'] = np.where((e_df.subEventName == 'Goal kick'), 50, e_df.location_y)

# add time in minutes and seconds
e_df['minute'] = e_df.eventSec//60
e_df.minute = e_df.minute.astype(int)
e_df['seconds'] = e_df.eventSec % 60
e_df.seconds = e_df.seconds.astype(int)
e_df['time'] = e_df.minute.astype(str).str.zfill(2) + ':' + e_df.seconds.astype(str).str.zfill(2)
e_df.drop(columns=['minute','seconds'],inplace=True)

e_df.drop(columns=['location','target'], inplace=True)
e_df.rename(columns={'player':'source','played_to':'target'},inplace=True)

# include starting lineup data
matches = load_json(f'matches_{national_league}')
match_df = pd.DataFrame(matches)
match_lineups = {}
for match in match_df.wyId.unique():
    match_lineups[match] = {}
    for team in match_df.loc[match_df.wyId == match].teamsData.values[0].keys():
        match_lineups[match][teams_dict[int(team)]] = list(pd.Series(pd.DataFrame(match_df.loc[
            match_df.wyId == match].teamsData.str[str(team)].str['formation'].str['lineup'].values[0]).playerId.values).map(players_dict))

def firstXI(row):
    return row.source in match_lineups[row.matchId][row.team]

e_df['FirstXI'] = e_df.apply(firstXI, axis=1)

# include home_or_away data
h_a_map = {}
for match in match_df.wyId.unique():
    h_a_map[match] = {}
    for team in match_df.loc[match_df.wyId == match].teamsData.values[0].keys():
        status = match_df.loc[match_df.wyId==match].teamsData.str[str(team)].str['side'].values[0].capitalize()
        h_a_map[match][team] = status

def home_or_away(row):
    return h_a_map[row.matchId][str(row.teamId)]

e_df['home_or_away'] = ev_df.apply(home_or_away, axis=1)
e_df['to_team'] = np.where(e_df.to_team == e_df.team, e_df.home_or_away, e_df.to_team)
opposition = {'Home':'Away','Away':'Home'}
e_df['to_team'] = np.where(e_df.to_team == 'Opposition', e_df.home_or_away.map(opposition), e_df.to_team)
e_df['to_team'] = np.where(e_df.tags.str.contains('own goal'), 'Own Goal', e_df.to_team)
whose_goal = {'Home':'Home Goal', 'Away':'Away Goal'}
e_df['to_team'] = np.where(e_df.to_team == 'Goal', e_df.home_or_away.map(whose_goal), e_df.to_team)
whose_own_goal = {'Home':'Away Goal', 'Away':'Home Goal'}
e_df['to_team'] = np.where(e_df.to_team == 'Own Goal', e_df.home_or_away.map(whose_own_goal), e_df.to_team)

e_df.drop(columns=['tags','in_play'],inplace=True)

```

### 2020-07-10 nb/analysis_in_R.Rmd
function works but points totals aren't right for some reason

@@
 
+    for (m in unique(league$matchId)){
+
+        match <- league[league$matchId==m,]
+        home_team <- unique(match[match$home_or_away=='Home',]$team)
+        away_team <- unique(match[match$home_or_away=='Away',]$team)
+        home_goals <- nrow(match[match$to_team=='Home Goal',])
+        away_goals <- nrow(match[match$to_team=='Away Goal',])
+
+        home_score[home_team, away_team] <- home_goals
+        away_score[home_team, away_team] <- away_goals
+
+        if (home_goals > away_goals){
+            league_table[home_team,'W'] <- league_table[home_team,'W'] + 1
+            league_table[away_team,'L'] <- league_table[away_team,'L'] + 1
+        } else if (away_goals > home_goals){
+            league_table[home_team,'L'] <- league_table[home_team,'L'] + 1
+            league_table[away_team,'W'] <- league_table[away_team,'W'] + 1    
+        } else if (home_goals == away_goals) {
+            league_table[home_team,'D'] <- league_table[home_team,'D'] + 1
+            league_table[away_team,'D'] <- league_table[away_team,'D'] + 1
+        }
+
+        league_table[home_team,'GF'] <- league_table[home_team,'GF'] + home_goals
+        league_table[away_team,'GF'] <- league_table[away_team,'GF'] + away_goals
+        league_table[home_team,'GA'] <- league_table[home_team,'GA'] + away_goals
+        league_table[away_team,'GA'] <- league_table[away_team,'GA'] + home_goals
+
+        for (team in c(home_team, away_team)){
+            league_table[team,'GD'] <- league_table[team,'GF'] - league_table[team,'GA']
+            league_table[team,'PTS'] <- 3*league_table[team,'W'] + 1*league_table[team,'D']
+        }
+
+    }
     return(list(league_table, home_score, away_score))
 }
+```
+
+```{r}
+results <- LeagueResults(events)
+```
+
+```{r}
+league_table <- data.frame(results[[1]])
+```
+
+```{r}
+league_table[order(-league_table$PTS),]
+```
+
+```{r}
+results[[2]]
+```
+
+```{r}
+results[[3]]
+```
+
+```{r}
+
+```
+
+```{r}
+[order(league_table[,1])]
+```
+
+```{r}
+table
+```
+
+```{r}
+
 ```
@@
 ```{r}
-for m in unique(league$matchId){
+for (m in unique(league$matchId)){
 
     match <- league[league$matchId==m,]
-    home_team <- unique(match[match$home_or_away=='Home']$team)
-    away_team <- unique(match[match$home_or_away=='Away']$team)
+    home_team <- unique(match[match$home_or_away=='Home',]$team)
+    away_team <- unique(match[match$home_or_away=='Away',]$team)
     home_goals <- nrow(match[match$to_team=='Home Goal',])
@@
 
-    league_table[]
+    if (home_goals > away_goals){
+        league_table[home_team,'W'] <- league_table[home_team,'W'] + 1
+        league_table[away_team,'L'] <- league_table[away_team,'L'] + 1
+    } else if (away_goals > home_goals){
+        league_table[home_team,'L'] <- league_table[home_team,'L'] + 1
+        league_table[away_team,'W'] <- league_table[away_team,'W'] + 1    
+    } else if (home_goals == away_goals) {
+        league_table[home_team,'D'] <- league_table[home_team,'D'] + 1
+        league_table[away_team,'D'] <- league_table[away_team,'D'] + 1
+    }
 
-        league_table[team]
+    league_table[home_team,'GF'] <- league_table[home_team,'GF'] + home_goals
+    league_table[away_team,'GF'] <- league_table[away_team,'GF'] + away_goals
+    league_table[home_team,'GA'] <- league_table[home_team,'GA'] + away_goals
+    league_table[away_team,'GA'] <- league_table[away_team,'GA'] + home_goals
+
+    for (team in c(home_team, away_team)){
+        league_table[team,'GD'] <- league_table[team,'GF'] - league_table[team,'GA']
+        league_table[team,'PTS'] <- 3*league_table[team,'W'] + 1*league_table[team,'D']
     }
@@
 league_table
+```
+
+```{r}
+
+```
+
+```{r}
+
+```
+
+```{r}
+
+```
+
+```{r}
+
+```
+
+```{r}
+league_table['Arsenal',]
 ```

### 2020-07-11 nb/analysis_in_R.Rmd
correct de bruyne goal data

@@
 LeagueResults <- function(league){
+    #' Calculate results and table for given national league.
 
@@
 ```{r}
-results[[2]]
-```
-
-```{r}
-results[[3]]
-```
-
-```{r}
-
-```
-
-```{r}
-[order(league_table[,1])]
-```
-
-```{r}
-table
-```
-
-```{r}
-
-```
-
-```{r}
-
-```
-
-```{r}
-v <- LeagueResults(events)
-```
-
-```{r}
-hs <- v[[2]]
-```
-
-```{r}
-league <- events
-```
-
-```{r}
-teams <- sort(unique(league$team))
-
-home_score <- matrix(0, nrow=20, ncol=20)
-colnames(home_score) <- teams
-rownames(home_score) <- teams
-
-away_score <- home_score
-
-league_table <- matrix(0, nrow=20, ncol=8)
-rownames(league_table) <- teams
-colnames(league_table) <- c('POS','PTS','W','D','L','GF','GA','GD')
-```
-
-```{r}
-for (m in unique(league$matchId)){
-
-    match <- league[league$matchId==m,]
-    home_team <- unique(match[match$home_or_away=='Home',]$team)
-    away_team <- unique(match[match$home_or_away=='Away',]$team)
-    home_goals <- nrow(match[match$to_team=='Home Goal',])
-    away_goals <- nrow(match[match$to_team=='Away Goal',])
-
-    home_score[home_team, away_team] <- home_goals
-    away_score[home_team, away_team] <- away_goals
-
-    if (home_goals > away_goals){
-        league_table[home_team,'W'] <- league_table[home_team,'W'] + 1
-        league_table[away_team,'L'] <- league_table[away_team,'L'] + 1
-    } else if (away_goals > home_goals){
-        league_table[home_team,'L'] <- league_table[home_team,'L'] + 1
-        league_table[away_team,'W'] <- league_table[away_team,'W'] + 1    
-    } else if (home_goals == away_goals) {
-        league_table[home_team,'D'] <- league_table[home_team,'D'] + 1
-        league_table[away_team,'D'] <- league_table[away_team,'D'] + 1
-    }
-
-    league_table[home_team,'GF'] <- league_table[home_team,'GF'] + home_goals
-    league_table[away_team,'GF'] <- league_table[away_team,'GF'] + away_goals
-    league_table[home_team,'GA'] <- league_table[home_team,'GA'] + away_goals
-    league_table[away_team,'GA'] <- league_table[away_team,'GA'] + home_goals
-
-    for (team in c(home_team, away_team)){
-        league_table[team,'GD'] <- league_table[team,'GF'] - league_table[team,'GA']
-        league_table[team,'PTS'] <- 3*league_table[team,'W'] + 1*league_table[team,'D']
-    }
-
-}
-```
-
-```{r}
-league_table
-```
-
-```{r}
 
@@
 
-```
-
-```{r}
-
-```
-
-```{r}
-league_table['Arsenal',]
-```
-
-```{r}
-
-```
-
-```{r}
-for (match in unique(events$matchId)){
-
-    goals <- 0
-
-}
-```
-
-```{r}
-
-```
-
-```{r}
-
-```
-
-```{r}
-mId <- unique(events$matchId)[1]
-```
-
-```{r}
-m <- events[events$matchId==mId,]
-```
-
-```{r}
-home_goals <- m[m$to_team=='Home Goal',]
-```
-
-```{r}
-away_goals <- m[m$to_team=='Away Goal',]
-```
-
-```{r}
-home_goals <- matrix(0, nrow=20, ncol=20)
-```
-
-```{r}
-away_goals <- matrix(0, nrow=20, ncol=20)
-```
-
-```{r}
-colnames(home_goals) <- teams
-```
-
-```{r}
-rownames(home_goals) <- teams
-```
-
-```{r}
-home
-```
-
-```{r}
-home_goals
-```
-
-```{r}
-teams <- sort(unique(events$team))
 ```

### 2020-07-11 nb/preprocessing.Rmd
correct de bruyne goal data

@@
 ```{python}
+# Correct data for De Bruyne's goal vs Chelsea (not an inaccurate pass!)
 
+england_events = pd.read_csv('../data/processed/England.csv')
+
+england_events.at[102730,'eventName'] = 'Shot'
+england_events.at[102730,'subEventName'] = 'Goal'
+england_events.at[102730,'accurate'] = 1
+england_events.at[102730,'to_team'] = 'Away Goal'
+england_events.at[102730,'target'] = 'Goal'
+england_events.at[102730,'target_y'] = 50
+
+england_events.to_csv('../data/processed/England.csv',index=False)
 ```
-
-```{python}
-national_league='England'
-
-# load JSON event-logs
-events = load_json(f'events_{national_league}')
-
-# create dataframe from JSON event-logs
-ev_df = pd.DataFrame(events)
-
-# decode tags
-for i in range(6):
-    ev_df[f'tag{i}'] = ev_df['tags'].str[i].str['id'].map(tag_dict).str.lower()
-
-# decode player ids
-players = load_json('players')
-players_dict = dict(zip(pd.DataFrame(players).wyId,pd.DataFrame(players).shortName))
-ev_df['player'] = ev_df['playerId'].map(players_dict)
-
-# decode team ids
-teams = load_json('teams')
-teams_dict = dict(zip(pd.DataFrame(teams).wyId,pd.DataFrame(teams).name))
-ev_df['team'] = ev_df['teamId'].map(teams_dict)
-
-# unpack location and target
-ev_df['location'] = ev_df['positions'].str[0]
-ev_df['target'] = ev_df['positions'].str[1]
-
-# create simpler dataframe
-e_df = ev_df[['matchId','eventSec','matchPeriod','team','location','target','player','eventName','subEventName','tag0','tag1','tag2','tag3','tag4','tag5']]
-e_df['notes'] = e_df[['tag0','tag1','tag2','tag3','tag4','tag5']].values.tolist()
-e_df['tags'] = e_df['notes'].astype(str)
-e_df['accurate'] = e_df.tags.str.contains('accurate')
-e_df['accurate'] = np.where(e_df.tags.str.contains('not accurate'), 0, e_df.accurate)
-e_df.drop(columns=['tag0','tag1','tag2','tag3','tag4','tag5','notes'],inplace=True)
-
-# label in_play sequences
-play = 0
-plays = []
-next_action = []
-match_play = []
-matchId = e_df.iloc[0].matchId
-for i in range(len(e_df)):
-    if i>1:
-        row_minus2 = row_minus1
-    if i>0:
-        row_minus1 = row
-    row = e_df.iloc[i]
-    if i == 0:
-        row_minus1 = row
-        row_minus2 = row
-
-    if row.eventName == 'Free Kick':
-        # when the ball goes 'out of play', the game resumes with a 'free kick' (here includes Throw-Ins)
-        play += 1
-    elif 'goal' in row_minus2.tags and 'goal' in row_minus1.tags:
-        # (there are generally two events marked goal -- the shot and the attempted save)
-        play +=1
-
-    if row.matchId == matchId:
-        pass
-    else:
-        play = 0
-        matchId = row.matchId
-    plays.append(play)
-    match_play.append(f'{matchId}-{play}')
-
-e_df['in_play'] = pd.Series(plays)
-e_df['match_play'] = pd.Series(match_play)
-
-# label possession sequences
-ps_sequence = np.full(len(e_df),'0000000000000000000000000000000000')
-played_to = np.full(len(e_df), '                                    ')
-to_team = np.full(len(e_df), '                                       ')
-
-for match in e_df.matchId.unique():
-    match_df = e_df.loc[e_df.matchId == match]
-    for team in match_df.team.unique():
-        for play in match_df.in_play.unique():
-            team_play_df = match_df.loc[match_df.in_play == play].loc[match_df.team == team]
-            possession_sequence = 0
-            for i in range(len(team_play_df)):
-                if i>0:
-                    previous_row = row
-
-                row = team_play_df.iloc[i]
-
-                if i > 0:
-                    if row.location != target:
-                        possession_sequence += 1
-                        played_to[previous_row.name] = 'Opposition' 
-                        to_team[previous_row.name] = 'Opposition'
-                    else:
-                        played_to[previous_row.name] = row.player
-                        to_team[previous_row.name] = row.team
-                team_name = team.split(' ')
-                team_abbr = ''
-                for word in team_name:
-                    team_abbr = team_abbr + word[:3]
-                ps_sequence[row.name] = f'{match}-{play}-{team_abbr}-{possession_sequence}'
-                target = row.target
-
-e_df['possession'] = ps_sequence            
-e_df['played_to'] = played_to
-e_df['to_team'] = to_team
-
-e_df['played_to'] = np.where(e_df['played_to'] == '                                    ', 'Opposition', e_df['played_to'])
-e_df['to_team'] = np.where(e_df['to_team'] == '                                       ', 'Opposition', e_df['to_team'])
-
-# unpack x and y coorinates
-for foo in ('location','target'):
-    for goo in ('x','y'):
-        e_df[f'{foo}_{goo}'] = e_df[foo].str[goo]
-
-# change target coordinates of goal to inside goal
-e_df['target_x'] = np.where((e_df.tags.str.contains("'goal'")) & (e_df.eventName != 'Save attempt'),
-                            100, e_df.target_x )
-e_df['target_y'] = np.where((e_df.tags.str.contains("'goal'")) & (e_df.eventName != 'Save attempt'),
-                            50, e_df.target_y )
-
-e_df['subEventName'] = np.where((e_df.tags.str.contains("'goal'")) & (e_df.eventName != 'Save attempt'),
-                            'Goal', e_df.subEventName )
-
-e_df['eventName'] = np.where(e_df['eventName'] == 'Others on the ball', 'On the Ball', e_df['eventName'])
-
-e_df['eventName'] = np.where(e_df['eventName'] == 'Duel', 'Challenge', e_df['eventName'])
-
-e_df['played_to'] = np.where(e_df.subEventName == 'Goal', 'Goal', e_df.played_to)
-e_df['to_team'] = np.where(e_df.subEventName == 'Goal', 'Goal', e_df.to_team)
-
-# change location coordinates of goal kick
-e_df['location_x'] = np.where((e_df.subEventName == 'Goal kick'), 5, e_df.location_x)
-e_df['location_y'] = np.where((e_df.subEventName == 'Goal kick'), 50, e_df.location_y)
-
-# add time in minutes and seconds
-e_df['minute'] = e_df.eventSec//60
-e_df.minute = e_df.minute.astype(int)
-e_df['seconds'] = e_df.eventSec % 60
-e_df.seconds = e_df.seconds.astype(int)
-e_df['time'] = e_df.minute.astype(str).str.zfill(2) + ':' + e_df.seconds.astype(str).str.zfill(2)
-e_df.drop(columns=['minute','seconds'],inplace=True)
-
-e_df.drop(columns=['location','target'], inplace=True)
-e_df.rename(columns={'player':'source','played_to':'target'},inplace=True)
-
-# include starting lineup data
-matches = load_json(f'matches_{national_league}')
-match_df = pd.DataFrame(matches)
-match_lineups = {}
-for match in match_df.wyId.unique():
-    match_lineups[match] = {}
-    for team in match_df.loc[match_df.wyId == match].teamsData.values[0].keys():
-        match_lineups[match][teams_dict[int(team)]] = list(pd.Series(pd.DataFrame(match_df.loc[
-            match_df.wyId == match].teamsData.str[str(team)].str['formation'].str['lineup'].values[0]).playerId.values).map(players_dict))
-
-def firstXI(row):
-    return row.source in match_lineups[row.matchId][row.team]
-
-e_df['FirstXI'] = e_df.apply(firstXI, axis=1)
-
-# include home_or_away data
-h_a_map = {}
-for match in match_df.wyId.unique():
-    h_a_map[match] = {}
-    for team in match_df.loc[match_df.wyId == match].teamsData.values[0].keys():
-        status = match_df.loc[match_df.wyId==match].teamsData.str[str(team)].str['side'].values[0].capitalize()
-        h_a_map[match][team] = status
-
-def home_or_away(row):
-    return h_a_map[row.matchId][str(row.teamId)]
-
-e_df['home_or_away'] = ev_df.apply(home_or_away, axis=1)
-e_df['to_team'] = np.where(e_df.to_team == e_df.team, e_df.home_or_away, e_df.to_team)
-opposition = {'Home':'Away','Away':'Home'}
-e_df['to_team'] = np.where(e_df.to_team == 'Opposition', e_df.home_or_away.map(opposition), e_df.to_team)
-e_df['to_team'] = np.where(e_df.tags.str.contains('own goal'), 'Own Goal', e_df.to_team)
-whose_goal = {'Home':'Home Goal', 'Away':'Away Goal'}
-e_df['to_team'] = np.where(e_df.to_team == 'Goal', e_df.home_or_away.map(whose_goal), e_df.to_team)
-whose_own_goal = {'Home':'Away Goal', 'Away':'Home Goal'}
-e_df['to_team'] = np.where(e_df.to_team == 'Own Goal', e_df.home_or_away.map(whose_own_goal), e_df.to_team)
-
-e_df.drop(columns=['tags','in_play'],inplace=True)
-
-```

### 2020-07-13 nb/analysis_in_R.Rmd
define MarkovExpectedGoals

@@
 ```{r}
+events[1:10,]
+```
+
+```{r}
 LeagueResults <- function(league){
@@
 
-    home_score <- matrix(0, nrow=20, ncol=20)
+    home_score <- matrix(0, nrow=length(teams), ncol=length(teams))
     colnames(home_score) <- teams
@@
 
-    league_table <- matrix(0, nrow=20, ncol=8)
+    league_table <- matrix(0, nrow=length(teams), ncol=9)
     rownames(league_table) <- teams
-    colnames(league_table) <- c('POS','PTS','W','D','L','GF','GA','GD')
+    colnames(league_table) <- c('Pos','Pld','W','D','L','GF','GA','GD','Pts')
 
@@
             league_table[team,'GD'] <- league_table[team,'GF'] - league_table[team,'GA']
-            league_table[team,'PTS'] <- 3*league_table[team,'W'] + 1*league_table[team,'D']
+            league_table[team,'Pts'] <- 3*league_table[team,'W'] + 1*league_table[team,'D']
+            league_table[team,'Pld'] <- league_table[team,'Pld'] + 1
         }
@@
     }
-    return(list(league_table, home_score, away_score))
+
+    league_table <- league_table[order(-league_table[,'Pts']),]
+
+    league_table[,'Pos'] <- 1:20
+
+    return(league_table)
 }
@@
 ```{r}
-results <- LeagueResults(events)
+prem <- LeagueResults(events)
 ```
@@
 ```{r}
-league_table <- data.frame(results[[1]])
+pl <- as.data.frame(prem)
 ```
@@
 ```{r}
-league_table[order(-league_table$PTS),]
+pl['Team'] <- rownames(pl)
 ```
@@
 ```{r}
+pl2 <- as.data.frame(pl[,c(10,2:9)])
+```
+
+```{r}
+rownames(pl2) <- c()
+```
+
+```{r}
+pl2[1:5,]
+```
+
+```{r}
+prem[1:5,]
+```
+
+```{r}
+NationalResults <- function(nation) {
+
+    data <- LoadAndScaleData(nation)
+    results <- LeagueResults(data)
+    return(results)
+}
 
@@
 ```{r}
-
+correct_res <- read.csv('../data/league_results/England17-18.csv')
 ```
@@
 ```{r}
+correct_res
+```
 
+```{r}
+france <- NationalResults('France')
+germany <- NationalResults('Germany')
+italy <- NationalResults('Italy')
+spain <- NationalResults('Spain')
+```
+
+```{r}
+premiership <- NationalResults('England')
 ```
@@
 ```{r}
+#away vs Crystal Palace
 match_id <- '2500032'
+```
+
+```{r}
+# vs man city
+match_id <- '2499943'
 ```
@@
 ```{r}
-PassNetwork(events, match_id, team_colour='blue')
-PassNetwork(events, match_id, home_or_away='Away', flip=T)
+PassNetwork(events, match_id, team_colour='red')
+PassNetwork(events, match_id, team_colour='blue', home_or_away='Away', flip=T)
+```
+
+```{r}
+2499943
+```
+
+```{r}
+
 ```
@@
     game <- events[events$matchId == match_id,]
-    game <- game[game$source!='' & game$target!='nan',]
+    game <- game[game$source!='' & game$target!='' & game$target!='nan',]
 
@@
     team_graph <- graph_from_adjacency_matrix(team_adj, mode='directed')
+
+    team_graph$team <- team_name
+    team_graph$opposition <- opposition_team
+    team_graph$status <- home_or_away
 
@@
 ```{r}
-x_score_time <- ExpectedScoringTime(
-    TransitionMatrix(
-        GameGraph(events, match_id, home_or_away='Home'))
+SeasonAdjacencyMatrix <- function(events, team) {
 
-)
+    players <- unique(events[events$team==team & events$source!='' & events$target!='',]$source)
+    n <- length(players) + 2
+    squad_adj <- matrix(0,nrow=n,ncol=n)
+    rownames(squad_adj) <- c(as.character(players), 'Opposition', 'Goal')
+    colnames(squad_adj) <- rownames(squad_adj)
+
+    games <- unique(events[events$team==team,]$matchId)
+
+    for (game in games){
+
+        # get graph and adjacency matrix for game
+        game_graph <- GameGraph(events, game, team)
+        game_adj <- AdjacencyMatrix(game_graph)
+
+        # add values of game adjacency matrix to full-season matrix
+        for (source in rownames(game_adj)){
+            for (target in colnames(game_adj)){
+                squad_adj[source,target] <- squad_adj[source,target] + game_adj[source,target]
+            }
+        }
+    }
+
+    return(squad_adj)
+}
 ```
@@
 ```{r}
-x_score_time/mean(x_score_time)
+MarkovExpectedGoals <- function(events, season_adjacency_matrix) {
+
+    season_graph <- graph_from_adjacency_matrix(season_adjacency_matrix)
+    season_transition <- TransitionMatrix(season_graph)
+    seasonXST <- ExpectedScoringTime(season_transition)
+    avg_events_per_game <- nrow(events)/380
+    mXG <- sort(avg_events_per_game/seasonXST, decreasing=T)
+
+    return(mXG)
+}
 ```
@@
 ```{r}
-x_score_time/mean(x_score_time)
+GoalProbabilities <- function(season_adjacency_matrix) {
+
+    season_graph <- graph_from_adjacency_matrix(season_adjacency_matrix)
+    season_transition <- TransitionMatrix(season_graph)
+    goal_probs <- sort(season_transition[,'Goal'],T)[2:nrow(season_transition)] * 100
+
+    return(goal_probs)
+}
 ```
@@
 ```{r}
-names(transition_matrix[1,])[1:(nrow(transition_matrix)-2)]
+liverpool_adj <- SeasonAdjacencyMatrix(events,'Liverpool')
+
 ```
@@
 ```{r}
-max_steps <- 10000
-transition_matrix <- TransitionMatrix(GameGraph(events, match_id))
-state_probabilities <- matrix(NA,nrow=max_steps+1,ncol=ncol(transition_matrix),dimnames=list(0:max_steps,(ncol(transition_matrix)-1):0))
-vector <- rep(0,ncol(transition_matrix))
-vector[6] <- 1
-state_probabilities[1,] <- vector
-
-for ( kk in 1:max_steps ) {
-    state_probabilities[kk+1,] <- t(transition_matrix)%*%state_probabilities[kk,]
-}
-
-probs <- diff(state_probabilities[,ncol(transition_matrix)])
-sum(probs*seq_along(probs))
+mXG <- MarkovExpectedGoals(events, liverpool_adj)
 ```
@@
 ```{r}
-transition_matrix
+mXG
+```
+
+```{r}
+GoalProbabilities(liverpool_adj)
+```
+
+```{r}
+sort(liverpool_adj[,'Goal'],T)
+```
+
+```{r}
+
+```
+
+```{r}
+
+```
+
+```{r}
+
+```
+
+```{r}
+g <- GameGraph(events, match_id, home_or_away='Home')
 ```

### 2020-07-14 nb/analysis_in_R.Rmd
get mXG for all leagues

@@
 }
+
+```
+
+```{r}
 
@@
 
-        state_probabilities <- matrix(NA,nrow=max_steps+1,ncol=ncol(transition_matrix),dimnames=list(0:max_steps,(ncol(transition_matrix)-1):0))
+        state_probabilities <- matrix(NA,nrow=max_steps+1,ncol=ncol(transition_matrix),
+                                      dimnames=list(0:max_steps,(ncol(transition_matrix)-1):0))
         vector <- rep(0,ncol(transition_matrix))
@@
     seasonXST <- ExpectedScoringTime(season_transition)
+    seasonMXG <- 1/seasonXST
+    normalize <- seasonMXG/(mean(seasonMXG))
     avg_events_per_game <- nrow(events)/380
-    mXG <- sort(avg_events_per_game/seasonXST, decreasing=T)
+    scale <- normalize * avg_events_per_game - avg_events_per_game
+    sorted <- sort(scale, decreasing=T)
 
-    return(mXG)
+    return(sorted)
 }
@@
 ```{r}
-liverpool_adj <- SeasonAdjacencyMatrix(events,'Liverpool')
+leagueMXG <- function(events){
 
+    all_mXG <- c()
+    teams <- unique(events$team)
+
+    for (team in teams){
+
+        season_adjacency_matrix <- SeasonAdjacencyMatrix(events, team)
+        mXG <- MarkovExpectedGoals(events, season_adjacency_matrix)
+
+        all_mXG <- append(all_mXG, mXG)
+
+    }
+
+    return(all_mXG)
+}
 ```
@@
 ```{r}
-mXG <- MarkovExpectedGoals(events, liverpool_adj)
+mXG <- c()
+leagues = c('England','France','Germany','Italy','Spain')
+for (league in leagues){
+    events <- LoadAndScaleData(league)
+    mXG <- append(mXG, leagueMXG(events))
+}
 ```
@@
 ```{r}
-mXG
-```
-
-```{r}
-GoalProbabilities(liverpool_adj)
-```
-
-```{r}
-sort(liverpool_adj[,'Goal'],T)
-```
-
-```{r}
-
+sort(mXG,T)
 ```

### 2020-07-15 rmd/draft1.Rmd
defining terms...

---
jupyter:
  jupytext:
    formats: ipynb,Rmd
    text_representation:
      extension: .Rmd
      format_name: rmarkdown
      format_version: '1.2'
      jupytext_version: 1.4.2
  kernelspec:
    display_name: R
    language: R
    name: ir
---

# Random Walks along Social Networks: The Markovian Movement of a Stochastic Soccerball

## Executive Summary

## Introduction

## Data

## Methodology

We begin by defining our terms.

A *social* network is a network involving *persons*. I take as axiomatic the proposition that human beings are, individually and sometimes collectively, persons; we will not here entertain the question of whether such a designation should be extended to animal, divine, or robotic agents.

A *network* is a graph in which some or all of the nodes, edges, and graph are *objects* with *attributes* besides those given in the description of the network as a graph.

A *graph* is an ordered tuple $G = (V,E)$, consisting of a set of *nodes* (or *vertices*) $V = \{v_{i}\}$, and a set of *edges* $E = \{e_{ij}\}$, where the edge $e_{ij}$ is the ordered pair $\{i,j\}$ representing some connection from $v_{i}$ to $v_{j}$.

Two nodes connected by an edge are said to be *adjacent* to each other.

An edge $e_{ii}$ connecting a node $v_{i}$ to itself is called a *loop*.

A graph is called *undirected* if $e_{ij} \iff e_{ji}$ (otherwise it is a *directed* graph), and *simple* if the edges are distinct and unrepeated (otherwise it is a *multigraph*).

A graph is *weighted* if there exists some function $w: E \mapsto \mathbb{R}$ assigning some weight value to each edge.

An unweighted multigraph can thus be viewed as a weighted simple graph by defining $w(e_{ij}) = |\{e \in E: e = e_{ij}\}| \in \mathbb{N_{0}}$, so that the weight of an edge is the number of times it is repeated.

A weighted simple graph can be uniquely described by an *adjacency matrix*:
$$
M_{ij} = \begin{cases}
    \begin{array}{ll}
         w(e_{ij}), & e_{ij} \in E \\
       0, & otherwise
    \end{array}
\end{cases}
$$

A *walk* of length L is a sequence of adjacent (but not necessarily distinct) nodes $(v_{0},...,v_{L})$.

We also need some definitions from *probability theory* and *statistics*.

A *sample space* $ \Omega = \{\omega_{i}\}$ is a set of possible *outcomes* of an *observation*.

An *event* $F$ is some subset of $\Omega$, and the *event space* $\mathcal{F}$ is the set of subsets of $\Omega$. In particular, $\Omega \subset \mathcal{F}$.

A *probability* $\mathbb{P}$ is a function $\mathbb{P}: \mathcal{F} \mapsto [0,1] \in \mathbb{R}$, such that $\sum\limits_{\omega_{i} \in \Omega}\mathbb{P}(\omega_{i}) = 1$. We can also talk about the *conditional probability* of an event $G$ given another event $F$, which we write $\mathbb{P}(G | F)$.

A *random variable* is a function $X: \Omega \mapsto S$. If $S \subseteq \mathbb{R}$ then we also have the *expectation* $\mathbb{E}[X]$ and *variance* $\mathrm{Var}[X]$, but it is acceptable to have an $S \not\subset \mathbb{R}$.

A *stochastic process* is an indexed set (that is, a *sequence*) of random variables, where the index commonly refers to points in time $t \in T$, which may be considered discretely (so $t \in \mathbb{N}_{0}$ ) or continuously (so $t \in \mathbb{R}_{ \ge 0}$ ). The range $S$ of a stochastic process $\{X(t,\omega):t\in T,\omega \in \Omega\}$ is called its *state space*, and the value $X_{t} = X(t)$ is its state at time $t$.

A stochastic process has the *Markov property* if $\forall t \ge 0, \mathbb{P}(X_{t+1}=s| X_{0},...,X_{t}) = \mathbb{P}(X_{t+1}=s| X_{t})$; this means that future states are dependent only on the present state, regardless of the previous history of past states. Such a process can be called a Markov process, and we call the probability $\mathbb{P}(X_{t+1}=s_{j}|X_{t}=s_{i})$ the *transition probability* $p_{ij}$ .

A Markov process is *time-homogenous* if the transition probabilities stay constant through time, ie. $\mathbb{P}(X_{n+1}=s_{i}|X_{n}=s_{j}) = \mathbb{P}(X_{m+1}=x|X_{m}=y) \forall m,n \in T$. If $S$ is countable, then the probability distribution of such a process can be described by its initial *distribution vector* $\pi_{0}$ and its *transition matrix* $P$, where each $i$th row is made up of the transition probabilities $p_{ij}$ and $\sum\limits_{j} p_{ij} = 1 \forall i$. Specifically, after $k$ transitions we have the distribution $\pi_{k} = \pi_{0}P^{k}$.

If a Markov process develops in discrete time and takes a values from a discrete state space, then we call it a *Markov chain* and can straightforwardly draw it as a weighted simple graph by treating its transition matrix as an adjacency matrix. Conversely, we can transform an adjacency matrix $M$ into a transition matrix $P$ by defining the elements $p_{ij} := \frac{m_{ij} }{\sum\limits_{j}m_{ij}}$, so that each row sums to 1 as required. We can then picture the development of a Markov process as a *random walk* on the graph defined by its transition probabilities, noting that since the transition probabilities are not uniform this walk is *biased*.

The *hitting time* $\tau_{R}$ of some state(s) $R \subset S$ is given by $\tau_{R} = min\{ t \ge 0 \colon X_{t} \in R \}$.

A discrete-time *stochastic process* $\{X_{n} : n \ge 0 \}$ on a countable set $S$ is a collection of $S$-valued random variables defined on a probability space $(\Omega,\mathcal{F},\mathbb{P})$. The $\mathbb{P}$ is a probability measure on $\mathcal{F}$, the field of subsets of the sample-space $\Omega$. The set $S$ is the *state space* of the process, and the value $X_{n} \in S$ is the *state* of the process at *time* $n$.

A stochastic process has the *Markov property* if $\forall n \ge 0$,

[@RSerfozo2009]

https://stats.stackexchange.com/questions/64167/how-to-define-sample-space-for-discrete-random-variable

https://math.stackexchange.com/questions/2042801/state-space-and-sample-space-difference/2322791

sample space, state space, event space

sample space is the set of all possible outcomes,

event space, the set of sets of outcomes

probability mass function $p: \Omega \mapsto [0,1]$ such that $\sum\limits_{\omega \in \Omega} p(\omega) = 1$

$(\Omega,\mathcal{F},\mathbb{P})$

"a network can be defined as a graph in which nodes and/or edges have attributes"

how to define social

For our purposes, the

## Results

## Conclusion

## Bibliography

### 2020-07-16 rmd/analysis_in_R.Rmd
writing draft

--- 2020-07-11 nb/analysis_in_R.Rmd
@@
   jupytext:
-    formats: ipynb,Rmd
     text_representation:
@@
 
-# Executive Summary
-
-# Introduction
-
 ```{r}
-library(plyr) # data manipulation @HWickham2009
-library(dplyr) # data manipulation @HWickham@2017
-library(knitr) # dynamic report generation @YXie2012
-library(igraph) # for graph/network analysis @GCsardiTNepusz2006
-library(ggraph) # graph visualization in the style of ggplot @TPederson2016a
-library(netrankr) # network centrality @DSchoch2015
-library(ggplot2) # for plotting visualizations @HWickham2007
-library(ggforce) # adds functionality to ggplot2, eg. geom_circle @TPederson2016b
-library(ggrepel) # adds 'repellent' non-overlapping labels to ggplot @KSlowikowski2016
-library(pheatmap) # prettier heat maps
-library(RColorBrewer) # color palettes
-library(docstring) # allows ?help to display function descriptions analogous to Pythonic docstrings
+library(plyr)
+library(dplyr)
+library(ggplot2)
+library(MASS)
+library(cowplot)
+library(forcats)
+library(ggforce)
+library(ggrepel)
+library(scales)
+library(tidyr)
+library(xts)
+library(zoo)
 ```
@@
 ```{r}
-DrawPitch <- function(lengthPitch=105, widthPitch=68, arrow=c("none", "r", "l"), theme=c("light", "dark", "grey", "grass")) {
-    #' Draw regulation football pitch with penalty areas and centre circle.
-    #' Adapted from https://github.com/JoGall/soccermatics
+library(ggraph)
+```
+
+```{r}
+library(igraph)
+```
+
+```{r}
+soccerPitch <- function(lengthPitch = 105, widthPitch = 68, arrow = c("none", "r", "l"), title = NULL, subtitle = NULL, theme = c("light", "dark", "grey", "grass")) {
 
@@
 
+  # outer border (t,r,b,l)
+#   border <- c(10, 6, 5, 6)
   border <- c(5, 5, 5, 5)
@@
     geom_arc(aes(x0= 11, y0 = widthPitch/2, r = 9.15, start = pi/2 + 0.9259284, end = pi/2 - 0.9259284), col = colPitch, lwd = lwd) +
-    geom_arc(aes(x0 = lengthPitch - 11, y0 = widthPitch/2, r = 9.15, start = pi/2*3 - 0.9259284, end = pi/2*3 + 0.9259284), 
-             col = colPitch, lwd = lwd) +
+    geom_arc(aes(x0 = lengthPitch - 11, y0 = widthPitch/2, r = 9.15, start = pi/2*3 - 0.9259284, end = pi/2*3 + 0.9259284), col = colPitch, lwd = lwd) +
     # penalty areas
     geom_rect(aes(xmin = 0, xmax = 16.5, ymin = widthPitch/2 - 20.15, ymax = widthPitch/2 + 20.15), fill = NA, col = colPitch, lwd = lwd) +
-    geom_rect(aes(xmin = lengthPitch - 16.5, xmax = lengthPitch, ymin = widthPitch/2 - 20.15, ymax = widthPitch/2 + 20.15), 
-              fill = NA, col = colPitch, lwd = lwd) +
+    geom_rect(aes(xmin = lengthPitch - 16.5, xmax = lengthPitch, ymin = widthPitch/2 - 20.15, ymax = widthPitch/2 + 20.15), fill = NA, col = colPitch, lwd = lwd) +
     # penalty spots
@@
     geom_rect(aes(xmin = 0, xmax = 5.5, ymin = (widthPitch/2) - 9.16, ymax = (widthPitch/2) + 9.16), fill = NA, col = colPitch, lwd = lwd) +
-    geom_rect(aes(xmin = lengthPitch - 5.5, xmax = lengthPitch, ymin = (widthPitch/2) - 9.16, ymax = (widthPitch/2) + 9.16), 
-              fill = NA, col = colPitch, lwd = lwd) +
+    geom_rect(aes(xmin = lengthPitch - 5.5, xmax = lengthPitch, ymin = (widthPitch/2) - 9.16, ymax = (widthPitch/2) + 9.16), fill = NA, col = colPitch, lwd = lwd) +
     # goals
     geom_rect(aes(xmin = -2, xmax = 0, ymin = (widthPitch/2) - 3.66, ymax = (widthPitch/2) + 3.66), fill = NA, col = colPitch, lwd = lwd) +
-    geom_rect(aes(xmin = lengthPitch, xmax = lengthPitch + 2, ymin = (widthPitch/2) - 3.66, ymax = (widthPitch/2) + 3.66), 
-              fill = NA, col = colPitch, lwd = lwd) +
+    geom_rect(aes(xmin = lengthPitch, xmax = lengthPitch + 2, ymin = (widthPitch/2) - 3.66, ymax = (widthPitch/2) + 3.66), fill = NA, col = colPitch, lwd = lwd) +
     coord_fixed() +
@@
 
-  # add arrow
-  if(arrow[1] == "r") {
-    p <- p + 
-      geom_segment(aes(x = 0, y = -2, xend = lengthPitch / 3, yend = -2), 
-                   colour = arrowCol, size = 1.5, arrow = arrow(length = unit(0.2, "cm"), type="closed"), linejoin='mitre')
-  } else if(arrow[1] == "l") {
-    p <- p + 
-      geom_segment(aes(x = lengthPitch, y = -2, xend = lengthPitch / 3 * 2, yend = -2), 
-                   colour = arrowCol, size = 1.5, arrow = arrow(length = unit(0.2, "cm"), type="closed"), linejoin='mitre')
-  }
+#   # add arrow
+#   if(arrow[1] == "r") {
+#     p <- p + 
+#       geom_segment(aes(x = 0, y = -2, xend = lengthPitch / 3, yend = -2), colour = arrowCol, size = 1.5, arrow = arrow(length = unit(0.2, "cm"), type="closed"), linejoin='mitre')
+#   } else if(arrow[1] == "l") {
+#     p <- p + 
+#       geom_segment(aes(x = lengthPitch, y = -2, xend = lengthPitch / 3 * 2, yend = -2), colour = arrowCol, size = 1.5, arrow = arrow(length = unit(0.2, "cm"), type="closed"), linejoin='mitre')
+#   }
+
+#   # add title and/or subtitle
+#   theme_buffer <- ifelse(theme[1] == "light", 0, 4)
+#   if(!is.null(title) & !is.null(subtitle)) {
+#     p <- p +
+#       cowplot::draw_text(title, 
+#                          x = 0, y = widthPitch + 9, hjust = 0, vjust = 1,
+#                          size = 15, fontface = 'bold', col = colText) +
+#       cowplot::draw_text(subtitle, 
+#                          x = 0, y = widthPitch + 4.5, hjust = 0, vjust = 1,
+#                          size = 13, col = colText) +
+#       theme(plot.margin = unit(c(-0.525,-0.9,-0.7,-0.9), "cm"))
+#   } else if(!is.null(title) & is.null(subtitle)) {
+#     p <- p +
+#       cowplot::draw_text(title, 
+#                          x = 0, y = widthPitch + 4.5, hjust = 0, vjust = 1,
+#                          size = 15, fontface = 'bold', col = colText) +
+#       theme(plot.margin = unit(c(-0.9,-0.9,-0.7,-0.9), "cm"))
+#   } else if(is.null(title) & !is.null(subtitle)) {
+#     p <- p +
+#       cowplot::draw_text(subtitle, 
+#                          x = 0, y = widthPitch + 4.5, hjust = 0, vjust = 1,
+#                          size = 13, col = colText) +
+#       theme(plot.margin = unit(c(-0.9,-0.9,-0.7,-0.9), "cm"))
+#   } else if(is.null(title) & is.null(subtitle)){
+#     p <- p +
+#       theme(plot.margin = unit(c(-1.2,-0.9,-0.7,-0.9), "cm"))
+#   }
 
   return(p)
+
 }
@@
 
-# Data
-
 ```{r}
-LoadAndScaleData <- function(national_league) {
-    #' Load preprocessed event logs for given national league.
-
-    events <- read.csv(paste0('../data/processed/', national_league, '.csv'))
+events <- read.csv('../data/processed_events_England.csv')
 
@@
 events$target_y <- y_scale - (events$target_y/100 * y_scale)
+```
 
-    return(events)
-}
+```{r}
+game <- events[events$matchId == '2500032',]
+```
+
+```{r}
+game_events <- game[game$eventName == 'Pass'| game$eventName == 'Shot' | game$eventName == 'Duel' | game$eventName == 'Free Kick',]
+```
+
+```{r}
+soccerPitch(theme='grey') + 
+    geom_point(data = game_events, 
+           aes(location_x , location_y, fill=eventName), pch=21, alpha=0.5, size=2 ) + 
+    theme(legend.position='bottom', legend.title = element_blank()) +
+    scale_fill_manual(values=c("red", "black", "blue", 'white'))
+```
+
+```{r}
+id <-  '2500032-91-Liv-0'
+# id <- best_goal_sequences[1]
+data <- events[events$possession == id,]
+# 534 is brighton vs bournemouth
+# data <- events[events$possession == '2499759-46-Bri&HovAlb-0',]
+# data <- events[events$possession == '2499889-18-Liv-0',]
+# data <- events[events$possession == '2499774-85-ManCit-0',] #2500065-23-ManCit-1',]#2499913-77-Liv-0',]
+
+soccerPitch(theme='grass') + geom_point(data = data,
+               aes(location_x , location_y)) + 
+    geom_label_repel(data = data, aes(location_x, location_y, label = source), label.padding=0.1, size=2.3, alpha=1) +
+ geom_curve(data = data,
+              aes(x = location_x, xend = target_x, y = location_y, yend = target_y*.99, col = X), show.legend=FALSE, size=1, alpha = 1)
 
@@
 
-```{r}
-events <- LoadAndScaleData('England')
-```
+<img src="../figures/youtube_clip.gif">
 
 ```{r}
-LeagueResults <- function(league){
-    #' Calculate results and table for given national league.
-
-    teams <- sort(unique(league$team))
-
-    home_score <- matrix(0, nrow=20, ncol=20)
-    colnames(home_score) <- teams
-    rownames(home_score) <- teams
-
-    away_score <- home_score
-
-    league_table <- matrix(0, nrow=20, ncol=8)
-    rownames(league_table) <- teams
-    colnames(league_table) <- c('POS','PTS','W','D','L','GF','GA','GD')
-
-    for (m in unique(league$matchId)){
-
-        match <- league[league$matchId==m,]
-        home_team <- unique(match[match$home_or_away=='Home',]$team)
-        away_team <- unique(match[match$home_or_away=='Away',]$team)
-        home_goals <- nrow(match[match$to_team=='Home Goal',])
-        away_goals <- nrow(match[match$to_team=='Away Goal',])
-
-        home_score[home_team, away_team] <- home_goals
-        away_score[home_team, away_team] <- away_goals
-
-        if (home_goals > away_goals){
-            league_table[home_team,'W'] <- league_table[home_team,'W'] + 1
-            league_table[away_team,'L'] <- league_table[away_team,'L'] + 1
-        } else if (away_goals > home_goals){
-            league_table[home_team,'L'] <- league_table[home_team,'L'] + 1
-            league_table[away_team,'W'] <- league_table[away_team,'W'] + 1    
-        } else if (home_goals == away_goals) {
-            league_table[home_team,'D'] <- league_table[home_team,'D'] + 1
-            league_table[away_team,'D'] <- league_table[away_team,'D'] + 1
-        }
-
-        league_table[home_team,'GF'] <- league_table[home_team,'GF'] + home_goals
-        league_table[away_team,'GF'] <- league_table[away_team,'GF'] + away_goals
-        league_table[home_team,'GA'] <- league_table[home_team,'GA'] + away_goals
-        league_table[away_team,'GA'] <- league_table[away_team,'GA'] + home_goals
-
-        for (team in c(home_team, away_team)){
-            league_table[team,'GD'] <- league_table[team,'GF'] - league_table[team,'GA']
-            league_table[team,'PTS'] <- 3*league_table[team,'W'] + 1*league_table[team,'D']
-        }
-
-    }
-    return(list(league_table, home_score, away_score))
-}
-```
-
-```{r}
-results <- LeagueResults(events)
-```
-
-```{r}
-league_table <- data.frame(results[[1]])
-```
-
-```{r}
-league_table[order(-league_table$PTS),]
+nodes <- unique(c(as.character(data$source),as.character(data$target)))
+edges <- data.frame(data$source, data$target)
+g <- graph_from_data_frame(d=edges, vertices=nodes, directed=TRUE)
+ggraph(g, layout='linear') + 
+    geom_edge_arc(aes(start_cap = label_rect(node1.name), 
+                       end_cap = label_rect(node2.name)), 
+                  arrow=arrow(length=unit(4,'mm')), 
+                  color='grey', width=0.7, fold=F) +
+    geom_edge_loop(color='grey', linetype='dashed') +
+    geom_node_point(size=2, color='black', fill='white') +
+    geom_node_text(aes(label = name), repel=T, angle=90, hjust=2, ) +
+    theme_void()
 ```
@@
 ```
-
-```{r}
-
-```
-
-```{r}
-
-```
-
-```{r}
-ShowMatchEvents <- function(events, match_id, team_name, home_or_away='Home', flip=F) {
-    #' Visualize events on pitch for given match and team.
-
-    # select match
-    game <- events[events$matchId==match_id,]
-    goals <- game[game$subEventName=='Goal',]
-
-    # note teams
-    teams <- unique(game$team)
-
-    # limit events to team specified by name or home/away
-    if(!missing(team_name)){
-        game <- game[game$team==team_name,]
-        home_or_away <- unique(game$home_or_away)
-    } else {
-        game <- game[game$home_or_away==home_or_away,]
-        team_name <- unique(game$team)
-    }
-
-    # note opposition
-    opposition <- teams[teams!=team_name]
-
-    # get score
-    team_score <- table(goals$team)[team_name]
-    opposition_score <- table(goals$team)[opposition]
-
-    # flip coordinates if desired
-    if (flip==T) {
-        game$location_x <- 105 - game$location_x
-        game$location_y <- 68 - game$location_y
-        direction_of_play = 'l'
-    } else { 
-        direction_of_play = 'r'
-    }
-
-    # limit attention to main events
-    game_events <- game[game$eventName == 'Pass'| 
-                        game$eventName == 'Shot' | 
-                        game$eventName == 'On the Ball' | 
-                        game$eventName == 'Challenge' | 
-                        game$eventName == 'Free Kick',]
-
-    # draw pitch
-    p <- DrawPitch(theme='grey', arrow=direction_of_play) + 
-        geom_point(data = game_events, 
-               aes(location_x , location_y, fill=eventName, shape=eventName), pch=21, alpha=0.6, size=2 ) + 
-        geom_point(data=game_events[game_events$subEventName=='Goal',], 
-                   aes(location_x, location_y), shape=13, size=5) +
-        geom_label_repel(data=game_events[game_events$subEventName=='Goal',], 
-                         aes(location_x, location_y, label = paste0(source,'(',matchPeriod,' ',time,')')), label.padding=0.1, size=2.3, alpha=1) +
-        theme(legend.direction='horizontal', legend.position=c(0.5,0)) +
-        scale_fill_manual(values=c("red", "black", 'white', "yellow", 'blue'), name='Event Type') +
-        ggtitle(paste('Event Map:', team_name,'playing',home_or_away,'versus',opposition,'(',team_score,'-',opposition_score,')'))
-
-    return(p)
-
-}
-```
-
-```{r}
-match_id <- '2500032'
-```
-
-```{r}
-ShowMatchEvents(events, match_id, home_or_away='Home')
-ShowMatchEvents(events, match_id, home_or_away='Away', flip=T)
-```
-
-```{r}
-ShowPassesAndShots <- function(events, match_id, team_name, home_or_away='Home', flip=F) {
-    #' Visualize passes on pitch for given match and team.
-
-    # select match
-    game <- events[events$matchId==match_id,]
-    goals <- game[game$subEventName=='Goal',]
-
-    # note teams
-    teams <- unique(game$team)
-
-    # limit events to team specified by name or home/away
-    if(!missing(team_name)){
-        game <- game[game$team==team_name,]
-        home_or_away <- unique(game$home_or_away)
-    } else {
-        game <- game[game$home_or_away==home_or_away,]
-        team_name <- unique(game$team)
-    }
-
-    # note opposition
-    opposition <- teams[teams!=team_name]
-
-    # get score
-    team_score <- table(goals$team)[team_name]
-    opposition_score <- table(goals$team)[opposition]
-
-    # flip coordinates if desired
-    if (flip==T) {
-        game$location_x <- 105 - game$location_x
-        game$target_x <- 105 - game$target_x
-        game$location_y <- 68 - game$location_y
-        game$target_y <- 68 - game$target_y
-        direction_of_play = 'l'
-    } else { 
-        direction_of_play = 'r'
-    }
-
-    # limit attention to passes and shots
-    passes <- game[game$eventName == 'Pass' & game$team == team_name,]
-    shots <- game[(game$eventName=='Shot' | game$subEventName=='Goal') & game$team==team_name,]
-
-    # draw pitch
-    p <- DrawPitch(theme='grass', arrow=direction_of_play) + 
-        geom_segment(data=na.exclude(passes), 
-                     aes(x=location_x, y=location_y, xend=target_x, yend=target_y, color=subEventName), 
-                     alpha=0.6, arrow = arrow(length = unit(0.1,"cm"))) +
-        geom_label_repel(data=game[game$subEventName=='Goal',], 
-                     aes(location_x, location_y, label = paste0(source,'(',matchPeriod,' ',time,')')), label.padding=0.1, size=2.3, alpha=1) + 
-        geom_segment(data=na.exclude(shots), 
-                 aes(x=location_x, y=location_y, xend=target_x, yend=target_y, color=subEventName), 
-                 alpha=0.7, arrow = arrow(length = unit(0.1,"cm"))) +
-        theme(legend.position=c(0.5,-0.01), legend.direction='horizontal',
-              legend.background=element_rect(fill='#008000', linetype='solid')) +
-        scale_color_manual(values=c("red", "black", 'blue', 'orange', "yellow", 'grey', 'purple', 'white','brown'), name='Pass Type') +
-        ggtitle(paste('Pass Map:',team_name,'playing',home_or_away,'versus',opposition,'(',team_score,'-',opposition_score,')'))
-
-    return(p)
-} 
-```
-
-```{r}
-ShowPassesAndShots(events, match_id, home_or_away='Home')
-ShowPassesAndShots(events, match_id, home_or_away='Away', flip=T)
-```
-
-# Methodology
-
-```{r}
-PassNetwork <- function(events, match_id, team_name, team_colour='red', home_or_away='Home', flip=F, lower_threshold=1, high_threshold=10) {
-    #' Draw Pass Map of First XI with nodes placed on mean (x,y) pitch-coordinates.
-
-    game <- events[events$matchId == match_id,]
-    goals <- game[game$subEventName=='Goal',]
-
-    # note teams
-    teams <- unique(game$team)
-
-    # limit events to team specified by name or home/away
-    if(!missing(team_name)){
-        game <- game[game$team==team_name,]
-        home_or_away <- unique(game$home_or_away)
-    } else {
-        game <- game[game$home_or_away==home_or_away,]
-        team_name <- unique(game$team)
-    }
-
-    # note opposition
-    opposition <- teams[teams!=team_name]
-
-    # get score
-    team_score <- table(goals$team)[team_name]
-    opposition_score <- table(goals$team)[opposition]
-
-    firstXI <- game[game$FirstXI == 'True',]
-    mean_positions <- firstXI[firstXI$location_x>0 & firstXI$location_y>0 & firstXI$location_x<105 & firstXI$location_y<68,] %>% 
-        group_by(team, matchId, source) %>%
-          dplyr::summarise(x_mean = mean(location_x), y_mean = mean(location_y)) %>% 
-          ungroup() %>%
-          mutate(team = as.factor(team), id = as.factor(matchId)) %>%
-          as.data.frame()
-
-    pass_counts <- ddply(data.frame(game$source, game$target),.(game.source,game.target),nrow)
-    names(pass_counts) <- c('source','target','passcount')
-    step1 <- merge(mean_positions, pass_counts, by='source')
-    step2 <- step1[,c(1,4,5,7,8)]
-    names(step2)[2:3] <- c('source_x','source_y')
-    names(mean_positions)[3] <- 'target'
-    step3 <- merge(mean_positions, step2, by='target')
-    team <- step3[step3$team==team_name,]
-
-    if (flip==T) {
-        team$source_x <- 105 - team$source_x
-        team$x_mean <- 105 - team$x_mean
-        team$source_y <- 68 - team$source_y
-        team$y_mean <- 68 - team$y_mean
-        mean_positions$x_mean <- 105 - mean_positions$x_mean
-        mean_positions$y_mean <- 68 - mean_positions$y_mean
-    }
-
-    return (DrawPitch() +
-        geom_segment(data=team[team$passcount>=lower_threshold,], size=1, colour=team_colour,
-                     aes(x=source_x, y=source_y, xend=x_mean, yend=y_mean, alpha=passcount)) +
-        geom_segment(data=team[team$passcount>=high_threshold,], size=1.5, colour='black',
-                     aes(x=source_x, y=source_y, xend=x_mean, yend=y_mean, alpha=passcount)) +
-        geom_label_repel(data = mean_positions[mean_positions$team==team_name,], 
-                         aes(x_mean, y_mean, label = target), label.padding=0.1, size=2.3, alpha=1) +
-        geom_point(data=team, aes(x_mean, y_mean,), fill=team_colour, colour='black', pch=21, size=3) +
-        ggtitle(paste('Passing Network:',team_name,'playing',home_or_away,'versus',opposition,'(',team_score,'-',opposition_score,')')) +
-        theme(legend.position=c(0.5,0.07), legend.direction='horizontal'))
-
-}
-```
-
-```{r}
-PassNetwork(events, match_id, team_colour='blue')
-PassNetwork(events, match_id, home_or_away='Away', flip=T)
-```
-
-```{r}
-GameGraph <- function(events, match_id, team_name, home_or_away='Home') {
-    #' Return Possession Graph for given Match and Team.
-
-    game <- events[events$matchId == match_id,]
-    game <- game[game$source!='' & game$target!='nan',]
-
-     # note teams
-    teams <- unique(game$team)
-
-    # limit events to team specified by name or home/away
-    if(!missing(team_name)){
-        team_game <- game[game$team==team_name,]
-        home_or_away <- unique(team_game$home_or_away)
-    } else {
-        team_game <- game[game$home_or_away==home_or_away,]
-        team_name <- as.character(unique(team_game$team))
-    }
-
-    opposition_game <- game[game$team != team_name,]
-    opposition_team <- as.character(unique(opposition_game$team))
-
-    team_nodes <- unique(c(as.character(team_game$source),as.character(team_game$target)))
-    team_edges <- data.frame(team_game$source, team_game$target)
-    team_graph <- graph_from_data_frame(d=team_edges, vertices=team_nodes, directed=TRUE)
-    team_adj <- as.matrix(team_graph[])
-
-    possession_seqs <- unique(team_game$possession)
-
-    seq_start <- c()
-
-    for (pseq in possession_seqs){
-        if (sum(team_game$possession==pseq)>1){
-            player <- as.character(team_game[team_game$possession==pseq,][1,]$source)
-            if (player!='' & player!='nan'){
-                seq_start <- append(seq_start, player)
-            }
-        }
-    }
-
-    start_counts <- table(as.factor(seq_start))
-
-    for (i in (2:(length(start_counts)-1))){
-        team_adj[nrow(team_adj)-1,][names(start_counts[i])] <- as.numeric(start_counts[i])
-    }
-
-    opposition_status <- as.character(unique(opposition_game$home_or_away))
-    opp_to_opp <- as.numeric(summary(game[game$team != team_name,]$to_team)[opposition_status])
-
-    opp_nodes <- unique(c(as.character(opposition_game$source),as.character(opposition_game$target)))
-    opp_edges <- data.frame(opposition_game$source, opposition_game$target)
-    opp_graph <- graph_from_data_frame(d=opp_edges, vertices=opp_nodes, directed=TRUE)
-    opp_adj <- as.matrix(opp_graph[])
-
-    opp_to_opp <- sum(opp_adj[,ncol(opp_adj)-1])
-    team_adj[(nrow(team_adj)-1),(ncol(team_adj)-1)] <- opp_to_opp
-
-    team_status <- unique(team_game$home_or_away)
-    opposition_own_goals <- as.numeric(nrow(game[game$home_or_away==opposition_status & game$to_team==paste(team_status,'Goal'),]))
-    team_adj[(nrow(team_adj)-1),(ncol(team_adj))] <- opposition_own_goals
-
-    team_graph <- graph_from_adjacency_matrix(team_adj, mode='directed')
-
-    return(team_graph)
-}
-```
-
-```{r}
-AdjacencyMatrix <- function(graph){
-    #' Return Adjacency Matrix for Graph
-
-    return(as.matrix(graph[]))
-}
-```
-
-```{r}
-TransitionMatrix <- function(graph){
-    #' Return Transition Matrix for Graph considered as time-homogeneous Markov Process
-
-    matrix <- as.matrix(graph[])
-
-    for (i in 1:nrow(matrix)){
-        matrix[i,] <- matrix[i,]/sum(matrix[i,])
-    }
-
-    matrix[nrow(matrix),ncol(matrix)] <- 1
-    matrix[nrow(matrix),] <- rep(0, ncol(matrix))
-
-    # treat 'Goal' as absorbtion state
-    matrix[nrow(matrix),ncol(matrix)] <- 1
-
-    return(matrix)
-}
-```
-
-```{r}
-ExpectedScoringTime <- function(transition_matrix){
-    #' Calculate expected scoring time based on transition matrix.
-
-    max_steps <- 10000
-#     transition_matrix <- TransitionMatrix(GameGraph(events, match_id))
-    hitting_times <- rep(0,(ncol(transition_matrix)-2))
-
-    for (i in (1:(ncol(transition_matrix)-2))){
-
-        state_probabilities <- matrix(NA,nrow=max_steps+1,ncol=ncol(transition_matrix),dimnames=list(0:max_steps,(ncol(transition_matrix)-1):0))
-        vector <- rep(0,ncol(transition_matrix))
-
-        vector[i] <- 1
-        state_probabilities[1,] <- vector
-
-        for ( kk in 1:max_steps ) {
-            state_probabilities[kk+1,] <- t(transition_matrix)%*%state_probabilities[kk,]
-        }
-
-        probs <- diff(state_probabilities[,ncol(transition_matrix)])
-        hitting_time <- sum(probs*seq_along(probs))
-
-        hitting_times[i] <- hitting_time
-    }
-
-    names(hitting_times) <- names(transition_matrix[1,])[1:(nrow(transition_matrix)-2)]
-
-    return(hitting_times)
-
-}
-```
-
-```{r}
-x_score_time <- ExpectedScoringTime(
-    TransitionMatrix(
-        GameGraph(events, match_id, home_or_away='Home'))
-
-)
-```
-
-```{r}
-x_score_time/mean(x_score_time)
-```
-
-```{r}
-x_score_time/mean(x_score_time)
-```
-
-```{r}
-names(transition_matrix[1,])[1:(nrow(transition_matrix)-2)]
-```
-
-```{r}
-max_steps <- 10000
-transition_matrix <- TransitionMatrix(GameGraph(events, match_id))
-state_probabilities <- matrix(NA,nrow=max_steps+1,ncol=ncol(transition_matrix),dimnames=list(0:max_steps,(ncol(transition_matrix)-1):0))
-vector <- rep(0,ncol(transition_matrix))
-vector[6] <- 1
-state_probabilities[1,] <- vector
-
-for ( kk in 1:max_steps ) {
-    state_probabilities[kk+1,] <- t(transition_matrix)%*%state_probabilities[kk,]
-}
-
-probs <- diff(state_probabilities[,ncol(transition_matrix)])
-sum(probs*seq_along(probs))
-```
-
-```{r}
-transition_matrix
-```
-
-```{r}
-MatrixHeatMap <- function(matrix, number_format='%.d', title='Matrix Heatmap'){
-    #' Show Matrix as HeatMap
-
-    hm <- pheatmap(matrix,color=brewer.pal(15,'YlGnBu'),cluster_rows=F,cluster_cols=F,legend=F,
-                    display_numbers=T,number_format=number_format,fontsize_number=9,angle_col='315',
-                    main=title)
-
-    return(hm)
-}
-```
-
-```{r}
-MatrixHeatMap(
-    AdjacencyMatrix(g)[1:11,1:11]
-
-)
-```
-
-```{r}
-MatrixHeatMap(
-    AdjacencyMatrix(
-        GameGraph(events, match_id, home_or_away='Away'))[1:11,1:11]
-)
-```
-
-```{r}
-MatrixHeatMap(
-    TransitionMatrix(
-        GameGraph(events, match_id, home_or_away='Home')), 
-    number_format='%.2f',
-    title=''
-)
-```
-
-```{r}
-MatrixHeatMap(
-    TransitionMatrix(
-        GameGraph(events, match_id, home_or_away='Away')), 
-    number_format='%.2f',
-    title=''
-)
-```
-
-```{r}
-GoalSeqs <- function(events, match_id) {
-    match <- events[events$matchId==match_id,]
-    goals <- match[match$subEventName=='Goal',]
-    goal_seqs <- unique(goals$possession)   
-    return(goal_seqs)
-}
-```
-
-```{r}
-TabulateSequence <- function(events, possession_sequence) {
-    # Return a markdown table starting from the row before a possession sequence begins, and continuing until it ends.
-
-    start <- min(events[events$possession == possession_sequence,]$X)
-    stop <- max(events[events$possession == possession_sequence,]$X) + 1
-    table <- events[start:stop,]
-    df <- data.frame(table$matchPeriod, table$time, table$team, table$source, table$subEventName)
-    names(df) <- c('Half', 'Time', 'Team', 'Event Description', 'Player')
-
-    return(kable(df))
-}
-```
-
-```{r}
-sequence <- '2500032-91-Liv-0'
-```
-
-```{r}
-TabulateSequence(events, sequence)
-```
-
-```{r}
-SequenceOnPitch <- function(events, possession_sequence){
-    #' Draw Possession Sequence on Pitch
-
-    data <- events[events$possession == possession_sequence,]
-
-    p <- DrawPitch(theme='grass') + 
-        geom_label_repel(data = data, aes(location_x, location_y, label = paste(source,time)), label.padding=0.1, size=2.3, alpha=1) +
-        geom_curve(data = data, aes(x = location_x, xend = target_x, y = location_y, yend = target_y*.99, col = X), 
-                   show.legend=FALSE, size=1, alpha = 1) +
-        geom_point(data = data, aes(location_x , location_y, shape=eventName), size=2) + 
-        ggtitle(paste0(nrow(data),'-part ', unique(data$team)[1], ' Possession Sequence ending with ', 
-                       data[nrow(data),]$subEventName, ' by ', data[nrow(data),]$source))  +
-        theme(legend.position=c(0.5,0.08), legend.direction='horizontal', legend.title=element_blank())
-
-    return(p)
-}
-```
-
-```{r}
-SequenceOnPitch(events, sequence)
-```
-
-youtube clip
-<img src="../figures/youtube_clip.gif" width='700'>
-
-```{r}
-SequenceGraph <- function(events, possession_sequence){
-    #' Return Graph of Possession Sequence
-
-    data <- events[events$possession == possession_sequence,]
-    data <- data[data$source!='' & data$target!='nan',]
-    nodes <- unique(c(as.character(data$source),as.character(data$target)))
-    edges <- data.frame(data$source, data$target)
-    g <- graph_from_data_frame(d=edges, vertices=nodes, directed=TRUE)
-    g$id <- possession_sequence
-    g$data <- data
-    g$team <- as.character(unique(data$team))
-
-    return(g)
-}
-```
-
-```{r}
-VisualizeGraph <- function(possession_sequence_graph) {
-    #' Visualize Possession Sequence as Linear Graph with Looping Edges
-
-    g <- possession_sequence_graph
-    data <- possession_sequence_graph$data
-
-    visualization <- ggraph(g, 'linear') + 
-        geom_edge_arc(aes(color=data$eventName), 
-                      arrow=arrow(length=unit(4,'mm')), 
-                      fold=F,
-                     width=1) +
-        geom_edge_loop(aes(color=data$eventName),
-                      width=1) +
-        geom_node_point(color='black', 
-                        size=2,
-                        alpha=0.5) +
-        geom_node_text(aes(label = name), 
-                        repel=T, 
-                        angle=90, hjust=2, ) +
-        scale_edge_colour_manual(
-            values=c('indianred3', 'wheat4', 'grey', 'grey30',
-                    'red','blue','green','orange','purple','brown','pink'
-                    ),
-            name='Event Type') +
-        theme_void()
-
-    return(visualization)
-
-}
-```
-
-```{r}
-g <- SequenceGraph(events, sequence)
-```
-
-```{r}
-VisualizeGraph(g)
-```
-
-```{r}
-GetGraphMetrics <- function(possession_graph) {
-    #' Return basic metrics for given Graph
-
-    g <- possession_graph
-    path_metrics <- c(g$id, g$team, length(V(g)), length(E(g)), 
-                      diameter(g), mean_distance(g), edge_density(g, loops=T), 
-                      girth(g)$girth, transitivity(g), authority_score(g)$value, hub_score(g)$value)
-    names(path_metrics) <- c('Label', 'Team', 'Nodes', 'Edges', 'Diameter', 'Mean_Distance', 
-                             'Edge_Density', 'Girth', 'Transitivity', 'Authority_Score', 'Hub_Score')
-
-    return(path_metrics)
-}
-```
-
-```{r}
-GetNodeMetrics <- function(possession_graph) {
-    #' Return metrics for each Node of given Graph
-
-    g <- possession_graph
-    df <- data.frame(page_rank(g)$vector)
-    names(df) <- 'pagerank'
-    df$in_degree <- degree(g, mode='in')
-    df$out_degree <- degree(g, mode='out')
-    df$eccentricity <- eccentricity(g)
-    df$closeness <- closeness(g)
-    df$betweenness <- betweenness(g)
-    df$norm_betweenness <- betweenness(g, normalized=T)
-    df$alpha_centrality <- alpha_centrality(g)
-    df$power_centrality <- power_centrality(g)
-    df$eigen_centrality <- eigen_centrality(g)$vector
-    df$authority <- authority_score(g)$vector
-    df$hub <- hub_score(g)$vector
-
-    return(df)
-}
-```
-
-```{r}
-GetEdgeMetrics <- function(possession_graph) {
-    #' Return metrics for each Edge of given Graph
-
-    g <- possession_graph
-    df <- data.frame(as_edgelist(g))
-    names(df) <- c('Source','Target')
-    df$sequence <- 1:length(E(g))
-    df$edge_betweenness <- edge_betweenness(g)
-
-    return(df)
-}
-```
-
-```{r}
-GetGraphMetrics(g)
-```
-
-```{r}
-GetNodeMetrics(match_graph)
-```
-
-```{r}
-GetEdgeMetrics(g)
-```
-
-```{r}
-sequence2 <- '2499858-18-Liv-5'
-```
-
-```{r}
-TabulateSequence(sequence2)
-```
-
-```{r}
-SequenceOnPitch(sequence2)
-```
-
-<img src='../figures/mane_vs_stoke.gif'>
-
-```{r}
-g2 <- SequenceGraph(sequence2)
-VisualizeGraph(g2)
-```
-
-```{r}
-GetGraphMetrics(g2)
-```
-
-```{r}
-sequence3 <-  '2499889-18-Liv-0'
-```
-
-```{r}
-SequenceOnPitch(sequence3)
-```
-
-```{r}
-g3 <- SequenceGraph(sequence3)
-VisualizeGraph(g3)
-```
-
-```{r}
-GetGraphMetrics(g3)
-```
-
-<img src="../figures/youtube_clip2.gif">
-
-```{r}
-
-```

### 2020-07-16 rmd/draft.Rmd
writing draft

---
title: "Random Walks across Social Networks: The Markovian Movement of a Stochastic Soccerball"

jupyter:
  jupytext:
    formats: nb//ipynb,rmd//Rmd
    text_representation:
      extension: .Rmd
      format_name: rmarkdown
      format_version: '1.2'
      jupytext_version: 1.4.2
  kernelspec:
    display_name: R
    language: R
    name: ir
---

## Executive Summary

## Introduction

@RPollardReep1997
@RPollard2002
@RPollardEtAl1988
@CReepBenjamin1968
@CReepEtAl1971

even earlier @JEaves2015

David Barrett [@ECochraneEtAl1910]

## Data

dataset
@LPappalardoEtAl2019a

nested JSON

- load JSON
- decode the 'tags', player IDs and team IDs.
- unpack the location and target positions.
- distinguish play sequences, as indicated by a 'Free Kick' (includes Throw-in) or a 'Goal'.
- distinguish possession sequences
- include starting XI and Home/Away status info
- convert time-stamp in seconds to classic minutes

load into R, test that league table given by preprocessed data matches actual results. Table 1
- found that De Bruyne's goal was misclassified as an inaccurate pass, rather than a goal-scoring shot.

Table 1
Premiership Top 5

Initial exploratory visualizations:

- points showing events in a match for home and away team

Figure 1a,b
<img src="../figures/EventMapHome.png">

Figure 2a,b
- arrows showing passes and shots
<img src="../figures/PassMapHome.png">

More specifically: we consider a specific possession sequence to check that our preprocessing strategy has been successfuly.
Table 2: Events of Possession Sequence
Figure 3: Pitchmap showing Possession Sequence
<img src="../figures/PossessionSequence.png">

Figure 4: YouTube Clip showing successful conclusion of possession sequence
<img src="../figures/youtube_clip.png">

## Methodology

### Graph Theory Definitions

*Social network analysis* uses the tools of mathematical graph theory to analyse social structures. We therefore begin by defining several necessary theoretical terms.

A *graph* is an ordered tuple $G = (V,E)$, consisting of a set of *nodes* (or *vertices*) $V = \{v_{i}\}$, and a set of *edges* $E = \{e_{ij}\}$, where the edge $e_{i,j}$ is the ordered pair $(i,j)$ representing some connection from the *source* $v_{i}$ to the *target* $v_{j}$.

Two nodes connected by an edge are said to be *adjacent* to each other. An edge $e_{ii}$ connecting a node $v_{i}$ to itself is called a *loop*; the node is then *self-adjacent*.

A graph is *undirected* if $e_{ij} \iff e_{ji}$ -- otherwise it is *directed*. It is a *simple* graph if the edges are distinct and unrepeated -- otherwise it is a *multigraph*.

Given a graph G, we can describe a *walk* of length L as a sequence of adjacent (but not necessarily distinct) nodes $(v_{0},...,v_{L})$; or, equivalently, as a sequence of edges $e_{0,1},...,e_{L-1,L}$. Conversely, given a set of walks, we can construct the (minimal) underlying graph containing all the nodes and edges involved in the walks.

A graph is *weighted* if there exists some function $w: E \mapsto \mathbb{R}$ assigning a weight value to each edge.

An unweighted multigraph can thus be viewed as a weighted simple graph by defining $w(e_{i,j}) = |\{e \in E: e = e_{i,j}\}| \in \mathbb{N}_{0}$, so that the weight of an edge is the number of times it is repeated.

A weighted simple graph can be uniquely described by an *adjacency matrix*:
$$
M_{ij} = \begin{cases}
    \begin{array}{ll}
         w(e_{i,j}), & e_{i,j} \in E \\
       0, & \mathrm{otherwise}
    \end{array}
\end{cases}
$$

There are then a large variety of different graph metrics which can be calculated, including in particular an ever-increasing number of measures of *centrality*.

### Graph Theory Application

We consider the possession sequences of a football match as *walks* of the football across a social graph in which the nodes are players (considered either individually or collectively), and the connecting edges are defined by the movement of the ball, as it travels from player to player. We can then construct the underlying multigraph for any particular possession sequence (as in Figure i), or indeed for the whole game.

Figure 5
<img src="../figures/PossessionGraph.png">

There are several different 'weights' we could assign an edge: the Euclidean distance moved by the ball; the polar angle; the time interval. However, for this particular analysis we ignore the spatial dynamics and the subtler aspects of continuous time, and focus on the network connections between players.

We do not restrict our analysis to actual passes but include touches on the ball, challenges, and shots. Regardless of the type of event, we consider any movement of the ball from player $U$ to player $V$ as an edge $e_{U,V}$, including the loop $e_{U,U}$ where the player dribbles with the ball. As suggested above, we let the weight of the edge $e_{U,V}$ be the number of times the ball moves from player U to player V. We can then generate the adjacency matrix for a team in a particular match.

Figure 6
<img src="../figures/AdjacencyMatrixHome.png">

It is increasingly common [@TKnutson2018] to visualize a possession network by positioning each node according to the mean location of the player's touches, and varying the appearance (width, colour or transparency) of the edges to show the frequency of that pass. This swiftly communicates a team's formation, and gives a qualitative sense of a team's strategy. However, visualizing the mean position of a player could be misleading: if a player always plays on the wing, but is switched half-way through the game from one side of the pitch to the other, his 'average position' will be central, although he may never have touched the ball centrally. It is also not possible in a static image to distinguish a pass $e_{U,V}$ from its reverse $e_{V,U}$. @KSingh2018 therefore suggests creating interactive visualizations in which the user can adjust the weighting to highlight forward or backwards passes as desired.

Figure 7
<img src="../figures/PassingNetworkHome.png">

### Probability and Statistics

We then also need some definitions from *probability theory* and *statistics*. [@RSerfozo2009a]

A *sample space* $\Omega = \{\omega_{i}\}$ is the set of possible *outcomes* of an *observation*. If we are observing who has the ball, then the sample space is the set of players on the pitch. An *event* $F$ is some subset of $\Omega$, and the *event space* $\mathcal{F}$ is the set of subsets of $\Omega$. In particular, $\Omega \subset \mathcal{F}$.

To illustrate the difference between the sample space and the event space, we could observe whether the *home team* is in possession of the ball. The outcome of observing who has the ball will not be a whole team, but merely some particular player $\omega_{x}$, whom we could call $h_{x}$ if he is on the home team, or $a_{x}$ if he is on the away team. So our sample space $\Omega = \{\omega_{i}\} = \{h_{i}\} + \{a_{i}\}$, and the *event* that the home team has possession $H = \{h_{i}\} \in \mathcal{F}$.

A *probability* $\mathbb{P}$ is then a function $\mathbb{P}: \mathcal{F} \mapsto [0,1] \in \mathbb{R}$, such that $\sum\limits_{\omega_{i} \in \Omega}\mathbb{P}(\omega_{i}) = 1$ and $\mathbb{P}(\O) = 0$. We can also talk about the *conditional probability* of an event $X$ given another event $Y$, which we write $\mathbb{P}(X | Y)$.

A *random variable* is a function $X: \Omega \mapsto S$. If $S \subseteq \mathbb{R}$ then we also have the *expectation* $\mathbb{E}[X]$ and *variance* $\mathrm{Var}[X]$, but it is acceptable to have an $S \not\subset \mathbb{R}$. Again, to illustrate what we mean -- the number of goals in the match could be considered as a random variable, but so could our previous question of which player has the ball at a particular time.

A *stochastic process* is an indexed set (that is, a *sequence*) of random variables, where the index commonly refers to points in time $t \in T$, which may be considered discretely (so $t \in \mathbb{N}_{0}$ ) or continuously (so $t \in \mathbb{R}_{ \ge 0}$ ). The range $S$ of a stochastic process $\{X(t,\omega):t\in T,\omega \in \Omega\}$ is called its *state space*, and the value $X_{t}$ is its *state* at time $t$.

A stochastic process has the *Markov property* if $\forall t \ge 0, \mathbb{P}(X_{t+1}=s| X_{0},...,X_{t}) = \mathbb{P}(X_{t+1}=s| X_{t})$; this means that future states are dependent only on the present state, regardless of the previous history of past states. Such a process can be called a Markov process, and we call the probability $\mathbb{P}(X_{t+1}=s_{j}|X_{t}=s_{i})$ the *transition probability* $p_{ij}$ .

A Markov process is *time-homogenous* if the transition probabilities stay constant through time, ie. $\mathbb{P}(X_{n+1}=s|X_{n}) = \mathbb{P}(X_{m+1}=x|X_{m}) \forall m,n \in T$. If $S$ is countable, then the probability distribution of such a process can be described by its initial *distribution vector* $\pi_{0}$ and its *transition matrix* $P$, where each $i$th row is made up of the transition probabilities $p_{ij}$ and $\sum\limits_{j} p_{ij} = 1 \forall i$. Specifically, after $k$ transitions we have the distribution $\pi_{k} = \pi_{0}P^{k}$.

If a Markov process develops in discrete time and takes a values from a discrete state space, then we call it a *Markov chain* and can straightforwardly draw it as a weighted simple graph by treating its transition matrix as an adjacency matrix. Conversely, we can transform an adjacency matrix $M$ into a transition matrix $P$ by defining the elements $p_{ij} := \frac{m_{ij} }{\sum\limits_{j}m_{ij}}$, so that each row sums to 1 as required. We can then picture the development of a Markov process as a *random walk* on the graph defined by its transition probabilities, noting that since the transition probabilities are not uniform this walk is *biased*.

Figure 8
<img src="../figures/TransitionMatrixHome.png">

In Figure 8 we show the heatmap of the Transition Matrix derived from the Adjacency Matrix shown in Figure 6. There we showed only the adjacency counts for the starting eleven, but now we include the subsitutes, as well as treating the 'Opposition' collectively as a single node. Most significantly, we now include a 'Goal' as a node. In every other case where the ball goes out of play, we ignore the details of whether it was for a corner, throw-in, offside, or foul, and simply consider the ball to have moved from whichever player had the ball to the opposition. (Consequently, the players which the Adjacency Matrix indicates get the ball most often from the opposition are generally full-backs who take throw-ins, rather than players whose challenges are actually winning the ball, as one might initially suppose). But a goal is of unique significance, and for a social network analysis of football to be anything more than a diverting curiosity, we need to be able to say something meaningful about what a team's possession graph tells us about the likelihood of goals.

To do this, as we have said, we treat a goal as a node in its own right; in particular, in the terms of Markov theory, we treat it as an *absorbing state*. Clearly a football match continues after a goal is scored, but from the perspective of our model, the match is considered as a series of stochastic processes which end and immediately start anew if and when a goal is scored.

This allows us to calculate the expected *hitting time* of a goal, given that any particular player has the ball.

The hitting time $\tau_{R}$ of some state(s) $R \subset S$ is given by $\tau_{R} = min\{ t \ge 0 \colon X_{t} \in R \}$.

### Statistical Application
The extent of a sample space depends very much on the nature of its corresponding observation: if we are observing who has the ball at a particular moment, then the sample space will consist of the players; if we are considering

## Results

## Conclusion

## Bibliography

### 2020-07-17 rmd/draft.Rmd
explain MXG

@@
 
-### Graph Theory Definitions
+### Graph Theoretic Foundations
 
@@
 
-### Graph Theory Application
+### Graphing a Team's Possession Network
 
 We consider the possession sequences of a football match as *walks* of the football across a social graph in which the nodes are players (considered either individually or collectively), and the connecting edges are defined by the movement of the ball, as it travels from player to player.
-We can then construct the underlying multigraph for any particular possession sequence (as in Figure i), or indeed for the whole game.
+We can then construct the underlying graph for any particular possession sequence (as in Figure i), or indeed for the whole game.
 
@@
 
-### Probability and Statistics
+### Probabilistic Walks on a Graph
 
@@
 The outcome of observing who has the ball will not be a whole team, but merely some particular player $\omega_{x}$, whom we could call $h_{x}$ if he is on the home team, or $a_{x}$ if he is on the away team.
-So our sample space $\Omega = \{\omega_{i}\} = \{h_{i}\} + \{a_{i}\}$, and the *event* that the home team has possession $H = \{h_{i}\} \in \mathcal{F}$.
+So our sample space $\Omega = \{\omega_{i}\} = \{h_{i}\} \cup \{a_{i}\}$, and the *event* that the home team has possession $H = \{h_{i}\} \in \mathcal{F}$.
 
@@
 If $S \subseteq \mathbb{R}$ then we also have the *expectation* $\mathbb{E}[X]$ and *variance* $\mathrm{Var}[X]$, but it is acceptable to have an $S \not\subset \mathbb{R}$.
-Again, to illustrate what we mean -- the number of goals in the match could be considered as a random variable, but so could our previous question of which player has the ball at a particular time.
+Again, to illustrate what we mean -- the number of goals in the match could be considered as a real-valued random variable, and we could talk about the expectation and variation of the number of goals; but this is not the case for the question of which player has the ball at a particular time, though this also is a random variable.
 
@@
 
+### Calculating Markovian Expected Goals
+
 In Figure 8 we show the heatmap of the Transition Matrix derived from the Adjacency Matrix shown in Figure 6.
@@
 
-To do this, as we have said, we treat a goal as a node in its own right; in particular, in the terms of Markov theory, we treat it as an *absorbing state*.
+To do this, as we have said, we treat a goal as a node in its own right; in particular, in the terms of Markov theory, we treat it as an *absorbing state* which loops back on itself with certain probability.
 Clearly a football match continues after a goal is scored, but from the perspective of our model, the match is considered as a series of stochastic processes which end and immediately start anew if and when a goal is scored.
-
 This allows us to calculate the expected *hitting time* of a goal, given that any particular player has the ball.
 
-The hitting time $\tau_{R}$ of some state(s) $R \subset S$ is given by $\tau_{R} = min\{ t \ge 0 \colon X_{t} \in R \}$.
+The hitting time $\tau_{s}$ of some state $s \in S$ is given by $\tau_{s} := min\{ t \ge 0 \colon X_{t} = s \}$, the earliest time at which the process reaches that state.
 
-### Statistical Application
-The extent of a sample space depends very much on the nature of its corresponding observation: if we are observing who has the ball at a particular moment, then the sample space will consist of the players; if we are considering
+Given an initial distribution $\pi_{0}$, we know that after $k$ transitions, the state $\pi_{k} = \pi_{0}P^{k}$.
+
+We can define the *state vector* $\delta$ of any particular state $s{i}$ as
+$$
+\delta_{j} := \begin{cases}
+    \begin{array}{ll}
+    	1, & i=j \\
+	0, & \mathrm{otherwise}
+    \end{array}
+\end{cases}
+$$
+
+The probability $\mathbb{P}(X_{k}=s)$ is then given by the dot product $\langle \pi_{k}, \delta \rangle$.
+
+If $s_{i}$ is an absorbing state, then the probability that it is reached *for the first time* is given by  $p_{k}:= \langle \pi_{k}, \delta \rangle - \langle \pi_{k-1}, \delta \rangle$.
+
+If the expectation of the hitting time is finite it can then be approximated directly by iterating up to some large $K$:
+$$\mathbb{E}[\tau_{s}] = \sum_{k=0}^\infty kp_k \approx \sum_{k=0}^K kp_k$$
+
+Thus we can quantify, for each player on the team, the expected number of events there will be before the team scores a goal.
+We invert this to give a measure of the expected number of goals; and scale it by the team's average number of events per game, and by the average player's expected goals values, and call this measure *Markov Expected Goals* (MXG).
 

### 2020-07-20 rmd/draft.Rmd
just need to write results and conclusions, but i've used up wordcount

@@
 
+Association football is the world's most popular game and a multibillion dollar industry.
+Statistical analysis of the sport was pioneered more than fifty years ago, but has only recently begun to make a more widespread impact.
+In this paper I explain how Social Network Analysis can provide the tools to model a football team's ball-possession as a dynamic probabilistic system, and hence how we can quantify the effect of each player on the team's goal-scoring capability.
+
 ## Introduction
 
-@RPollardReep1997
-@RPollard2002
-@RPollardEtAl1988
-@CReepBenjamin1968
-@CReepEtAl1971
+After thirty years of waiting, Liverpool F.C.
+are once again English champions [@BBC2020].
+But while manager Jurgen Klopp is now guaranteed his place in footballing folklore, less well known is the name of data scientist Ian Graham, whose analysis led to Klopp being hired in the first place [@BSchoenfeld2019], and whose research team is regarded as "the strongest of its kind in English football" [@SAustin2020].
 
-even earlier @JEaves2015
-
-David Barrett [@ECochraneEtAl1910]
+In this paper, I develop a rigorous account of a football match as a stochastic social network, the graph of which is generated by the movement of the ball from player to player.
+I then show how the adjacency matrix of a player, considered as a node in a football match's possession sequence graph, can be used to quantify his contribution to the likelihood of the team scoring.
+Finally, I discuss the potential and pitfalls of this new metric, which I call *Relative Markovian Expected Goals*.
 
@@
 
-dataset
-@LPappalardoEtAl2019a
+### JSON Event-logs
 
-nested JSON
+The dataset is provided by WyScout, one of a growing number of companies in the football analytics industry, and is available from FigShare.com [@LPappalardoMassucco2020].
+It includes more than three million *event-logs* for every match from the five major European leagues, in JSON format (see Figure 1), with  the details of players, teams, and informative tags (such as the crucial 'Goal') numerically encoded.
+Each event is associated with an event type (such as 'Shot', 'Pass' or 'Free Kick'), identified with a player, and includes a nested array of the locational *positions* of the event itself and its target.
 
-- load JSON
-- decode the 'tags', player IDs and team IDs.
-- unpack the location and target positions.
-- distinguish play sequences, as indicated by a 'Free Kick' (includes Throw-in) or a 'Goal'.
-- distinguish possession sequences
-- include starting XI and Home/Away status info
-- convert time-stamp in seconds to classic minutes
+### Preprocessing
 
-load into R, test that league table given by preprocessed data matches actual results. Table 1
-- found that De Bruyne's goal was misclassified as an inaccurate pass, rather than a goal-scoring shot.
+I used Python to convert this into a *tidy* dataframe [@HWickham2014], that could then be analyzed and visualized using R.
+For the sake of reproducible research, all code is attached as appendices, as well as available on Github.
 
-Table 1
-Premiership Top 5
+As part of this preprocessing stage I also labelled each event according first to its *in-play sequence*, and then its *possession sequence*.
+An in-play sequence ends when the ball goes out of play, as indicated by the subsequent event having the *eventName* 'Free Kick' (a category which includes throw-ins and corners, as well as actual free-kicks given for fouls or offsides); or by the final two events (the shot and the attempted save) being labelled with the *tag* 'Goal'.
+(Although every goal is followed by a 'free kick' from the centre of the pitch, this never seems to be included in the event logs, presumably because they are created from camera footage which focusses on the goal celebrations and tends to miss the immediate restart of play after a goal).
 
-Initial exploratory visualizations:
+Intuitively, to identify a possession sequence we should just be able to iterate through the event-logs of a given in-play sequence, and note when the team of the player changes.
+However, since the event-logs include challenges for the ball even when they are unsuccessful, this would lead to systematically undercounting the length of possession sequences.
+The solution I adopt is to iterate through the event-logs for a single team within an in-play sequence, paying close attention to the location data: if the location of an event is identical to the target-location of the previous event associated with that same team, then I treat is as part of the same possession sequence, otherwise I treat it as a separate sequence.
 
-- points showing events in a match for home and away team
+### Exploratory Visualization
 
-Figure 1a,b
-<img src="../figures/EventMapHome.png">
+We can then visualize the location of the event-log data points in the context of a diagrammatic football pitch, to allow for an immediate and accessible view of any particular possession sequence (eg.
+Table 2, Figure 3, Figure 4).
 
-Figure 2a,b
-- arrows showing passes and shots
-<img src="../figures/PassMapHome.png">
-
-More specifically: we consider a specific possession sequence to check that our preprocessing strategy has been successfuly.
-Table 2: Events of Possession Sequence
 Figure 3: Pitchmap showing Possession Sequence
@@
 
+We can compare this to a tabulated account of all events from just before the beginning of the possession sequence until its end, we see that our algorithm has successfully handled the case where an opposition player's unsuccessful challenge interrupts the sequence of events associated with the team in possession.
+
+Table 2: Events of Possession Sequence
+
+And with possession sequences that end in a goal, it is generally possible to find video highlights of the events under consideration, and confirm that the data we have does correspond to what actually happened.
+
 Figure 4: YouTube Clip showing successful conclusion of possession sequence
-<img src="../figures/youtube_clip.png">
+<img src="../figures/youtube_clip.gif">
 
@@
 
-*Social network analysis* uses the tools of mathematical graph theory to analyse social structures. We therefore begin by defining several necessary theoretical terms.
+Social network analysis applies the concepts of mathematical *graph theory*.
 
@@
 A graph is *undirected* if $e_{ij} \iff e_{ji}$ -- otherwise it is *directed*. It is a *simple* graph if the edges are distinct and unrepeated -- otherwise it is a *multigraph*.
+
+On an undirected graph, we call the number of edges connecting a node its *degree*; in a directed graph we distinguish between the *indegree* and *outdegree*.
 
@@
 $$
-
-There are then a large variety of different graph metrics which can be calculated, including in particular an ever-increasing number of measures of *centrality*.
 
@@
 This swiftly communicates a team's formation, and gives a qualitative sense of a team's strategy.
-However, visualizing the mean position of a player could be misleading: if a player always plays on the wing, but is switched half-way through the game from one side of the pitch to the other, his 'average position' will be central, although he may never have touched the ball centrally.
-It is also not possible in a static image to distinguish a pass $e_{U,V}$ from its reverse $e_{V,U}$.
-@KSingh2018 therefore suggests creating interactive visualizations in which the user can adjust the weighting to highlight forward or backwards passes as desired.
-
 Figure 7
@@
 
-We then also need some definitions from *probability theory* and *statistics*. [@RSerfozo2009a]
+To consider the team's ball-possession in its dynamic development, rather than merely as a static structure, we also need some definitions from *probability theory*.
 
@@
 $\mathbb{P}(X_{n+1}=s|X_{n}) = \mathbb{P}(X_{m+1}=x|X_{m}) \forall m,n \in T$.
-If $S$ is countable, then the probability distribution of such a process can be described by its initial *distribution vector* $\pi_{0}$ and its *transition matrix* $P$, where each $i$th row is made up of the transition probabilities $p_{ij}$ and $\sum\limits_{j} p_{ij} = 1 \forall i$.
-Specifically, after $k$ transitions we have the distribution $\pi_{k} = \pi_{0}P^{k}$.
+If $S$ is countable, then the probability distribution of such a process can be described by its initial *distribution vector* $\boldsymbol{\alpha_{0}}$ and its *transition matrix* $\mathbf{P}$, where each $i$th row is made up of the transition probabilities $p_{ij}$ and $\sum\limits_{j} p_{ij} = 1 \forall i$.
+Specifically, after $k$ transitions we have the distribution $\boldsymbol{\alpha_{k}} = \boldsymbol{\alpha_{0}}\mathbf{P}^{k}$.
 
 If a Markov process develops in discrete time and takes a values from a discrete state space, then we call it a *Markov chain* and can straightforwardly draw it as a weighted simple graph by treating its transition matrix as an adjacency matrix.
-Conversely, we can transform an adjacency matrix $M$ into a transition matrix $P$ by defining the elements $p_{ij} := \frac{m_{ij} }{\sum\limits_{j}m_{ij}}$, so that each row sums to 1 as required.
-We can then picture the development of a Markov process as a *random walk* on the graph defined by its transition probabilities, noting that since the transition probabilities are not uniform this walk is *biased*.
+Conversely, we can transform an adjacency matrix $\mathbf{M}$ into a transition matrix $\mathbf{P}$ by defining the elements so that each row sums to 1 as required:
+
+$p_{ij} := \frac{m_{ij} }{\sum\limits_{j}m_{ij}}$
+
+We can then picture the development of a Markov process as a *random walk* on the graph in which the nodes ${V}$ of the graph are the state space $S$ of the process, and the weights of the edges are the transition probabilities.
 
@@
 
-Given an initial distribution $\pi_{0}$, we know that after $k$ transitions, the state $\pi_{k} = \pi_{0}P^{k}$.
+Given an initial distribution $\boldsymbol{\alpha_{0}}$, we know that after $k$ transitions, the state $\boldsymbol{\alpha_{k}} = \boldsymbol{\alpha_{0}}\mathbf{P}^{k}$.
 
-We can define the *state vector* $\delta$ of any particular state $s{i}$ as
+We can define the *state vector* $\boldsymbol{\delta_{i}}$ of any particular state $s_{i}$ as
 $$
-\delta_{j} := \begin{cases}
+\delta_{ij} := \begin{cases}
     \begin{array}{ll}
     	1, & i=j \\
-	0, & \mathrm{otherwise}
+	0, & \mathrm{otherwise.}
     \end{array}
@@
 
-The probability $\mathbb{P}(X_{k}=s)$ is then given by the dot product $\langle \pi_{k}, \delta \rangle$.
+The probability $\mathbb{P}(X_{k}=s_{i})$ is then given by the dot product $\langle \boldsymbol{\alpha_{k}}, \boldsymbol{\delta_{i}} \rangle$.
 
-If $s_{i}$ is an absorbing state, then the probability that it is reached *for the first time* is given by  $p_{k}:= \langle \pi_{k}, \delta \rangle - \langle \pi_{k-1}, \delta \rangle$.
+If $s_{i}$ is an absorbing state, then the probability that it is reached *for the first time* is given by $\pi_{k}:= \langle \boldsymbol{\alpha_{k}}, \boldsymbol{\delta_{i}} \rangle - \langle \boldsymbol{\alpha_{k-1}}, \boldsymbol{\delta_{i}} \rangle$.
 
 If the expectation of the hitting time is finite it can then be approximated directly by iterating up to some large $K$:
-$$\mathbb{E}[\tau_{s}] = \sum_{k=0}^\infty kp_k \approx \sum_{k=0}^K kp_k$$
+$$\mathbb{E}[\tau_{s}] = \sum_{k=0}^\infty k\pi_k \approx \sum_{k=0}^K k\pi_k$$
 
 Thus we can quantify, for each player on the team, the expected number of events there will be before the team scores a goal.
-We invert this to give a measure of the expected number of goals; and scale it by the team's average number of events per game, and by the average player's expected goals values, and call this measure *Markov Expected Goals* (MXG).
+We invert this to give a measure of the expected number of goals; and scale it by the team's average number of events per game, we call this measure *Markovian Expected Goals* (MXG).
+We then scale this by the team's mean expected goals values to give a *Relative MXG* value, so that values can be compared fairly across teams to see which individual players contribute the most to their teams' goalscoring chances.
 
 ## Results
+
+T
 

### 2020-07-20 rmd/draft.Rmd
almost done writing

@@
 
-I used Python to convert this into a *tidy* dataframe [@HWickham2014], that could then be analyzed and visualized using R.
-For the sake of reproducible research, all code is attached as appendices, as well as available on Github.
+I used Python to convert this into a *tidy* dataframe [@HWickham2014], which I then analyzed and visualized using R.
+For the sake of reproducible research [@RPeng2011], all code is attached as appendices, as well as available on Github.
 
@@
 Most significantly, we now include a 'Goal' as a node.
-In every other case where the ball goes out of play, we ignore the details of whether it was for a corner, throw-in, offside, or foul, and simply consider the ball to have moved from whichever player had the ball to the opposition.
-(Consequently, the players which the Adjacency Matrix indicates get the ball most often from the opposition are generally full-backs who take throw-ins, rather than players whose challenges are actually winning the ball, as one might initially suppose).
-But a goal is of unique significance, and for a social network analysis of football to be anything more than a diverting curiosity, we need to be able to say something meaningful about what a team's possession graph tells us about the likelihood of goals.
-
-To do this, as we have said, we treat a goal as a node in its own right; in particular, in the terms of Markov theory, we treat it as an *absorbing state* which loops back on itself with certain probability.
-Clearly a football match continues after a goal is scored, but from the perspective of our model, the match is considered as a series of stochastic processes which end and immediately start anew if and when a goal is scored.
+In particular, in the terms of Markov theory, we treat it as an *absorbing state* which loops back on itself with certain probability.
 This allows us to calculate the expected *hitting time* of a goal, given that any particular player has the ball.
@@
 
-## Results
+### Results
 
-T
+The
 
@@
 
+We have here focussed on the attacking contribution a player makes towards his team's goal-scoring capability, but we could equally consider the defensive aspect by letting the absorbing state be the possibility that the opposition score.
+
+One apparent flaw of this analysis is that it pays no attention to the *spatial* dynamics of the game.
+But it would be fairly straightforward to develop the analysis articulated here to also take into account a player's location on the pitch when he plays the ball.
+Rather than simply representing the players, the nodes' state space could instead be the Cartesian product of players with some set of discrete areas of the pitch, say $\{\Delta, \Phi, \Xi \}$ could represent the Defensive Half, Attacking Half, and Attacking Penalty Area respectively.
+One could simply add the spatial indicator to the player name at the preprocessing stage, and the rest of the analysis would then proceed as described here.
+
+But the major weakness is one inherent in the data itself -- although the event-log dataset offers a wealth of information when compared to the manual *notational analysis* first developed by Charles Reep in the 1950s [@RPollard2002], a convincing analysis of football will only really be possible with data that includes the positions of off-the-ball players, as well as on-the-ball events.
+Such data can be captured [eg.
+@SPettersenEtAl2014], but is not yet widely (at least openly) available.
+
+## Acknowledgments
+
 ## Bibliography
+
+## Appendix 1: Analysis in R
+
+## Appendix 2: Preprocessing in Python

### 2020-07-24 rmd/SocialNetworkAnalysis_R.Rmd
done almost

--- 2020-07-13 nb/analysis_in_R.Rmd
@@
 
-# Executive Summary
-
-# Introduction
-
 ```{r}
-library(plyr) # data manipulation @HWickham2009
-library(dplyr) # data manipulation @HWickham@2017
-library(knitr) # dynamic report generation @YXie2012
-library(igraph) # for graph/network analysis @GCsardiTNepusz2006
-library(ggraph) # graph visualization in the style of ggplot @TPederson2016a
-library(netrankr) # network centrality @DSchoch2015
-library(ggplot2) # for plotting visualizations @HWickham2007
-library(ggforce) # adds functionality to ggplot2, eg. geom_circle @TPederson2016b
-library(ggrepel) # adds 'repellent' non-overlapping labels to ggplot @KSlowikowski2016
+library(plyr) # data manipulation 
+library(dplyr) # data manipulation
+library(igraph) # for graph/network analysis 
+library(ggraph) # graph visualization in the style of ggplot
+library(netrankr) # network centrality
+library(ggplot2) # for plotting visualizations
+library(ggforce) # adds functionality to ggplot2, eg. geom_circle
+library(ggrepel) # adds 'repellent' non-overlapping labels to ggplot
 library(pheatmap) # prettier heat maps
@@
 library(docstring) # allows ?help to display function descriptions analogous to Pythonic docstrings
+```
+
+```{r}
+LoadAndScaleData <- function(national_league) {
+    #' Load preprocessed event logs for given national league.
+
+    events <- read.csv(paste0('../data/processed/', national_league, '.csv'))
+
+    x_scale <- 105
+    y_scale <- 68
+
+    events$location_x <- events$location_x/100 * x_scale
+    events$target_x <- events$target_x/100 * x_scale
+
+    events$location_y <- y_scale - (events$location_y/100 * y_scale)
+    events$target_y <- y_scale - (events$target_y/100 * y_scale)
+
+    return(events)
+}
+
+```
+
+```{r}
+LeagueResults <- function(league){
+    #' Test data integrity by calculating results and table for given national league.
+
+    teams <- sort(unique(league$team))
+
+    league_table <- matrix(0, nrow=length(teams), ncol=9)
+    rownames(league_table) <- teams
+    colnames(league_table) <- c('Pos','Pld','W','D','L','GF','GA','GD','Pts')
+
+    for (m in unique(league$matchId)){
+
+        match <- league[league$matchId==m,]
+        home_team <- unique(match[match$home_or_away=='Home',]$team)
+        away_team <- unique(match[match$home_or_away=='Away',]$team)
+        home_goals <- nrow(match[match$to_team=='Home Goal',])
+        away_goals <- nrow(match[match$to_team=='Away Goal',])
+
+        if (home_goals > away_goals){
+            league_table[home_team,'W'] <- league_table[home_team,'W'] + 1
+            league_table[away_team,'L'] <- league_table[away_team,'L'] + 1
+        } else if (away_goals > home_goals){
+            league_table[home_team,'L'] <- league_table[home_team,'L'] + 1
+            league_table[away_team,'W'] <- league_table[away_team,'W'] + 1    
+        } else if (home_goals == away_goals) {
+            league_table[home_team,'D'] <- league_table[home_team,'D'] + 1
+            league_table[away_team,'D'] <- league_table[away_team,'D'] + 1
+        }
+
+        league_table[home_team,'GF'] <- league_table[home_team,'GF'] + home_goals
+        league_table[away_team,'GF'] <- league_table[away_team,'GF'] + away_goals
+        league_table[home_team,'GA'] <- league_table[home_team,'GA'] + away_goals
+        league_table[away_team,'GA'] <- league_table[away_team,'GA'] + home_goals
+
+        for (team in c(home_team, away_team)){
+            league_table[team,'GD'] <- league_table[team,'GF'] - league_table[team,'GA']
+            league_table[team,'Pts'] <- 3*league_table[team,'W'] + 1*league_table[team,'D']
+            league_table[team,'Pld'] <- league_table[team,'Pld'] + 1
+        }
+
+    }
+
+    league_table <- league_table[order(-league_table[,'Pts']),]
+
+    league_table[,'Pos'] <- 1:20
+
+    return(league_table)
+}
 ```
@@
 
-# Data
-
-```{r}
-LoadAndScaleData <- function(national_league) {
-    #' Load preprocessed event logs for given national league.
-
-    events <- read.csv(paste0('../data/processed/', national_league, '.csv'))
-
-    x_scale <- 105
-    y_scale <- 68
-
-    events$location_x <- events$location_x/100 * x_scale
-    events$target_x <- events$target_x/100 * x_scale
-
-    events$location_y <- y_scale - (events$location_y/100 * y_scale)
-    events$target_y <- y_scale - (events$target_y/100 * y_scale)
-
-    return(events)
-}
-
-```
-
-```{r}
-events <- LoadAndScaleData('England')
-```
-
-```{r}
-events[1:10,]
-```
-
-```{r}
-LeagueResults <- function(league){
-    #' Calculate results and table for given national league.
-
-    teams <- sort(unique(league$team))
-
-    home_score <- matrix(0, nrow=length(teams), ncol=length(teams))
-    colnames(home_score) <- teams
-    rownames(home_score) <- teams
-
-    away_score <- home_score
-
-    league_table <- matrix(0, nrow=length(teams), ncol=9)
-    rownames(league_table) <- teams
-    colnames(league_table) <- c('Pos','Pld','W','D','L','GF','GA','GD','Pts')
-
-    for (m in unique(league$matchId)){
-
-        match <- league[league$matchId==m,]
-        home_team <- unique(match[match$home_or_away=='Home',]$team)
-        away_team <- unique(match[match$home_or_away=='Away',]$team)
-        home_goals <- nrow(match[match$to_team=='Home Goal',])
-        away_goals <- nrow(match[match$to_team=='Away Goal',])
-
-        home_score[home_team, away_team] <- home_goals
-        away_score[home_team, away_team] <- away_goals
-
-        if (home_goals > away_goals){
-            league_table[home_team,'W'] <- league_table[home_team,'W'] + 1
-            league_table[away_team,'L'] <- league_table[away_team,'L'] + 1
-        } else if (away_goals > home_goals){
-            league_table[home_team,'L'] <- league_table[home_team,'L'] + 1
-            league_table[away_team,'W'] <- league_table[away_team,'W'] + 1    
-        } else if (home_goals == away_goals) {
-            league_table[home_team,'D'] <- league_table[home_team,'D'] + 1
-            league_table[away_team,'D'] <- league_table[away_team,'D'] + 1
-        }
-
-        league_table[home_team,'GF'] <- league_table[home_team,'GF'] + home_goals
-        league_table[away_team,'GF'] <- league_table[away_team,'GF'] + away_goals
-        league_table[home_team,'GA'] <- league_table[home_team,'GA'] + away_goals
-        league_table[away_team,'GA'] <- league_table[away_team,'GA'] + home_goals
-
-        for (team in c(home_team, away_team)){
-            league_table[team,'GD'] <- league_table[team,'GF'] - league_table[team,'GA']
-            league_table[team,'Pts'] <- 3*league_table[team,'W'] + 1*league_table[team,'D']
-            league_table[team,'Pld'] <- league_table[team,'Pld'] + 1
-        }
-
-    }
-
-    league_table <- league_table[order(-league_table[,'Pts']),]
-
-    league_table[,'Pos'] <- 1:20
-
-    return(league_table)
-}
-```
-
-```{r}
-prem <- LeagueResults(events)
-```
-
-```{r}
-pl <- as.data.frame(prem)
-```
-
-```{r}
-pl['Team'] <- rownames(pl)
-```
-
-```{r}
-pl2 <- as.data.frame(pl[,c(10,2:9)])
-```
-
-```{r}
-rownames(pl2) <- c()
-```
-
-```{r}
-pl2[1:5,]
-```
-
-```{r}
-prem[1:5,]
-```
-
-```{r}
-NationalResults <- function(nation) {
-
-    data <- LoadAndScaleData(nation)
-    results <- LeagueResults(data)
-    return(results)
-}
-
-```
-
-```{r}
-correct_res <- read.csv('../data/league_results/England17-18.csv')
-```
-
-```{r}
-correct_res
-```
-
-```{r}
-france <- NationalResults('France')
-germany <- NationalResults('Germany')
-italy <- NationalResults('Italy')
-spain <- NationalResults('Spain')
-```
-
-```{r}
-premiership <- NationalResults('England')
-```
-
 ```{r}
@@
         geom_point(data = game_events, 
-               aes(location_x , location_y, fill=eventName, shape=eventName), pch=21, alpha=0.6, size=2 ) + 
+               aes(location_x , location_y, fill=eventName, shape=eventName), pch=21, alpha=1, size=2 ) + 
         geom_point(data=game_events[game_events$subEventName=='Goal',], 
                    aes(location_x, location_y), shape=13, size=5) +
-        geom_label_repel(data=game_events[game_events$subEventName=='Goal',], 
-                         aes(location_x, location_y, label = paste0(source,'(',matchPeriod,' ',time,')')), label.padding=0.1, size=2.3, alpha=1) +
+#         geom_label_repel(data=game_events[game_events$subEventName=='Goal',], 
+#                          aes(location_x, location_y, label = paste0(source,'(',matchPeriod,' ',time,')')), label.padding=0.1, size=2.3, alpha=1) +
         theme(legend.direction='horizontal', legend.position=c(0.5,0)) +
         scale_fill_manual(values=c("red", "black", 'white', "yellow", 'blue'), name='Event Type') +
-        ggtitle(paste('Event Map:', team_name,'playing',home_or_away,'versus',opposition,'(',team_score,'-',opposition_score,')'))
+        ggtitle(paste('Event Map:', team_name,home_or_away,'versus',opposition,'(',team_score,'-',opposition_score,')'))
+
+    p$figname <- paste0('EventMap',home_or_away)
 
@@
 }
-```
-
-```{r}
-#away vs Crystal Palace
-match_id <- '2500032'
-```
-
-```{r}
-# vs man city
-match_id <- '2499943'
-```
-
-```{r}
-ShowMatchEvents(events, match_id, home_or_away='Home')
-ShowMatchEvents(events, match_id, home_or_away='Away', flip=T)
 ```
@@
                      aes(x=location_x, y=location_y, xend=target_x, yend=target_y, color=subEventName), 
-                     alpha=0.6, arrow = arrow(length = unit(0.1,"cm"))) +
-        geom_label_repel(data=game[game$subEventName=='Goal',], 
-                     aes(location_x, location_y, label = paste0(source,'(',matchPeriod,' ',time,')')), label.padding=0.1, size=2.3, alpha=1) + 
+                     alpha=1, arrow = arrow(length = unit(0.1,"cm"))) +
+#         geom_label_repel(data=game[game$subEventName=='Goal',], 
+#                      aes(location_x, location_y, label = paste0(source,'(',matchPeriod,' ',time,')')), label.padding=0.1, size=2.3, alpha=1) + 
         geom_segment(data=na.exclude(shots), 
                  aes(x=location_x, y=location_y, xend=target_x, yend=target_y, color=subEventName), 
-                 alpha=0.7, arrow = arrow(length = unit(0.1,"cm"))) +
+                 alpha=1, arrow = arrow(length = unit(0.1,"cm"))) +
         theme(legend.position=c(0.5,-0.01), legend.direction='horizontal',
@@
         scale_color_manual(values=c("red", "black", 'blue', 'orange', "yellow", 'grey', 'purple', 'white','brown'), name='Pass Type') +
-        ggtitle(paste('Pass Map:',team_name,'playing',home_or_away,'versus',opposition,'(',team_score,'-',opposition_score,')'))
+        ggtitle(paste('Pass Map:',team_name,home_or_away,'versus',opposition,'(',team_score,'-',opposition_score,')'))
+
+    p$figname <- paste0('PassMap',home_or_away)
 
@@
 ```{r}
-ShowPassesAndShots(events, match_id, home_or_away='Home')
-ShowPassesAndShots(events, match_id, home_or_away='Away', flip=T)
+GoalSeqs <- function(events, match_id) {
+    match <- events[events$matchId==match_id,]
+    goals <- match[match$subEventName=='Goal',]
+    goal_seqs <- unique(goals$possession)   
+    return(goal_seqs)
+}
 ```
 
-# Methodology
+```{r}
+TabulateSequence <- function(events, possession_sequence) {
+    # Return a markdown table starting from the row before a possession sequence begins, and continuing until it ends.
+
+    start <- min(events[events$possession == possession_sequence,]$X)
+    stop <- max(events[events$possession == possession_sequence,]$X) + 1
+    table <- events[start:stop,]
+    df <- data.frame(table$matchPeriod, table$time, table$team, table$source, table$subEventName)
+    names(df) <- c('Half', 'Time', 'Team', 'Event Description', 'Player')
+
+    return(df)
+}
+```
+
+```{r}
+SequenceOnPitch <- function(events, possession_sequence){
+    #' Draw Possession Sequence on Pitch
+
+    data <- events[events$possession == possession_sequence,]
+
+    p <- DrawPitch(theme='grass') + 
+        geom_label_repel(data = data, 
+                         aes(location_x, location_y, label = paste(source,time)), 
+                         label.padding=0.1, size=2.3, alpha=1) +
+        geom_curve(data = data, 
+                   aes(x = location_x, xend = target_x, 
+                       y = location_y, yend = target_y*.99, col = X), 
+                   show.legend=FALSE, size=1, alpha = 1) +
+        geom_point(data = data, 
+                   aes(location_x , location_y, shape=eventName), size=2) + 
+        ggtitle(paste0(nrow(data),'-part ', unique(data$team)[1], 
+                       ' Possession Sequence \nending with ', 
+                       data[nrow(data),]$subEventName, ' by ', 
+                       data[nrow(data),]$source))  +
+        theme(legend.position=c(0.5,0.08), 
+              legend.direction='horizontal', 
+              legend.title=element_blank())
+
+    return(p)
+}
+```
+
+```{r}
+SequenceGraph <- function(events, possession_sequence){
+    #' Return Graph of Possession Sequence
+
+    data <- events[events$possession == possession_sequence,]
+    data <- data[data$source!='' & data$target!='nan',]
+    nodes <- unique(c(as.character(data$source),as.character(data$target)))
+    edges <- data.frame(data$source, data$target)
+    g <- graph_from_data_frame(d=edges, vertices=nodes, directed=TRUE)
+    g$id <- possession_sequence
+    g$data <- data
+    g$team <- as.character(unique(data$team))
+
+    return(g)
+}
+```
+
+```{r}
+VisualizeGraph <- function(possession_sequence_graph) {
+    #' Visualize Possession Sequence as Linear Graph with Looping Edges
+
+    g <- possession_sequence_graph
+    data <- possession_sequence_graph$data
+
+    visualization <- ggraph(g, 'linear') + 
+        geom_edge_arc(aes(color=data$eventName), 
+                      arrow=arrow(length=unit(4,'mm')), 
+                      fold=F,
+                     width=1) +
+        geom_edge_loop(aes(color=data$eventName),
+                      width=1) +
+        geom_node_point(color='black', 
+                        size=2,
+                        alpha=0.5) +
+        geom_node_text(aes(label = name), 
+                        repel=T, 
+                        angle=90, hjust=2, ) +
+        scale_edge_colour_manual(
+            values=c('indianred3', 'wheat4', 'grey', 'grey30',
+                    'red','blue','green','orange','purple','brown','pink'
+                    ),
+            name='Event Type') +
+        theme_void()
+
+    return(visualization)
+
+}
+```
 
@@
 
-    return (DrawPitch() +
-        geom_segment(data=team[team$passcount>=lower_threshold,], size=1, colour=team_colour,
-                     aes(x=source_x, y=source_y, xend=x_mean, yend=y_mean, alpha=passcount)) +
-        geom_segment(data=team[team$passcount>=high_threshold,], size=1.5, colour='black',
-                     aes(x=source_x, y=source_y, xend=x_mean, yend=y_mean, alpha=passcount)) +
+    p <- (DrawPitch() +
+        geom_segment(data=team[team$passcount>=lower_threshold,], 
+                     size=1, colour=team_colour,
+                     aes(x=source_x, y=source_y, 
+                         xend=x_mean, yend=y_mean, alpha=passcount)) +
+        geom_segment(data=team[team$passcount>=high_threshold,], 
+                     size=1.5, colour='black',
+                     aes(x=source_x, y=source_y, 
+                         xend=x_mean, yend=y_mean, alpha=passcount)) +
         geom_label_repel(data = mean_positions[mean_positions$team==team_name,], 
-                         aes(x_mean, y_mean, label = target), label.padding=0.1, size=2.3, alpha=1) +
-        geom_point(data=team, aes(x_mean, y_mean,), fill=team_colour, colour='black', pch=21, size=3) +
-        ggtitle(paste('Passing Network:',team_name,'playing',home_or_away,'versus',opposition,'(',team_score,'-',opposition_score,')')) +
+                         aes(x_mean, y_mean, label = target), 
+                         label.padding=0.5, size=4, alpha=0.8) +
+        geom_point(data=team, aes(x_mean, y_mean,), 
+                   fill=team_colour, colour='black', pch=21, size=3) +
+        ggtitle(paste('Passing Network:',team_name,home_or_away,
+                      'versus',opposition,'(',team_score,'-',opposition_score,')')) +
         theme(legend.position=c(0.5,0.07), legend.direction='horizontal'))
 
+    p$figname <- paste0('PassingNetwork',home_or_away)
+
+    return(p)
 }
-```
-
-```{r}
-PassNetwork(events, match_id, team_colour='red')
-PassNetwork(events, match_id, team_colour='blue', home_or_away='Away', flip=T)
-```
-
-```{r}
-2499943
-```
-
-```{r}
-
 ```
@@
 ```{r}
+MatrixHeatMap <- function(matrix, color, number_format, title){
+    #' Show Matrix as HeatMap
+
+    hm <- pheatmap(matrix,color=color,
+                   cluster_rows=F,cluster_cols=F,legend=F,
+                    display_numbers=T,number_format=number_format,
+                   fontsize_number=9,angle_col='315',
+                    main=title)
+
+    return(hm)
+}
+```
+
+```{r}
 ExpectedScoringTime <- function(transition_matrix){
@@
     max_steps <- 10000
-#     transition_matrix <- TransitionMatrix(GameGraph(events, match_id))
     hitting_times <- rep(0,(ncol(transition_matrix)-2))
@@
 
-        state_probabilities <- matrix(NA,nrow=max_steps+1,ncol=ncol(transition_matrix),dimnames=list(0:max_steps,(ncol(transition_matrix)-1):0))
+        state_probabilities <- matrix(NA,
+                                      nrow=max_steps+1,
+                                      ncol=ncol(transition_matrix),
+                        dimnames=list(0:max_steps,(ncol(transition_matrix)-1):0))
         vector <- rep(0,ncol(transition_matrix))
@@
 SeasonAdjacencyMatrix <- function(events, team) {
+    #' Return Adjacency Matrix for Team's full season.
 
@@
 MarkovExpectedGoals <- function(events, season_adjacency_matrix) {
+    #' Calculate MXG for each player from team's Season Adjacency Matrix.
 
@@
     seasonXST <- ExpectedScoringTime(season_transition)
-    avg_events_per_game <- nrow(events)/380
-    mXG <- sort(avg_events_per_game/seasonXST, decreasing=T)
+    seasonMXG <- 1/seasonXST
+    stdize <- seasonMXG/(mean(seasonMXG))
+    avg_events_per_game <- sum(season_adjacency_matrix)/38
+    scale <- stdize * avg_events_per_game - avg_events_per_game
 
-    return(mXG)
+    return(scale)
 }
@@
 ```{r}
-GoalProbabilities <- function(season_adjacency_matrix) {
+MXGTable <- function(events, team){
+    #' Return dataframe with Degree, Loops, Goals, and MXG for each player in team.
 
-    season_graph <- graph_from_adjacency_matrix(season_adjacency_matrix)
-    season_transition <- TransitionMatrix(season_graph)
-    goal_probs <- sort(season_transition[,'Goal'],T)[2:nrow(season_transition)] * 100
+    Team <- team
+    m <- SeasonAdjacencyMatrix(events,Team)
+    Degree <- rowSums(m)[1:(nrow(m)-2)]
+    Loops <- diag(m)[1:(nrow(m)-2)]
+    Goals <- m[,ncol(m)][1:(nrow(m)-2)]
+    Player <- rownames(m)[1:(nrow(m)-2)]
+    MXG <- MarkovExpectedGoals(events, m)[1:(nrow(m)-2)]
+    df <- data.frame(Player, Team, MXG, Goals, Loops, Degree)   
+    rownames(df) <- c()
 
-    return(goal_probs)
+    return(df)
 }
-```
-
-```{r}
-liverpool_adj <- SeasonAdjacencyMatrix(events,'Liverpool')
 
@@
 ```{r}
-mXG <- MarkovExpectedGoals(events, liverpool_adj)
-```
+CalculateMXG <- function() {
+    #' Get MXG and other stats for each player in all leagues.
 
-```{r}
-mXG
-```
+    mxgtable_df <- data.frame(Player = character(),
+                    Team = character(),
+                    MXG = double(),
+                    Goals = integer(),
+                    Loops = integer(),
+                    Degree = integer())
 
-```{r}
-GoalProbabilities(liverpool_adj)
-```
+    leagues = c('England','France','Germany','Italy','Spain')
+    for (league in leagues){
+        events <- LoadAndScaleData(league)
+        for (team in unique(events$team)){
+            mxgtable_df <- rbind(mxgtable_df, MXGTable(events,team))
 
-```{r}
-sort(liverpool_adj[,'Goal'],T)
-```
+        }
+    }
 
-```{r}
-
-```
-
-```{r}
-
-```
-
-```{r}
-
-```
-
-```{r}
-g <- GameGraph(events, match_id, home_or_away='Home')
-```
-
-```{r}
-MatrixHeatMap <- function(matrix, number_format='%.d', title='Matrix Heatmap'){
-    #' Show Matrix as HeatMap
-
-    hm <- pheatmap(matrix,color=brewer.pal(15,'YlGnBu'),cluster_rows=F,cluster_cols=F,legend=F,
-                    display_numbers=T,number_format=number_format,fontsize_number=9,angle_col='315',
-                    main=title)
-
-    return(hm)
-}
-```
-
-```{r}
-MatrixHeatMap(
-    AdjacencyMatrix(g)[1:11,1:11]
-
-)
-```
-
-```{r}
-MatrixHeatMap(
-    AdjacencyMatrix(
-        GameGraph(events, match_id, home_or_away='Away'))[1:11,1:11]
-)
-```
-
-```{r}
-MatrixHeatMap(
-    TransitionMatrix(
-        GameGraph(events, match_id, home_or_away='Home')), 
-    number_format='%.2f',
-    title=''
-)
-```
-
-```{r}
-MatrixHeatMap(
-    TransitionMatrix(
-        GameGraph(events, match_id, home_or_away='Away')), 
-    number_format='%.2f',
-    title=''
-)
-```
-
-```{r}
-GoalSeqs <- function(events, match_id) {
-    match <- events[events$matchId==match_id,]
-    goals <- match[match$subEventName=='Goal',]
-    goal_seqs <- unique(goals$possession)   
-    return(goal_seqs)
-}
-```
-
-```{r}
-TabulateSequence <- function(events, possession_sequence) {
-    # Return a markdown table starting from the row before a possession sequence begins, and continuing until it ends.
-
-    start <- min(events[events$possession == possession_sequence,]$X)
-    stop <- max(events[events$possession == possession_sequence,]$X) + 1
-    table <- events[start:stop,]
-    df <- data.frame(table$matchPeriod, table$time, table$team, table$source, table$subEventName)
-    names(df) <- c('Half', 'Time', 'Team', 'Event Description', 'Player')
-
-    return(kable(df))
-}
-```
-
-```{r}
-sequence <- '2500032-91-Liv-0'
-```
-
-```{r}
-TabulateSequence(events, sequence)
-```
-
-```{r}
-SequenceOnPitch <- function(events, possession_sequence){
-    #' Draw Possession Sequence on Pitch
-
-    data <- events[events$possession == possession_sequence,]
-
-    p <- DrawPitch(theme='grass') + 
-        geom_label_repel(data = data, aes(location_x, location_y, label = paste(source,time)), label.padding=0.1, size=2.3, alpha=1) +
-        geom_curve(data = data, aes(x = location_x, xend = target_x, y = location_y, yend = target_y*.99, col = X), 
-                   show.legend=FALSE, size=1, alpha = 1) +
-        geom_point(data = data, aes(location_x , location_y, shape=eventName), size=2) + 
-        ggtitle(paste0(nrow(data),'-part ', unique(data$team)[1], ' Possession Sequence ending with ', 
-                       data[nrow(data),]$subEventName, ' by ', data[nrow(data),]$source))  +
-        theme(legend.position=c(0.5,0.08), legend.direction='horizontal', legend.title=element_blank())
-
-    return(p)
-}
-```
-
-```{r}
-SequenceOnPitch(events, sequence)
-```
-
-youtube clip
-<img src="../figures/youtube_clip.gif" width='700'>
-
-```{r}
-SequenceGraph <- function(events, possession_sequence){
-    #' Return Graph of Possession Sequence
-
-    data <- events[events$possession == possession_sequence,]
-    data <- data[data$source!='' & data$target!='nan',]
-    nodes <- unique(c(as.character(data$source),as.character(data$target)))
-    edges <- data.frame(data$source, data$target)
-    g <- graph_from_data_frame(d=edges, vertices=nodes, directed=TRUE)
-    g$id <- possession_sequence
-    g$data <- data
-    g$team <- as.character(unique(data$team))
-
-    return(g)
-}
-```
-
-```{r}
-VisualizeGraph <- function(possession_sequence_graph) {
-    #' Visualize Possession Sequence as Linear Graph with Looping Edges
-
-    g <- possession_sequence_graph
-    data <- possession_sequence_graph$data
-
-    visualization <- ggraph(g, 'linear') + 
-        geom_edge_arc(aes(color=data$eventName), 
-                      arrow=arrow(length=unit(4,'mm')), 
-                      fold=F,
-                     width=1) +
-        geom_edge_loop(aes(color=data$eventName),
-                      width=1) +
-        geom_node_point(color='black', 
-                        size=2,
-                        alpha=0.5) +
-        geom_node_text(aes(label = name), 
-                        repel=T, 
-                        angle=90, hjust=2, ) +
-        scale_edge_colour_manual(
-            values=c('indianred3', 'wheat4', 'grey', 'grey30',
-                    'red','blue','green','orange','purple','brown','pink'
-                    ),
-            name='Event Type') +
-        theme_void()
-
-    return(visualization)
+    write.csv(mxgtable_df[order(-mxgtable_df[,'MXG']),], file='../tables/MXG.csv', row.names=F)
 
@@
 ```{r}
-g <- SequenceGraph(events, sequence)
-```
+TallyEventTypes <- function(events, eventType=F){
+    #' Return dataframe with tallies for specified type.
 
-```{r}
-VisualizeGraph(g)
-```
+    if (eventType!=F){
+        events <- events[events$eventName==eventType,]
+        df <- data.frame(table(events$subEventName))
+        df <- df[df$Freq>0,]
+    } else {
+        df <- data.frame(table(events$eventName))
+    }
 
-```{r}
-GetGraphMetrics <- function(possession_graph) {
-    #' Return basic metrics for given Graph
+    df <- df[order(-df$Freq),]
+    colnames(df)[1:2] <- c('Description','Frequency')
 
-    g <- possession_graph
-    path_metrics <- c(g$id, g$team, length(V(g)), length(E(g)), 
-                      diameter(g), mean_distance(g), edge_density(g, loops=T), 
-                      girth(g)$girth, transitivity(g), authority_score(g)$value, hub_score(g)$value)
-    names(path_metrics) <- c('Label', 'Team', 'Nodes', 'Edges', 'Diameter', 'Mean_Distance', 
-                             'Edge_Density', 'Girth', 'Transitivity', 'Authority_Score', 'Hub_Score')
+    df$Description <- as.character(df$Description)
 
-    return(path_metrics)
-}
-```
+    total <- sum(df$Frequency)
+    df[nrow(df)+1,] <- c('**Total**',total)
+    df$Frequency <- as.numeric(df$Frequency)
 
-```{r}
-GetNodeMetrics <- function(possession_graph) {
-    #' Return metrics for each Node of given Graph
+    df$Pct <- df$Frequency / total * 100
+    rownames(df) <- c()
+    colnames(df)[3] <- '%'
 
-    g <- possession_graph
-    df <- data.frame(page_rank(g)$vector)
-    names(df) <- 'pagerank'
-    df$in_degree <- degree(g, mode='in')
-    df$out_degree <- degree(g, mode='out')
-    df$eccentricity <- eccentricity(g)
-    df$closeness <- closeness(g)
-    df$betweenness <- betweenness(g)
-    df$norm_betweenness <- betweenness(g, normalized=T)
-    df$alpha_centrality <- alpha_centrality(g)
-    df$power_centrality <- power_centrality(g)
-    df$eigen_centrality <- eigen_centrality(g)$vector
-    df$authority <- authority_score(g)$vector
-    df$hub <- hub_score(g)$vector
+    df[,3] <- formatC(signif(df[,3],digits=3), digits=3,format="fg", flag="#")
+
+    df[,'Frequency (%)'] <- paste0(as.character(
+                            prettyNum(df$Frequency, big.mark=",", scientific=F)),
+                           ' (', as.character(df[,3]),'%)')
+
+    df <- df[,c(1,4)]
 
@@
 ```{r}
-GetEdgeMetrics <- function(possession_graph) {
-    #' Return metrics for each Edge of given Graph
+ExploratoryTallies <- function(){
+    #' Get exploratory tallies.
 
-    g <- possession_graph
-    df <- data.frame(as_edgelist(g))
-    names(df) <- c('Source','Target')
-    df$sequence <- 1:length(E(g))
-    df$edge_betweenness <- edge_betweenness(g)
+    events <- LoadAndScaleData('England')
 
+    leagues = c('England','France','Germany','Italy','Spain')
+    for (league in leagues[2:5]){
+        events <- rbind(events, LoadAndScaleData(league))
+    }
+
+    types <- TallyEventTypes(events)
+    passes <- TallyEventTypes(events,'Pass')
+    shots <- TallyEventTypes(events,'Shot')
+
+    write.csv(types,'../tables/EventTypeTally.csv',row.names=F)
+    write.csv(passes,'../tables/PassTally.csv',row.names=F)
+    write.csv(shots,'../tables/ShotTally.csv',row.names=F)
+
+}
+```
+
+```{r}
+SavePairedFigures <- function(league, match_id){
+    #' Generate and Save Paired Figures.
+
+    events <- LoadAndScaleData(league)
+
+    EventMapHome <- ShowMatchEvents(events, match_id)
+    EventMapAway <- ShowMatchEvents(events, match_id, home_or_away='Away', flip=T)
+
+    PassMapHome <- ShowPassesAndShots(events, match_id)
+    PassMapAway <- ShowPassesAndShots(events, match_id, home_or_away='Away', flip=T)
+
+    PassingNetworkHome <- PassNetwork(events, match_id, team_colour='blue')
+    PassingNetworkAway <- PassNetwork(events, match_id, home_or_away='Away', 
+                                      team_colour='red',flip=T)
+
+    pairedFigures <- list(EventMapHome, EventMapAway,
+                      PassMapHome, PassMapAway,
+                      PassingNetworkHome, PassingNetworkAway)
+
+    for (fig in pairedFigures){
+        png(filename=paste0('../figures/',fig$figname,'.png'),
+            width=800, 
+            height=650,
+            units='px',
+           pointsize=4,
+           res=140)
+        print(fig)
+        dev.off()
+
+    }   
+}
+
+```
+
+```{r}
+CountSequences <- function(events) {
+    #' Count Possession Sequences and return DataFrame.
+
+    poss_seqs <- data.frame(table(events$possession))
+    df_psq <- data.frame(table(poss_seqs$Freq))
+    goals_seqs <- events[events$subEventName == 'Goal',]$possession
+    goals_seqs_events <- events[events$possession %in% goals_seqs,]
+    df_gpsq0 <- data.frame(table(goals_seqs_events$possession))
+    df_gpsq <- df_gpsq0[df_gpsq$Freq>0,]
+    df_g <- data.frame(table(df_gpsq$Freq))
+
+    m <- rbind(1:10, df_psq$Freq[1:10], df_g$Freq[1:10])
+    mdf <- data.frame(m)
+    rownames(mdf) <- c('Sequence Length', 'Frequency', 'Goal Scoring')
+
+    total_goals <- sum(df_g$Freq)
+    total_seqs <- sum(df_psq$Freq)
+    final <- c('**Total**', total_seqs, total_goals)
+    mdf <- cbind(mdf,final)
+    colnames(mdf) <- c()
+
+    df <- t(mdf)
+    df[,2] <- prettyNum(df[,2], big.mark=",", scientific=F)
+    df[,3] <- prettyNum(df[,3], big.mark=",", scientific=F)
     return(df)
@@
 ```{r}
-GetGraphMetrics(g)
+PossessionSequence <- function(league, sequence){
+    #' Generate Figures and Table for given Possession Sequence.
+
+    events <- LoadAndScaleData(league)
+    write.csv(TabulateSequence(events, sequence),'../tables/PossessionSequence.csv',row.names=F)
+    png(filename=paste0('../figures/PossessionSequence.png'),
+        width=800, 
+        height=600,
+        units='px',
+       pointsize=12,
+       res=140)
+    print(SequenceOnPitch(events, sequence))
+    dev.off()
+
+    g <- SequenceGraph(events, sequence)
+    png(filename=paste0('../figures/PossessionGraph.png'),
+        width=800, 
+        height=600,
+        units='px',
+       pointsize=12,
+       res=140)
+    print(VisualizeGraph(g))
+    dev.off()
+}
 ```
@@
 ```{r}
-GetNodeMetrics(match_graph)
+Top20MXG <- function(){
+    #' Save CSV with top 20 players by MXG score.
+
+    d <- read.csv('../tables/MXG.csv')
+    top <- d[d$Degree>1000,]
+    rownames(top) <- c()
+    top['Loops/Degree'] <- top$Loops/top$Degree
+    top['Goals/Degree'] <- top$Goals/(top$Degree)
+    t <- top[c('Player','Team','Degree','Goals',
+               'Goals/Degree','Loops','Loops/Degree','MXG')]
+    write.csv(t[1:20,], '../tables/2.csv',row.names=F)
+}
+
 ```
@@
 ```{r}
-GetEdgeMetrics(g)
+SaveMatrixHeatmaps <- function(league, match_id, team_name){
+    #' Save Heatmaps for Adjacency and Transition Matrices.
+
+    events <- LoadAndScaleData(league)
+    g <- GameGraph(events, match_id, team_name)
+    adj_m <- AdjacencyMatrix(g)
+    tr_m <- TransitionMatrix(g)
+
+    matrix_cols <- c('darkgrey',colorRampPalette(c('white','red'))(35))
+    adj_m_hm <- MatrixHeatMap(adj_m[1:11,1:11], color=matrix_cols, number_format='%.0f',
+                  title=paste('Adjacency Matrix for',g$team,'(Starting XI)\n',
+                              g$status,'vs.',g$opposition))
+    adj_m_hm$type <- 'Adjacency'
+
+    tr_m_hm <- MatrixHeatMap(tr_m, color=matrix_cols[2:36], number_format='%.2f',
+                  title=paste('Transition Matrix for',g$team,g$status,'vs.',g$opposition))
+    tr_m_hm$type <- 'Transition'
+
+    hm_vec <- list(adj_m_hm, tr_m_hm)
+
+    for (hm in hm_vec){
+        png(filename=paste0('../figures/',hm$type,'MatrixAway.png'),
+            width=1600, 
+            height=1600,
+            units='px',
+           pointsize=12,
+           res=240)
+
+        print(hm)
+
+        dev.off()   
+    }
+
+}
 ```
@@
 ```{r}
-sequence2 <- '2499858-18-Liv-5'
+GenerateTablesAndFigures <- function(league, match_id, possession_seq, team){
+    #' Generate Tables and Figures for Paper.
+
+    ExploratoryTallies()
+    SavePairedFigures(league, match_id)
+    PossessionSequence(league, possession_seq)
+    SaveMatrixHeatmaps(league, match_id, team)
+    CalculateMXG()
+    Top20MXG()
+}
 ```
@@
 ```{r}
-TabulateSequence(sequence2)
+GenerateTablesAndFigures('England','2500032','2500032-91-Liv-0','Liverpool')
 ```
-
-```{r}
-SequenceOnPitch(sequence2)
-```
-
-<img src='../figures/mane_vs_stoke.gif'>
-
-```{r}
-g2 <- SequenceGraph(sequence2)
-VisualizeGraph(g2)
-```
-
-```{r}
-GetGraphMetrics(g2)
-```
-
-```{r}
-sequence3 <-  '2499889-18-Liv-0'
-```
-
-```{r}
-SequenceOnPitch(sequence3)
-```
-
-```{r}
-g3 <- SequenceGraph(sequence3)
-VisualizeGraph(g3)
-```
-
-```{r}
-GetGraphMetrics(g3)
-```
-
-<img src="../figures/youtube_clip2.gif">
-
-```{r}
-
-```

### 2020-07-24 rmd/draft.Rmd
done almost

@@
   kernelspec:
-    display_name: R
-    language: R
-    name: ir
+    display_name: Python 3
+    language: python
+    name: python3
 ---
+
+<small>
 
@@
 
-Association football is the world's most popular game and a multibillion dollar industry.
-Statistical analysis of the sport was pioneered more than fifty years ago, but has only recently begun to make a more widespread impact.
-In this paper I explain how Social Network Analysis can provide the tools to model a football team's ball-possession as a dynamic probabilistic system, and hence how we can quantify the effect of each player on the team's goal-scoring capability.
+Association football is the world's most popular sport [@IPalacios-Huerta2004] and a multibillion dollar industry [@Deloitte2019]. 
+Statistical analysis of the sport was pioneered more than fifty years ago [@CReepBenjamin1968], but has only recently begun to make a more widespread impact [@PCintiaEtAl2015].
+Such analysis is of potential interest to coaches and scouts working professionally within the game, journalists and media commentators explaining the game, and gamblers and bookmakers putting money on the game.
+In this paper I explain how Social Network Analysis [@SWassermanEtAl1994] can provide the tools to model a football team's ball-possession as a dynamic probabilistic system, and hence how we can quantify the contribution of each player to the team's goal-scoring capability.
+
+</small>
 
@@
 In this paper, I develop a rigorous account of a football match as a stochastic social network, the graph of which is generated by the movement of the ball from player to player.
-I then show how the adjacency matrix of a player, considered as a node in a football match's possession sequence graph, can be used to quantify his contribution to the likelihood of the team scoring.
-Finally, I discuss the potential and pitfalls of this new metric, which I call *Relative Markovian Expected Goals*.
+I then show how the adjacency matrix of a player, considered as a node in a football match's possession sequence graph, can be used to quantify his contribution to the likelihood of the team scoring, which I call *Markov Expected Goals*.
+Finally, I discuss the potential and pitfalls of this new metric.
 
@@
 
-The dataset is provided by WyScout, one of a growing number of companies in the football analytics industry, and is available from FigShare.com [@LPappalardoMassucco2020].
-It includes more than three million *event-logs* for every match from the five major European leagues, in JSON format (see Figure 1), with  the details of players, teams, and informative tags (such as the crucial 'Goal') numerically encoded.
+The dataset is provided by WyScout, one of a growing number of companies in the football analytics industry, and is available from FigShare.com [@LPappalardoMassucco2019].
+It includes more than three million *event-logs* for every 2017-18 match from the five major European leagues, in JSON format (see Figure 1), with  the details of players, teams, and informative tags (such as the crucial 'Goal') numerically encoded.
 Each event is associated with an event type (such as 'Shot', 'Pass' or 'Free Kick'), identified with a player, and includes a nested array of the locational *positions* of the event itself and its target.
+
+<details open>
+<summary>Figure 1</summary>
+<figure>
+----
+
+```{python echo=False, wrap=False}
+
+with open('../figures/ExampleEvent.json','r') as f:
+	json_event = f.read()
+print(json_event)
+
+```
+
+<figcaption>
+Figure 1: Example of Event-Log in JSON Format
+</figcaption>
+----
+
+</figure>
+</details>
 
@@
 
-We can then visualize the location of the event-log data points in the context of a diagrammatic football pitch, to allow for an immediate and accessible view of any particular possession sequence (eg.
-Table 2, Figure 3, Figure 4).
+We can then visualize the location of the event-log data points in the context of a diagrammatic football pitch, to allow for an immediate and accessible view of the events of any match (see Figures 2 & 3), or particular possession sequence (see Figure 4).
 
-Figure 3: Pitchmap showing Possession Sequence
+<details open>
+<summary>Table 1</summary>
+<figure>
+----
+
+<figcaption>
+Table 1: Frequency of Event Types
+</figcaption>
+
+```{python echo=False, wrap=False, results='table'}
+
+import pandas as pd
+
+table1 = pd.read_csv('../tables/EventTypeTally.csv',
+	index_col=0,dtype=str)
+print(table1.to_markdown(floatfmt="#.3g"))
+
+```
+
+----
+
+</figure>
+</details>
+
+<details open>
+<summary>Table 2</summary>
+<figure>
+----
+
+<figcaption>
+Table 2: Frequency of Pass Types
+</figcaption>
+
+```{python echo=False, wrap=False, results='table'}
+
+import pandas as pd
+
+table1 = pd.read_csv('../tables/PassTally.csv',
+	index_col=0,dtype=str)
+print(table1.T.to_markdown(floatfmt="#.3g"))
+
+```
+
+----
+
+</figure>
+</details>
+
+<details open>
+<summary>Figure 2</summary>
+<figure>
+----
+
+<table><tr><td>
+<img src="../figures/EventMapHome.png">
+</td><td>
+<img src="../figures/EventMapAway.png">
+</td></tr></table>
+
+<figcaption>
+Figure 2: Events for Home and Away Team
+</figcaption>
+----
+
+</figure>
+</details>
+
+<details open>
+<summary>Figure 3</summary>
+<figure>
+----
+
+<table><tr><td>
+<img src="../figures/PassMapHome.png">
+</td><td>
+<img src="../figures/PassMapAway.png">
+</td></tr></table>
+
+<figcaption>
+Figure 3: Passes and Shots for Home and Away Team
+</figcaption>
+
+----
+
+</figure>
+</details>
+
+<details open>
+<summary>Figure 4</summary>
+<figure>
+----
+
 <img src="../figures/PossessionSequence.png">
+<figcaption>Figure 4: Pitchmap showing Possession Sequence</figcaption>
 
-We can compare this to a tabulated account of all events from just before the beginning of the possession sequence until its end, we see that our algorithm has successfully handled the case where an opposition player's unsuccessful challenge interrupts the sequence of events associated with the team in possession.
+----
 
-Table 2: Events of Possession Sequence
+</figure>
+</details>
 
-And with possession sequences that end in a goal, it is generally possible to find video highlights of the events under consideration, and confirm that the data we have does correspond to what actually happened.
+We can compare this to a tabulated account of all events from just before the beginning of the possession sequence until its end, and we see that our algorithm has successfully handled the case where an opposition player's unsuccessful challenge interrupts the sequence of events associated with the team in possession.
 
-Figure 4: YouTube Clip showing successful conclusion of possession sequence
-<img src="../figures/youtube_clip.gif">
+<details open>
+<summary>Table 1</summary>
+<figure>
+----
+
+<figcaption>
+Table 1: Events of Possession Sequence
+</figcaption>
+
+```{python echo=False, wrap=False, results='table'}
+
+import pandas as pd
+
+table1 = pd.read_csv('../tables/1.csv',index_col=0)
+print(table1.to_markdown(showindex=False))
+
+```
+
+----
+
+</figure>
+</details>
+
+With possession sequences that end in a goal, it is generally possible to find video highlights of the events under consideration, and confirm that the data we have does correspond to what actually happened.
+
+<details open>
+<summary>Figure 5</summary>
+<figure>
+----
+
+<img src="../figures/YouTubeClip.gif" width=800>
+<figcaption>Figure 5: YouTube Clip showing successful conclusion of possession sequence
+</figcaption>
+
+----
+
+</figure>
+</details>
 
@@
 
-Social network analysis applies the concepts of mathematical *graph theory*.
+Social Network Analysis applies the concepts of mathematical *graph theory*.
 
@@
          w(e_{i,j}), & e_{i,j} \in E \\
-       0, & \mathrm{otherwise}
+       0, & \mathrm{otherwise.}
     \end{array}
@@
 
-Figure 5
+<details open>
+<summary>Figure 6</summary>
+<figure>
+----
+
 <img src="../figures/PossessionGraph.png">
+<figcaption>Figure 6: Graph of Possession Sequence leading to Goal
+</figcaption>
+
+----
+
+</figure>
+</details>
 
 There are several different 'weights' we could assign an edge: the Euclidean distance moved by the ball; the polar angle; the time interval.
-However, for this particular analysis we ignore the spatial dynamics and the subtler aspects of continuous time, and focus on the network connections between players.
+However, for this particular analysis we focus simply on the network connections between players.
 
@@
 
-Figure 6
-<img src="../figures/AdjacencyMatrixHome.png">
+<details open>
+<summary>Figure 7</summary>
+<figure>
+----
+
+<img src="../figures/AdjacencyMatrixAway.png">
+
+<figcaption>
+Figure 7: Heatmap of Adjacency Matrix
+</figcaption>
+
+----
+
+</figure>
+</details>
 
@@
 This swiftly communicates a team's formation, and gives a qualitative sense of a team's strategy.
-Figure 7
+
+<details open>
+<summary>Figure 8</summary>
+<figure>
+----
+
+<table><tr><td>
 <img src="../figures/PassingNetworkHome.png">
+</td><td>
+<img src="../figures/PassingNetworkAway.png">
+</td></tr></table>
+
+<figcaption>
+Figure 8: Possession Networks for Home and Away Team
+</figcaption>
+
+----
+
+</figure>
+</details>
 
@@
 A *random variable* is a function $X: \Omega \mapsto S$.
-If $S \subseteq \mathbb{R}$ then we also have the *expectation* $\mathbb{E}[X]$ and *variance* $\mathrm{Var}[X]$, but it is acceptable to have an $S \not\subset \mathbb{R}$.
-Again, to illustrate what we mean -- the number of goals in the match could be considered as a real-valued random variable, and we could talk about the expectation and variation of the number of goals; but this is not the case for the question of which player has the ball at a particular time, though this also is a random variable.
+If $S \subseteq \mathbb{R}$ then we also have the *expectation* $\mathbb{E}[X]$ and *variance* $\mathrm{Var}[X]$.
+Again, to illustrate what we mean -- the number of goals in the match is an integer-valued random variable, and since $\mathbb{Z} \subset \mathbb{R}$ we could talk about the expectation and variation of the number of goals; but this is not the case for the question of which player has the ball at a particular time, though this also is a random variable.
 
@@
 
-Figure 8
-<img src="../figures/TransitionMatrixHome.png">
+<details open>
+<summary>Figure 9</summary>
+<figure>
+----
+
+<img src="../figures/TransitionMatrixAway.png">
+
+<figcaption>
+Figure 9: Heatmap of Transition Matrix
+</figcaption>
+
+----
+
+</figure>
+</details>
 
@@
 
+<details open>
+<summary>Table 2</summary>
+<figure>
+----
+
+<figcaption>
+Table 2: Top Twenty European Players by MXG Score
+</figcaption>
+
+```{python echo=False, wrap=False, results='table'}
+
+import pandas as pd
+
+MXGtable = pd.read_csv('../tables/2.csv',index_col=0)
+print(MXGtable.to_markdown(floatfmt="#.3g"))
+
+```
+
+----
+
+</figure>
+</details>
+
 ## Conclusion
@@
 But it would be fairly straightforward to develop the analysis articulated here to also take into account a player's location on the pitch when he plays the ball.
-Rather than simply representing the players, the nodes' state space could instead be the Cartesian product of players with some set of discrete areas of the pitch, say $\{\Delta, \Phi, \Xi \}$ could represent the Defensive Half, Attacking Half, and Attacking Penalty Area respectively.
+Rather than merely representing the players, the nodes' state space could instead be the Cartesian product of players with some set of discrete areas of the pitch, say $\{\Delta, \Phi, \Xi \}$ representing the Defensive Half, Attacking Half, and Attacking Penalty Area respectively.
 One could simply add the spatial indicator to the player name at the preprocessing stage, and the rest of the analysis would then proceed as described here.
 
-But the major weakness is one inherent in the data itself -- although the event-log dataset offers a wealth of information when compared to the manual *notational analysis* first developed by Charles Reep in the 1950s [@RPollard2002], a convincing analysis of football will only really be possible with data that includes the positions of off-the-ball players, as well as on-the-ball events.
+But the major weakness is one inherent in the data itself -- although the event-log dataset offers a wealth of information when compared to the manual *notational analysis* first developed by Charles Reep in the 1950s [@RPollard2002], a convincing analysis of football will only really be possible with data that includes the positions of off-the-ball players, as well as on-the-ball events [@RTavares2017].
 Such data can be captured [eg.
@@
 
-## Acknowledgments
+## Appendix 1: Social Network Analysis in R
+
+<details open>
+<summary>Code</summary>
+
+```{python echo=False, wrap=False, results='Rmd'}
+
+with open('../rmd/SocialNetworkAnalysis_R.Rmd','r') as f:
+	R_script = f.read()
+print(R_script)
+
+```
+
+</details>
+
+## Appendix 2: Preprocessing in Python
+
+<details open>
+<summary>Code</summary>
+
+```{python echo=False, wrap=False, results='Rmd'}
+
+with open('../rmd/Preprocessing_Python.Rmd','r') as f:
+	python_script = f.read()
+print(python_script)
+
+```
+
+</details>
 
 ## Bibliography
-
-## Appendix 1: Analysis in R
-
-## Appendix 2: Preprocessing in Python

### 2020-07-25 rmd/Preprocessing_Python.Rmd
done almost
done

--- 2020-07-10 nb/preprocessing.Rmd
@@
   jupytext:
-    formats: ipynb,Rmd
+    formats: nb//ipynb,rmd//Rmd
     text_representation:
@@
 ```{python}
+if __name__ == '__main__':
     for league in national_leagues:
@@
 ```
-
-```{python}
-
-```
-
-```{python}
-national_league='England'
-
-# load JSON event-logs
-events = load_json(f'events_{national_league}')
-
-# create dataframe from JSON event-logs
-ev_df = pd.DataFrame(events)
-
-# decode tags
-for i in range(6):
-    ev_df[f'tag{i}'] = ev_df['tags'].str[i].str['id'].map(tag_dict).str.lower()
-
-# decode player ids
-players = load_json('players')
-players_dict = dict(zip(pd.DataFrame(players).wyId,pd.DataFrame(players).shortName))
-ev_df['player'] = ev_df['playerId'].map(players_dict)
-
-# decode team ids
-teams = load_json('teams')
-teams_dict = dict(zip(pd.DataFrame(teams).wyId,pd.DataFrame(teams).name))
-ev_df['team'] = ev_df['teamId'].map(teams_dict)
-
-# unpack location and target
-ev_df['location'] = ev_df['positions'].str[0]
-ev_df['target'] = ev_df['positions'].str[1]
-
-# create simpler dataframe
-e_df = ev_df[['matchId','eventSec','matchPeriod','team','location','target','player','eventName','subEventName','tag0','tag1','tag2','tag3','tag4','tag5']]
-e_df['notes'] = e_df[['tag0','tag1','tag2','tag3','tag4','tag5']].values.tolist()
-e_df['tags'] = e_df['notes'].astype(str)
-e_df['accurate'] = e_df.tags.str.contains('accurate')
-e_df['accurate'] = np.where(e_df.tags.str.contains('not accurate'), 0, e_df.accurate)
-e_df.drop(columns=['tag0','tag1','tag2','tag3','tag4','tag5','notes'],inplace=True)
-
-# label in_play sequences
-play = 0
-plays = []
-next_action = []
-match_play = []
-matchId = e_df.iloc[0].matchId
-for i in range(len(e_df)):
-    if i>1:
-        row_minus2 = row_minus1
-    if i>0:
-        row_minus1 = row
-    row = e_df.iloc[i]
-    if i == 0:
-        row_minus1 = row
-        row_minus2 = row
-
-    if row.eventName == 'Free Kick':
-        # when the ball goes 'out of play', the game resumes with a 'free kick' (here includes Throw-Ins)
-        play += 1
-    elif 'goal' in row_minus2.tags and 'goal' in row_minus1.tags:
-        # (there are generally two events marked goal -- the shot and the attempted save)
-        play +=1
-
-    if row.matchId == matchId:
-        pass
-    else:
-        play = 0
-        matchId = row.matchId
-    plays.append(play)
-    match_play.append(f'{matchId}-{play}')
-
-e_df['in_play'] = pd.Series(plays)
-e_df['match_play'] = pd.Series(match_play)
-
-# label possession sequences
-ps_sequence = np.full(len(e_df),'0000000000000000000000000000000000')
-played_to = np.full(len(e_df), '                                    ')
-to_team = np.full(len(e_df), '                                       ')
-
-for match in e_df.matchId.unique():
-    match_df = e_df.loc[e_df.matchId == match]
-    for team in match_df.team.unique():
-        for play in match_df.in_play.unique():
-            team_play_df = match_df.loc[match_df.in_play == play].loc[match_df.team == team]
-            possession_sequence = 0
-            for i in range(len(team_play_df)):
-                if i>0:
-                    previous_row = row
-
-                row = team_play_df.iloc[i]
-
-                if i > 0:
-                    if row.location != target:
-                        possession_sequence += 1
-                        played_to[previous_row.name] = 'Opposition' 
-                        to_team[previous_row.name] = 'Opposition'
-                    else:
-                        played_to[previous_row.name] = row.player
-                        to_team[previous_row.name] = row.team
-                team_name = team.split(' ')
-                team_abbr = ''
-                for word in team_name:
-                    team_abbr = team_abbr + word[:3]
-                ps_sequence[row.name] = f'{match}-{play}-{team_abbr}-{possession_sequence}'
-                target = row.target
-
-e_df['possession'] = ps_sequence            
-e_df['played_to'] = played_to
-e_df['to_team'] = to_team
-
-e_df['played_to'] = np.where(e_df['played_to'] == '                                    ', 'Opposition', e_df['played_to'])
-e_df['to_team'] = np.where(e_df['to_team'] == '                                       ', 'Opposition', e_df['to_team'])
-
-# unpack x and y coorinates
-for foo in ('location','target'):
-    for goo in ('x','y'):
-        e_df[f'{foo}_{goo}'] = e_df[foo].str[goo]
-
-# change target coordinates of goal to inside goal
-e_df['target_x'] = np.where((e_df.tags.str.contains("'goal'")) & (e_df.eventName != 'Save attempt'),
-                            100, e_df.target_x )
-e_df['target_y'] = np.where((e_df.tags.str.contains("'goal'")) & (e_df.eventName != 'Save attempt'),
-                            50, e_df.target_y )
-
-e_df['subEventName'] = np.where((e_df.tags.str.contains("'goal'")) & (e_df.eventName != 'Save attempt'),
-                            'Goal', e_df.subEventName )
-
-e_df['eventName'] = np.where(e_df['eventName'] == 'Others on the ball', 'On the Ball', e_df['eventName'])
-
-e_df['eventName'] = np.where(e_df['eventName'] == 'Duel', 'Challenge', e_df['eventName'])
-
-e_df['played_to'] = np.where(e_df.subEventName == 'Goal', 'Goal', e_df.played_to)
-e_df['to_team'] = np.where(e_df.subEventName == 'Goal', 'Goal', e_df.to_team)
-
-# change location coordinates of goal kick
-e_df['location_x'] = np.where((e_df.subEventName == 'Goal kick'), 5, e_df.location_x)
-e_df['location_y'] = np.where((e_df.subEventName == 'Goal kick'), 50, e_df.location_y)
-
-# add time in minutes and seconds
-e_df['minute'] = e_df.eventSec//60
-e_df.minute = e_df.minute.astype(int)
-e_df['seconds'] = e_df.eventSec % 60
-e_df.seconds = e_df.seconds.astype(int)
-e_df['time'] = e_df.minute.astype(str).str.zfill(2) + ':' + e_df.seconds.astype(str).str.zfill(2)
-e_df.drop(columns=['minute','seconds'],inplace=True)
-
-e_df.drop(columns=['location','target'], inplace=True)
-e_df.rename(columns={'player':'source','played_to':'target'},inplace=True)
-
-# include starting lineup data
-matches = load_json(f'matches_{national_league}')
-match_df = pd.DataFrame(matches)
-match_lineups = {}
-for match in match_df.wyId.unique():
-    match_lineups[match] = {}
-    for team in match_df.loc[match_df.wyId == match].teamsData.values[0].keys():
-        match_lineups[match][teams_dict[int(team)]] = list(pd.Series(pd.DataFrame(match_df.loc[
-            match_df.wyId == match].teamsData.str[str(team)].str['formation'].str['lineup'].values[0]).playerId.values).map(players_dict))
-
-def firstXI(row):
-    return row.source in match_lineups[row.matchId][row.team]
-
-e_df['FirstXI'] = e_df.apply(firstXI, axis=1)
-
-# include home_or_away data
-h_a_map = {}
-for match in match_df.wyId.unique():
-    h_a_map[match] = {}
-    for team in match_df.loc[match_df.wyId == match].teamsData.values[0].keys():
-        status = match_df.loc[match_df.wyId==match].teamsData.str[str(team)].str['side'].values[0].capitalize()
-        h_a_map[match][team] = status
-
-def home_or_away(row):
-    return h_a_map[row.matchId][str(row.teamId)]
-
-e_df['home_or_away'] = ev_df.apply(home_or_away, axis=1)
-e_df['to_team'] = np.where(e_df.to_team == e_df.team, e_df.home_or_away, e_df.to_team)
-opposition = {'Home':'Away','Away':'Home'}
-e_df['to_team'] = np.where(e_df.to_team == 'Opposition', e_df.home_or_away.map(opposition), e_df.to_team)
-e_df['to_team'] = np.where(e_df.tags.str.contains('own goal'), 'Own Goal', e_df.to_team)
-whose_goal = {'Home':'Home Goal', 'Away':'Away Goal'}
-e_df['to_team'] = np.where(e_df.to_team == 'Goal', e_df.home_or_away.map(whose_goal), e_df.to_team)
-whose_own_goal = {'Home':'Away Goal', 'Away':'Home Goal'}
-e_df['to_team'] = np.where(e_df.to_team == 'Own Goal', e_df.home_or_away.map(whose_own_goal), e_df.to_team)
-
-e_df.drop(columns=['tags','in_play'],inplace=True)
-
-```

### 2020-07-25 rmd/SocialNetworkAnalysis_R.Rmd
done

@@
 TabulateSequence <- function(events, possession_sequence) {
-    # Return a markdown table starting from the row before a possession sequence begins, and continuing until it ends.
+    #' Return table starting from row before possession sequence begins, 
+    #' and continuing until it ends.
 
@@
     table <- events[start:stop,]
-    df <- data.frame(table$matchPeriod, table$time, table$team, table$source, table$subEventName)
-    names(df) <- c('Half', 'Time', 'Team', 'Event Description', 'Player')
+    df <- data.frame(table$matchPeriod, table$time, table$team, 
+                     table$source, table$subEventName, table$possession)
+    names(df) <- c('Half', 'Time', 'Team', 'Player', 
+                   'Event Description', 'Sequence ID' )
 
@@
     d <- read.csv('../tables/MXG.csv')
+
     top <- d[d$Degree>1000,]
@@
                'Goals/Degree','Loops','Loops/Degree','MXG')]
-    write.csv(t[1:20,], '../tables/2.csv',row.names=F)
+
+    t[,'Goals/Degree'] <- formatC(signif(t[,'Goals/Degree']*100,digits=3), 
+                              digits=3,format="fg", flag="#")
+
+    t[,'Goals (% Degree)'] <- paste0(
+                            as.character(t$Goals),
+                            ' (',
+                            as.character(t[,'Goals/Degree']),'%)')
+
+    t$Degree <- as.character(prettyNum(t$Degree, big.mark=',',scientific=F))
+
+    t[,'Loops/Degree'] <- formatC(signif(t[,'Loops/Degree']*100,digits=3), 
+                              digits=3,format="fg", flag="#")
+
+    t[,'Loops (% Degree)'] <- paste0(
+                            as.character(t$Loops),
+                            ' (',
+                            as.character(t[,'Loops/Degree']),'%)')
+
+    t <- t[1:20,c(1,2,8,3,10,9)]
+
+    write.csv(t[1:20,], '../tables/Top20.csv',row.names=F)
 }

### 2020-07-25 rmd/draft.Rmd
done

@@
 Such analysis is of potential interest to coaches and scouts working professionally within the game, journalists and media commentators explaining the game, and gamblers and bookmakers putting money on the game.
-In this paper I explain how Social Network Analysis [@SWassermanEtAl1994] can provide the tools to model a football team's ball-possession as a dynamic probabilistic system, and hence how we can quantify the contribution of each player to the team's goal-scoring capability.
+In this paper I build upon previous work applying Social Network Analysis [@TGrund2012]
+and Markov Modelling [@JPena2014] to football data, and introduce an original metric for quantifying the contribution of each player to a team's goal-scoring capability.
 
@@
 
-The dataset is provided by WyScout, one of a growing number of companies in the football analytics industry, and is available from FigShare.com [@LPappalardoMassucco2019].
+The dataset is provided by WyScout, and is available from FigShare.com [@LPappalardoMassucco2019].
 It includes more than three million *event-logs* for every 2017-18 match from the five major European leagues, in JSON format (see Figure 1), with  the details of players, teams, and informative tags (such as the crucial 'Goal') numerically encoded.
@@
 I used Python to convert this into a *tidy* dataframe [@HWickham2014], which I then analyzed and visualized using R.
-For the sake of reproducible research [@RPeng2011], all code is attached as appendices, as well as available on Github.
+For the sake of reproducible research [@RPeng2011], all code is attached as appendices.
 
@@
 
-### Exploratory Visualization
+### Initial Exploration
 
-We can then visualize the location of the event-log data points in the context of a diagrammatic football pitch, to allow for an immediate and accessible view of the events of any match (see Figures 2 & 3), or particular possession sequence (see Figure 4).
+We begin our exploration of the data by tabulating the frequency of event types (Table 1) and pass types (Table 2), confirming that the *simple pass* is football's primary event.
 
@@
 
+We can then visualize the location of the event-log data points in the context of a diagrammatic football pitch, to allow for an immediate and accessible view of the events of any match (see Figures 2 & 3).
+
 <details open>
@@
 <table><tr><td>
-<img src="../figures/EventMapHome.png">
+<img src="figures/EventMapHome.png">
 </td><td>
-<img src="../figures/EventMapAway.png">
+<img src="figures/EventMapAway.png">
 </td></tr></table>
@@
 <table><tr><td>
-<img src="../figures/PassMapHome.png">
+<img src="figures/PassMapHome.png">
 </td><td>
-<img src="../figures/PassMapAway.png">
+<img src="figures/PassMapAway.png">
 </td></tr></table>
@@
 
+As well as the general overview of a full match's events, we can also consider a particular possession sequence (see Figure 4).
+
 <details open>
@@
 
-<img src="../figures/PossessionSequence.png">
+<img src="figures/PossessionSequence.png">
 <figcaption>Figure 4: Pitchmap showing Possession Sequence</figcaption>
@@
 
-We can compare this to a tabulated account of all events from just before the beginning of the possession sequence until its end, and we see that our algorithm has successfully handled the case where an opposition player's unsuccessful challenge interrupts the sequence of events associated with the team in possession.
+We can compare this to a tabulated account of all events from just before the beginning of the possession sequence until its end, and we see that our algorithm has successfully handled the case where an opposition player's challenge interrupts the sequence of events associated with the team in possession.
 
 <details open>
-<summary>Table 1</summary>
+<summary>Table 3</summary>
 <figure>
@@
 <figcaption>
-Table 1: Events of Possession Sequence
+Table 3: Events of Possession Sequence
 </figcaption>
@@
 
-table1 = pd.read_csv('../tables/1.csv',index_col=0)
-print(table1.to_markdown(showindex=False))
+table = pd.read_csv('../tables/PossessionSequence.csv',index_col=False)
+print(table.to_markdown(showindex=False))
 
@@
 
-<img src="../figures/YouTubeClip.gif" width=800>
+<img src="figures/YouTubeClip.gif" width=800>
 <figcaption>Figure 5: YouTube Clip showing successful conclusion of possession sequence
@@
 
-Social Network Analysis applies the concepts of mathematical *graph theory*.
+Social Network Analysis [@SWassermanFaust1994] applies the concepts of mathematical *graph theory*.
 
@@
 
-On an undirected graph, we call the number of edges connecting a node its *degree*; in a directed graph we distinguish between the *indegree* and *outdegree*.
+On an undirected graph, we call the number of edges connecting a node its *degree*. In a directed graph we distinguish between the *indegree* and *outdegree*.
 
@@
 We consider the possession sequences of a football match as *walks* of the football across a social graph in which the nodes are players (considered either individually or collectively), and the connecting edges are defined by the movement of the ball, as it travels from player to player.
-We can then construct the underlying graph for any particular possession sequence (as in Figure i), or indeed for the whole game.
+We can then construct the underlying graph for any particular possession sequence (as in Figure 6), or indeed for the whole game.
 
@@
 
-<img src="../figures/PossessionGraph.png">
+<img src="figures/PossessionGraph.png">
 <figcaption>Figure 6: Graph of Possession Sequence leading to Goal
@@
 
-There are several different 'weights' we could assign an edge: the Euclidean distance moved by the ball; the polar angle; the time interval.
+There are several different 'weights' we could assign an edge: the Euclidean distance moved by the ball; the polar angle; the time duration.
 However, for this particular analysis we focus simply on the network connections between players.
@@
 
-<img src="../figures/AdjacencyMatrixAway.png">
+<img src="figures/AdjacencyMatrixAway.png">
 
@@
 
-It is increasingly common [@TKnutson2018] to visualize a possession network by positioning each node according to the mean location of the player's touches, and varying the appearance (width, colour or transparency) of the edges to show the frequency of that pass.
+It is increasingly common [@TKnutson2018] to visualize a possession network by positioning each node according to the mean location of the player's touches, and varying the appearance (width, colour or transparency) of the edges to show the frequency of that pass (Figure 8).
 This swiftly communicates a team's formation, and gives a qualitative sense of a team's strategy.
@@
 <table><tr><td>
-<img src="../figures/PassingNetworkHome.png">
+<img src="figures/PassingNetworkHome.png">
 </td><td>
-<img src="../figures/PassingNetworkAway.png">
+<img src="figures/PassingNetworkAway.png">
 </td></tr></table>
@@
 
-To consider the team's ball-possession in its dynamic development, rather than merely as a static structure, we also need some definitions from *probability theory*.
+To consider the team's ball-possession in its dynamic development, rather than merely as a static structure, we also need some definitions from *probability theory* [@RSerfozo2009a].
 
@@
 A stochastic process has the *Markov property* if $\forall t \ge 0, \mathbb{P}(X_{t+1}=s| X_{0},...,X_{t}) = \mathbb{P}(X_{t+1}=s| X_{t})$; this means that future states are dependent only on the present state, regardless of the previous history of past states.
-Such a process can be called a Markov process, and we call the probability $\mathbb{P}(X_{t+1}=s_{j}|X_{t}=s_{i})$ the *transition probability* $p_{ij}$ .
+Such a process can be called a Markov process [@JNorris1998], and we call the probability $\mathbb{P}(X_{t+1}=s_{j}|X_{t}=s_{i})$ the *transition probability* $p_{ij}$ .
 
@@
 
-<img src="../figures/TransitionMatrixAway.png">
+<img src="figures/TransitionMatrixAway.png">
 
@@
 
-### Calculating Markovian Expected Goals
+### Calculating 'Markov Expected Goals'
 
-In Figure 8 we show the heatmap of the Transition Matrix derived from the Adjacency Matrix shown in Figure 6.
+In Figure 9 we show the heatmap of the Transition Matrix derived from the Adjacency Matrix shown in Figure 7.
 There we showed only the adjacency counts for the starting eleven, but now we include the subsitutes, as well as treating the 'Opposition' collectively as a single node.
@@
 Thus we can quantify, for each player on the team, the expected number of events there will be before the team scores a goal.
-We invert this to give a measure of the expected number of goals; and scale it by the team's average number of events per game, we call this measure *Markovian Expected Goals* (MXG).
-We then scale this by the team's mean expected goals values to give a *Relative MXG* value, so that values can be compared fairly across teams to see which individual players contribute the most to their teams' goalscoring chances.
+We invert this to give a measure of the expected number of goals; and standardize it by the team's average number of events per game.
+We then scale this by the team's mean values, so that we can make a fair comparison across teams to see which individual players contribute the most to their teams' goalscoring chances.
+We call this measure *Markov Expected Goals* (MXG).
 
@@
 
-The
+Applying our method to the cumulative season's transition matrix for every team in the dataset, we find that Inter's Icardi is the player who contributed most to the chance of his team scoring (considering only players with degree of at least 1000).
+His MXG score is almost three times higher than that of Messi, who only just makes it into the Top Twenty (Table 4).
+
+It is instructive to consider Messi's score in order to understand what a high MXG score implies.
+Of the twenty, Messi has the highest number of goals, the highest number of loops, and the highest (out-)degree, but the very strength of his degree score means that his percentage of goals per degree (which correlates most strongly with MXG) is the lowest.
 
 <details open>
-<summary>Table 2</summary>
+<summary>Table 4</summary>
 <figure>
@@
 <figcaption>
-Table 2: Top Twenty European Players by MXG Score
+Table 4: Top Twenty Players in European Leagues by MXG Score (2017-18)
 </figcaption>
@@
 
-MXGtable = pd.read_csv('../tables/2.csv',index_col=0)
+MXGtable = pd.read_csv('../tables/Top20.csv',index_col=0)
 print(MXGtable.to_markdown(floatfmt="#.3g"))

### 2020-07-25 scripts/correct_data.py

@@ -1,0 +1,14 @@
+import pandas as pd
+
+# Correct data for De Bruyne's goal vs Chelsea (not an inaccurate pass!)
+
+england_events = pd.read_csv('../data/processed/England.csv')
+
+england_events.at[102730,'eventName'] = 'Shot'
+england_events.at[102730,'subEventName'] = 'Goal'
+england_events.at[102730,'accurate'] = 1
+england_events.at[102730,'to_team'] = 'Away Goal'
+england_events.at[102730,'target'] = 'Goal'
+england_events.at[102730,'target_y'] = 50
+
+england_events.to_csv('../data/processed/England.csv',index=False)

### 2020-07-25 scripts/print_event.py

@@ -1,0 +1,21 @@
+import json
+from random import randint
+
+def load_json(file):
+    "Load raw JSON data."
+    
+    with open(f'../data/raw/{file}.json','r') as f:
+        return json.loads(
+            f.read().replace('\\\\','\\').encode('utf-8')
+            )
+    
+england = load_json('events_England')
+event_num = randint(0,len(england))
+
+with open('../figures/ExampleEvent.json', 'w') as f:
+    txt = '{\n'
+    event = england[event_num]
+    for key in event.keys():
+        txt = f"{txt}\t'{key}' : {event[key]},\n"
+    txt = txt + "}"
+    f.write(txt)

### 2020-07-26 rmd/draft.Rmd
as handed in

@@
 Such analysis is of potential interest to coaches and scouts working professionally within the game, journalists and media commentators explaining the game, and gamblers and bookmakers putting money on the game.
-In this paper I build upon previous work applying Social Network Analysis [@TGrund2012]
-and Markov Modelling [@JPena2014] to football data, and introduce an original metric for quantifying the contribution of each player to a team's goal-scoring capability.
+In this paper I build upon previous work applying Social Network Analysis [@TGrund2012] and Markov Modelling [@JPena2014] to football data, and introduce an original method for quantifying the contribution of each player to a team's goal-scoring capability.
 
@@
 I then show how the adjacency matrix of a player, considered as a node in a football match's possession sequence graph, can be used to quantify his contribution to the likelihood of the team scoring, which I call *Markov Expected Goals*.
-Finally, I discuss the potential and pitfalls of this new metric.
+Finally, I discuss the potential and pitfalls of this new indicator.
 
@@
 
-The dataset is provided by WyScout, and is available from FigShare.com [@LPappalardoMassucco2019].
+The dataset is provided by WyScout, and is available from [FigShare.com](https://figshare.com/collections/Soccer_match_event_dataset/4415000) [@LPappalardoEtAl2019a].
 It includes more than three million *event-logs* for every 2017-18 match from the five major European leagues, in JSON format (see Figure 1), with  the details of players, teams, and informative tags (such as the crucial 'Goal') numerically encoded.
@@
 As part of this preprocessing stage I also labelled each event according first to its *in-play sequence*, and then its *possession sequence*.
-An in-play sequence ends when the ball goes out of play, as indicated by the subsequent event having the *eventName* 'Free Kick' (a category which includes throw-ins and corners, as well as actual free-kicks given for fouls or offsides); or by the final two events (the shot and the attempted save) being labelled with the *tag* 'Goal'.
-(Although every goal is followed by a 'free kick' from the centre of the pitch, this never seems to be included in the event logs, presumably because they are created from camera footage which focusses on the goal celebrations and tends to miss the immediate restart of play after a goal).
 
@@
 However, since the event-logs include challenges for the ball even when they are unsuccessful, this would lead to systematically undercounting the length of possession sequences.
+
 The solution I adopt is to iterate through the event-logs for a single team within an in-play sequence, paying close attention to the location data: if the location of an event is identical to the target-location of the previous event associated with that same team, then I treat is as part of the same possession sequence, otherwise I treat it as a separate sequence.
@@
 
-We can then visualize the location of the event-log data points in the context of a diagrammatic football pitch, to allow for an immediate and accessible view of the events of any match (see Figures 2 & 3).
+We can then visualize the event-log data as points (Figure 2) or trajectories (Figure 3) in the context of a diagrammatic football pitch, to allow for an immediately accessible overview of any match.
 
@@
 
-With possession sequences that end in a goal, it is generally possible to find video highlights of the events under consideration, and confirm that the data we have does correspond to what actually happened.
-
-<details open>
-<summary>Figure 5</summary>
-<figure>
-----
-
-<img src="figures/YouTubeClip.gif" width=800>
-<figcaption>Figure 5: YouTube Clip showing successful conclusion of possession sequence
-</figcaption>
-
-----
-
-</figure>
-</details>
-
 ## Methodology
@@
 We consider the possession sequences of a football match as *walks* of the football across a social graph in which the nodes are players (considered either individually or collectively), and the connecting edges are defined by the movement of the ball, as it travels from player to player.
-We can then construct the underlying graph for any particular possession sequence (as in Figure 6), or indeed for the whole game.
+We can then construct the underlying graph for any particular possession sequence (as in Figure 5), or indeed for the whole game.
 
 <details open>
-<summary>Figure 6</summary>
+<summary>Figure 5</summary>
 <figure>
@@
 <img src="figures/PossessionGraph.png">
-<figcaption>Figure 6: Graph of Possession Sequence leading to Goal
+<figcaption>Figure 5: Graph of Possession Sequence leading to Goal
 </figcaption>
@@
 <details open>
-<summary>Figure 7</summary>
+<summary>Figure 6</summary>
 <figure>
@@
 <figcaption>
-Figure 7: Heatmap of Adjacency Matrix
+Figure 6: Heatmap of Adjacency Matrix
 </figcaption>
@@
 
-It is increasingly common [@TKnutson2018] to visualize a possession network by positioning each node according to the mean location of the player's touches, and varying the appearance (width, colour or transparency) of the edges to show the frequency of that pass (Figure 8).
+It is increasingly common [@TKnutson2018] to visualize a possession network by positioning each node according to the mean location of the player's touches, and varying the appearance (width, colour or transparency) of the edges to show the frequency of that pass (Figure 7).
 This swiftly communicates a team's formation, and gives a qualitative sense of a team's strategy.
@@
 <details open>
-<summary>Figure 8</summary>
+<summary>Figure 7</summary>
 <figure>
@@
 <figcaption>
-Figure 8: Possession Networks for Home and Away Team
+Figure 7: Possession Networks for Home and Away Team
 </figcaption>
@@
 <details open>
-<summary>Figure 9</summary>
+<summary>Figure 8</summary>
 <figure>
@@
 <figcaption>
-Figure 9: Heatmap of Transition Matrix
+Figure 8: Heatmap of Transition Matrix
 </figcaption>
@@
 
-In Figure 9 we show the heatmap of the Transition Matrix derived from the Adjacency Matrix shown in Figure 7.
+In Figure 8 we show the heatmap of the Transition Matrix derived from the Adjacency Matrix shown in Figure 6.
 There we showed only the adjacency counts for the starting eleven, but now we include the subsitutes, as well as treating the 'Opposition' collectively as a single node.
